### 2022-03-27 paper.tex
init

%% ****** Start of file apstemplate.tex ****** %
%%
%%
%%   This file is part of the APS files in the REVTeX 4.2 distribution.
%%   Version 4.2a of REVTeX, January, 2015
%%
%%
%%   Copyright (c) 2015 The American Physical Society.
%%
%%   See the REVTeX 4 README file for restrictions and more information.
%%
%
% This is a template for producing manuscripts for use with REVTEX 4.2
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.
%
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showkeys' option to make keywords appear
%\documentclass[aps,prb,twocolumn,superscriptaddress]{revtex4-1}
\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-2}
%\documentclass[aps,prl,reprint,groupedaddress]{revtex4-2}

% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.
%\bibliographystyle{apsrev4-2}
\usepackage{graphicx}
\usepackage{color}
\frenchspacing
%\graphicspath{{~/data/KannanLu/exp623/article/}}

\begin{document}

% Use the \preprint command to place your local institutional report
% number in the upper righthand corner of the title page in preprint mode.
% Multiple \preprint commands are allowed.
% Use the 'preprintnumbers' class option to override journal defaults
% to display numbers if necessary
%\preprint{}

%Title of paper
\title{Independent Component Analysis}

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current
% author. Explanatory text should go in the []'s, actual e-mail
% address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro foreach each type of information

% \affiliation command applies to all authors since the last
% \affiliation command. The \affiliation command should follow the
% other information
% \affiliation can be followed by \email, \homepage, \thanks as well.

%\email[]{Your e-mail address}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\author{K. Krongchon}
\author{K. Lu}
\affiliation{Department of Physics, University of Illinois at Urbana-Champaign, Urbana, IL 61801, USA}

%Collaboration name if desired (requires use of superscriptaddress
%option in \documentclass). \noaffiliation is required (may also be
%used with the \author command).
%\collaboration can be followed by \email, \homepage, \thanks as well.
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}

\end{abstract}

% insert suggested keywords - APS authors don't need to do this
%\keywords{}

%\maketitle must follow title, authors, abstract, and keywords
\maketitle

% body of paper here - Use proper section commands
% References should be done using the \cite, \ref, and \label commands
\section{Introduction}

% Put \label in argument of \section for cross-refe(WHAT DOES THIS LAST SENTENCE MEAN?)rencing
%\section{\label{}}
%\subsection{}
%\subsubsection{}

\section{Mathematical Formulation}

\section{Algorithms and Applications}

\section{Discussions and Conclusions}

\begin{acknowledgments}

\end{acknowledgments}

% If in two-column mode, this environment will change to single-column
% format so that long equations can be displayed. Use
% sparingly.
%\begin{widetext}
% put long equation here
%\end{widetext}

% figures should be put into the text as floats.
% Use the graphics or graphicx packages (distributed with LaTeX2e)
% and the \includegraphics macro defined in those packages.
% See the LaTeX Graphics Companion by Michel Goosens, Sebastian Rahtz,
% and Frank Mittelbach for instance.
%
% Here is an example of the general form of a figure:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Use the figure* environment if the figure should span across the
% entire page. There is no need to do explicit centering.

% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}

% Surround figure environment with turnpage environment for landscape
% figure
% \begin{turnpage}
% \begin{figure}
% \includegraphics{}%
% \caption{\label{}}
% \end{figure}
% \end{turnpage}

% tables should appear as floats within the text
%
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
% The ruledtabular enviroment adds doubled rules to table and sets a
% reasonable default table settings.
% Use the table* environment to get a full-width table in two-column
% Add \usepackage{longtable} and the longtable (or longtable*}
% environment for nicely formatted long tables. Or use the the [H]
% placement option to break a long table (with less control than
% in longtable).
% \begin{table}%[H] add [H] placement to break table across pages
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% Lines of table here ending with \\
% \end{tabular}
% \end{ruledtabular}
% \end{table}

% Surround table environment with turnpage environment for landscape
% table
% \begin{turnpage}
% \begin{table}
% \caption{\label{}}
% \begin{ruledtabular}
% \begin{tabular}{}
% \end{tabular}
% \end{ruledtabular}
% \end{table}
% \end{turnpage}

% Specify following sections are appendices. Use \appendix* if there
% only one appendix.
%\appendix
%\section{}

% If you have acknowledgments, this puts in the proper section head.
%\begin{acknowledgments}
% put your acknowledgments here.
%\end{acknowledgments}

% Create the reference section using BibTeX:
%\bibliography{****.bib}

\end{document}
%
% ****** End of file apstemplate.tex ******

\grid
\grid
\grid

### 2022-04-17 scripts/test.py
add script

@@ -1,0 +1,87 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import numpy.linalg as la
+
+def exmaple_1():
+
+    def get_s1(j):
+        return 0.5*np.sin(j)
+
+    def get_s2(N):
+        return np.random.uniform(0, 1, N) - 0.5
+
+    def get_x(s1, s2, a, b):
+        return a*s1 + b*ss
+
+    N = 500
+    lin = np.linspace(1, 30, N)
+    s1 = get_s1(lin)
+    s2 = get_s2(N)
+    x1 = get_x(s1, s2, 0.5, 0.5)
+    x2 = get_x(s1, s2, 0.5, -0.5)
+    # plt.plot(lin, x2)
+    plt.plot(x1, x2, 'o')
+    plt.show()
+
+def example_2_1_2():
+    s_1 = [1, 2, 1, 2]
+    s_2 = [1, 1, 2, 2]
+    s = np.array([s_1, s_2])
+    A = np.array([
+        [1, 2],
+        [1, -1]
+        ])
+    x = A @ s
+    W = la.inv(A)
+    print(W @ x)
+
+def get_cov(A):
+    return (A.T @ A)/(A.shape[0] - 1)
+
+def example_3_3():
+    X = np.array([
+        [1, 1, 2, 0, 5, 4, 5, 3],
+        [3, 2, 3, 3, 4, 5, 5, 4]
+        ]).T
+    mu = X.mean(axis=0)
+    D = X - mu
+    sigma_D = get_cov(D)
+    lamda, V = la.eigh(sigma_D)
+    U = (V @ D.T).T
+    sigma_U = get_cov(U)
+    print(sigma_U)
+
+    Z = np.einsum('i,ij -> ij', lamda**-0.5, U.T).T
+    print(lamda.shape)
+    print(U.shape)
+    print(Z.T)
+    sigma_Z = get_cov(Z)
+    print(sigma_Z)
+
+    # plt.plot(X[:, 0], X[:, 1], 'o')
+    # plt.plot(U[:, 0], U[:, 1], 'o')
+    plt.plot(Z[:, 0], Z[:, 1], 'o')
+    plt.show()
+
+def pca():
+
+    def get_s1(N):
+        lin = np.linspace(1, 30, N)
+        return 0.5*np.sin(lin)
+
+    def get_s2(N):
+        return np.random.uniform(0, 1, N) - 0.5
+
+    N = 100
[…]

### 2022-04-17 paper.tex
separate sections
sections content summary

@@
 \section{Introduction}
+Why ICA: 1) In reality, source signals are often corrupted with noise, or in other words, data are composed by mutiple independent source signals.
+To isolate the signals from different sources are called 'Blind Source Separation' (BSS).
+Famous example is 'Cocktail party problem'.
+ICA is widely used in images, sounds, stock market and medicine, can be also viewed as a dimension reduction technique (or filtering operation).
+ICA is an extension of PCA, where the latter pertains to the second order statistics (covariance matrix of data).
+ICA optmizes higher-order statistics.
+
+ALgorithms: FastICA, projection pursuit and Infomax. Drawbacks: over-complete or under-complete.
 
@@
 \section{Mathematical Formulation}
+1) Optimizing Kurtosis
+2) Minimizing mutual information
+3) Maximum likelihood estimation
 
 \section{Algorithms and Applications}
+1) projection pursuit (kurtotis)
+2) FastICA (minimizing mutual information
+
+3) example: sound single decomposition
 
 \section{Discussions and Conclusions}
+1) review the traditional development from 1990s
+2) recent advances and development from 2000s
 

### 2022-04-17 paper.tex
more section summaries

@@
 \section{Mathematical Formulation}
-1) Optimizing Kurtosis
-2) Minimizing mutual information
-3) Maximum likelihood estimation
+0) Proof that gaussian signals cannot be separated (Kannan)
+1) Optimizing Kurtosis (Kannan)
+2) Minimizing mutual information (Mick)
+3) Maximum likelihood estimation (Mick)
 
 \section{Algorithms and Applications}
-1) projection pursuit (kurtotis)
-2) FastICA (minimizing mutual information
+1) projection pursuit (kurtotis) (Kannan)
+2) FastICA (minimizing mutual information (Mick)
 

### 2022-04-22 paper.tex
more section summaries
edited the paper.tex

@@
 \usepackage{graphicx}
+\usepackage{amsfonts}
+\usepackage{bm}
 \usepackage{color}
@@
 0) Proof that gaussian signals cannot be separated (Kannan)
-1) Optimizing Kurtosis (Kannan)
-2) Minimizing mutual information (Mick)
+//
+
+Before we delve into the detailed mathematical formulations, we first define the notations that we will use throughout the whole document.
+We will use boldface capital letters $\bf{A}$ for matrices, boldface lowercase letters $\bf{a}$ for vectors and ordinary lower case letters $a$ for scalars.
+We will use subscripts for denoting the components of matrices (i.e.
+$A_{ij})$ and vectors (i.e.
+$a_{i}$).
+Notice that these letters are not in boldface as they mean the specific component, which is just a scalar in $ \rm I\!R$.
+Superscripts are reserved for indexing different samples.
+For instance, $m$ samples of vector $\bf{a}$ can be denoted as $\bf{a}$$^{(k)}$ where $k = 1, 2, ..., m$.
+Following these conventions, we denote the original signal to be $\bf{s}$$^{(i)} \in \rm I\!R^{d}$.
+We consider the standard setting where the $d$ dimensional signals $\bf{s}$$^{(i)} \in \rm I\!R^{d}$ ($d$ original sources) are mixed and observed by $d$ detectors $\bf{x}$$^{(i)} \in \rm I\!R^{d}$ (i.e.
+the cocktail party scenario).
+That is there is a linear map between the $\bf{s}$$^{(i)}$ and $\bf{x}$$^{(i)}$, $\bf{x}$$^{(i)}$ $=$ $\bf{A}$$\bf{s}$$^{(i)}$, where $\bf{A} \in \rm I\!R^{d\times d}$ is known as the mixing matrix.
+The whole idea of ICA is to learn the inverse of the mixing matrix $\bf{W} = \bf{A}$$^{-1}$ from the observed data $\bf{x}$$^{(i)}$, known as the un-mixing.
+This can be done in several different ways based on ideas of seperating non-Gaussian signals, and the original signals being independent.
+We will in this section discuss various mathematical formulations.
+
+//
+
+First of all, we need to talk about some ambiguities embedded in the symmetry of the problem and justify that if the origianl signals are all Gaussians, they cannot be learned through the un-mixing.
+
+There are two ambiguities that usually do not affect the practical application, namely the permutation ambiguity and the scaling ambiguity.
+It is trivial to see that the ordering of the original sources \{$s_{j}$\} is ambiguous.
+In the case of scaling, if the original signal $\bf{s}$ is scaled by a non-zero constant to be $c\bf{s}$ where $c \neq 0$.
+Then the mixing matrix $\bf{A}$ can be scaled by $1/c$ and resulting in the same observed data $\bf{x}$$ = \frac{1}{c}\bf{A}$$c\bf{s}$.
+This scaling ambiguity can be furtherly extended with regard to each component of the original signal.
+That is, for a particular component $j$, if we scale the component by $c_{j}$ the corrsponding column of the mixing matrix can be scaled by $1/c_{j}$ to have the observed data unchanged (i.e.
+$x_{i} = \sum_{j}\frac{1}{c_{j}}A_{ij}c_{j}s_{j}$).
+
+The other ambiguity that matters for the practical application is that the original sources cannot all be distributed as Gaussians.
+Or to be more precise, in order to seperate the independent components, we require that the original signals can only have at most one component to be sampled from a Gaussian distribution, i.e.
+$s_{j}$ $\sim \mathcal{N}(\mu,\,\sigma^{2})$ for at most one $j$.
+Let's consider the case when all the original independent signals are Gaussians.
+That is, $s_{j}$ $\sim \mathcal{N}(\mu_{j},\,\sigma_{j}^{2})$ for $j = 1, 2, ..., d$.
+Then given the mixing matrix $\bf{A}$, we have
+\begin{equation}
+\mathbb{E}_{\bf{s}}[\bf{x}] = \bf{A}\mathbb{E}[\bf{s}] = \bf{A}\bm{\mu}
+\end{equation}
+ and 
+\begin{equation}
+Cov[\bf{x}] = \mathbb{E}_{\bf{s}}[\bf{As}\bf{s}^t\bf{A}^{t}]-\mathbb{E}_{\bf{s}}[\bf{As}]\mathbb{E}[(\bf{As})^{t}] =  \bf{A}\bm{\Sigma}\bf{A}^{t}-\bf{A}\bf{M}\bf{A}^{t}
+\end{equation} 
+, where $\bf{M}$ and $\bm{\Sigma}$ are diagonal matrices of $\rm I\!R^{d\times d}$ with diagonal elements to be \{$\mu_{j}^{2}$\} and \{$\sigma_{j}^{2}$\}, respectively.
+Due to the scale ambiguity, this is the same as considering normally distributed signals with unity variance.
+Since each column $j$ of the mixing matrix $\bf{A}$ can be scaled by $\sigma_{j}$, $\tilde{\bf{A}}_{:j} = \sigma_{j}\bf{A}_{:j}$ and correspondingly the signal needs to be scaled by $1/\sigma_{j}$, $\tilde{s_{j}} = \frac{s_{j}-\mu_{j}}{\sigma_{j}}$ (known as the whitening).
+Then,
+\begin{equation}
+Cov[\bf{x}] = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t}
+\end{equation}.
+Now, if we consider a rotation $\bf{R} $$\in O(d)$ acting on the scaled (whitened) source signals $\bf{\tilde{s}}$.
+The mixing matrix then changes to $\tilde{\bf{A}}\bf{R}$.
+Unpon this rotation, we would observe $\bf{x'}$ as $\bf{x'} = \tilde{\bf{A}}\bf{R}\tilde{\bf{s}}$.
+$\bf{x'}$ is again normally distributed and has covariance matrix,
+\begin{equation}
+Cov[\bf{x'}] = \tilde{\bf{A}}\bf{R}\bf{I}_{d\times d}\bf{R}^{t}\tilde{\bf{A}}^{t} = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t} 
+\end{equation}.
+This simply means that whether the sources are rotated or not the observed data will be distributed as $\mathcal{N}(0, \tilde{\bf{A}}\tilde{\bf{A}}^{t})$.
+Thus, because of the rotational symmetry of the multivariate Gaussian distribution, we cannot seperate and obtain the original source signals.
+
+//
+
+1) Optimizing Kurtosis (high-order based moment approach) (Kannan)
+
+The non-Gaussianity thus hints the methodologies used in finding the independent component.
+Notice that the PCA only finds the uncorrelated components but not necessarily independent.
+The ICA tries to find independent components where the joint distibutions can be factorized into marginal distributions by investigating into higher-order moments.
+There are multiple moment-based objective functions have been used along the development of ICA.
+We will define some of the notations and stick to one of the moment-based objective functions in this paper.
+
+As we discussed previously, we have the freedom to whiten the observed signals and we will denote the whitened observed data as $\tilde{\bf{x}}$, where each component is $\tilde{x_{j}} = \frac{x_{j}-\mathbb{E}[x_{j}]}{\sigma_{x,j}}$.
+The skewness, kurtosis and excess kutosis are defined as $$ \gamma(x) := \mathbb{E}[\tilde{x}^{3}], \beta(x) := \mathbb{E}[\tilde{x}^{4}], \kappa(x) := \beta(x)-3 $$.
+The objective function is a convex combination of the squared third and fourth cumulants defined as follows,
+\begin{equation}
+L_{\alpha}(\bf{u}) := \alpha \gamma^{2}(\bf{u}^{t}\tilde{\bf{x}}) + (1-\alpha)\kappa^{2}(\bf{u}^{t}\tilde{\bf{x}})
+\end{equation},
+where $\alpha \in [0, 1]$ and $\bf{u}^{t}\bf{u}$$ = 1$.
+
+2) Minimizing mutual information (min KL divergence) (Mick)
 3) Maximum likelihood estimation (Mick)
@@
 3) example: sound single decomposition
+4) tock price
+5) EGG
 
@@
 2) recent advances and development from 2000s  
+3) what happens when there are multiple Gaussian components
 

### 2022-04-24 paper.tex

@@
 \usepackage{graphicx}
+\usepackage{amsmath,amsthm,amssymb,mathtools}
 \usepackage{amsfonts}
@@
 
-\section{Algorithms and Applications}
+\section{Algorithms}
 1) projection pursuit (kurtotis) (Kannan)
-2) FastICA (minimizing mutual information (Mick)
 
+\subsection{FastICA}
+In this method, we maximize the non-Gaussianity given by
+\begin{align}
+J(Y) \propto \{E[G(Y)] - E[G(V)]\}^2,
+\end{align}
+where $Y$ is a random variable of zero mean and unit variance, which can be achieved by whitening the data, and $V$ is a Gaussian variable also of zero mean and unit variance.
+In function forms of $G_1$ and $G_2$ should not grow too fast to be robust.
+The following choice of $G$ is proposed by Hyvärinen and Oja.
+\begin{align}
+G_1(u) &= \frac{1}{a_1} \log (\cosh a_1 u). \\
+G_1^{\prime}(u) &= \tanh(a_1 u).
+\end{align}
+We want to find the extrema of $E[G(Y)]$ to maximize $J(Y)$ under the constraints
+\begin{align}
+E[(w^T x)^2] = \| w\|^2 = 1.
+\end{align}
+From the Kuhn--Tucker conditions, the extremum condition is satisfied when
+\begin{align}
+E[x g(w^Tx)] - \beta w &= 0.
+\end{align}
+This equation can be solved by using Newton's method.
+Let $F(w)$ denote the left-hand side of the equation, which we are trying to solve.
+We update the value of $w$ in each iteration according to the following equation.
+\begin{align}
+w_{n+1} &= w_n - \frac{F(w_n)}{F^{\prime}(w_n)} \\
+&= w_n - \frac{E[x g (w_n^T x)] - \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
+&= \frac{w_n E[g^{\prime}(w_n^T x)] - \beta w_n - E[x g (w_n^T x)] + \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
+&= \frac{w_n E[g^{\prime}(w_n^T x)] - E[x g (w_n^T x)]}{E[g^{\prime}(w_n^T x)] - \beta}.
+\end{align}
+
+\section{Applications}
 3) example: sound single decomposition

### 2022-04-24 paper.tex
update outline

@@
 
-1) Optimizing Kurtosis (high-order based moment approach) (Kannan)
-
 The non-Gaussianity thus hints the methodologies used in finding the independent component.
@@
 
-2) Minimizing mutual information (min KL divergence) (Mick)
-3) Maximum likelihood estimation (Mick)
+1) Optimizing Kurtosis (high-order based moment approach) (Mick)
+2) non-gaussianity  (Mick)
+3) Maximum likelihood estimation (Kannan)
+    - Minimizing mutual information (min KL divergence)
 
 \section{Algorithms}
-1) projection pursuit (kurtotis) (Kannan)
+\subsection{objective function based on maximum likelihood  (Kannan)}
 
 \subsection{FastICA}
+\subsection{alternative objective function, skewness and kurtosis (Kannan)}
 In this method, we maximize the non-Gaussianity given by

### 2022-04-28 scripts/fast_ica.py
add fast_ica

@@ -1,0 +1,133 @@
+import numpy as np
+import numpy.linalg as la
+from scipy import signal
+from scipy.io import wavfile
+from matplotlib import pyplot as plt
+
+def data1():
+    N = 2000
+    time = np.linspace(0, 8, N)
+    s1 = np.sin(2*time) # sinusoidal
+    s2 = np.sign(np.sin(3*time)) # square signal
+    s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
+
+    S = np.array([s1, s2, s3])
+    A = np.array([
+        [1, 1, 1],
+        [0.5, 2, 1],
+        [1.5, 1, 2]
+        ])
+    X = A @ S
+    return X, S
+
+def data2():
+    X = np.array([
+        [1, 1, 2, 0, 5, 4, 5, 3],
+        [3, 2, 3, 3, 4, 5, 5, 4]
+        ])
+    return X
+
+def get_cov(X):
+    return (X @ X.T)/(X.shape[1] - 1)
+
+def center(X, standardize=False):
+    mu = X.mean(axis=1, keepdims=True)
+    sd = X.std(axis=1, keepdims=True, ddof=0) if standardize else 1
+    D = (X - mu)/sd
+    return D
+
+def whiten(X):
+    cov = np.cov(X)
+    lamda, V = la.eigh(cov)
+    lamda_inv = np.sqrt(la.inv(np.diag(lamda)))
+    Z = lamda_inv @ V.T @ X
+    print(np.cov(Z))
+    return Z
+
+def g(x, a1=1):
+    return np.tanh(a1*x)
+
+def dg(x):
+    return 1 - g(x)**2
+
+def test_gram_schmidt(w, W, i):
+    w_new1 = w.copy()
+    for j in range(i):
+        w_new1 -= (w_new1 @ W[j].T) * W[j]
+    w_new2 = w.copy()
+    w_new2 -= W[:i] @ w_new2 @ W[:i]
+    assert np.allclose(w_new1, w_new2)
+
+def ica(X, iterations=1000, tol=1e-5):
+    X = whiten(center(X))
+    nrows = X.shape[0]
+    W = np.zeros((nrows, nrows))
+    for i in range(nrows):
+        w = np.random.random((nrows))
+        for j in range(iterations):
+            w_new = (X * g(w.T @ X)).mean(axis=1) - dg(w.T @ X).mean() * w
+            w_new -= W[:i] @ w_new @ W[:i]
[…]

### 2022-04-29 scripts/fast_ica.py
add tests

@@ -1,10 +1,15 @@
+from matplotlib import pyplot as plt
 import numpy as np
 import numpy.linalg as la
 from scipy import signal
 from scipy.io import wavfile
-from matplotlib import pyplot as plt
 
-def data1():
+import tests
+
+plt.rc('font', family='serif')
+plt.rc('text', usetex=True)
+
+def load_data():
     N = 2000
     time = np.linspace(0, 8, N)
     s1 = np.sin(2*time) # sinusoidal
@@ -20,28 +25,20 @@ def data1():
     X = A @ S
     return X, S
 
-def data2():
-    X = np.array([
-        [1, 1, 2, 0, 5, 4, 5, 3],
-        [3, 2, 3, 3, 4, 5, 5, 4]
-        ])
-    return X
-
-def get_cov(X):
-    return (X @ X.T)/(X.shape[1] - 1)
-
 def center(X, standardize=False):
     mu = X.mean(axis=1, keepdims=True)
     sd = X.std(axis=1, keepdims=True, ddof=0) if standardize else 1
     D = (X - mu)/sd
     return D
 
-def whiten(X):
+def whiten(X, test=True):
     cov = np.cov(X)
     lamda, V = la.eigh(cov)
     lamda_inv = np.sqrt(la.inv(np.diag(lamda)))
     Z = lamda_inv @ V.T @ X
-    print(np.cov(Z))
+
+    if test:
+        tests.test_identity(np.cov(Z))
     return Z
 
 def g(x, a1=1):
@@ -50,32 +47,29 @@ def g(x, a1=1):
 def dg(x):
     return 1 - g(x)**2
 
-def test_gram_schmidt(w, W, i):
-    w_new1 = w.copy()
-    for j in range(i):
-        w_new1 -= (w_new1 @ W[j].T) * W[j]
-    w_new2 = w.copy()
-    w_new2 -= W[:i] @ w_new2 @ W[:i]
-    assert np.allclose(w_new1, w_new2)
-
-def ica(X, iterations=1000, tol=1e-5):
+def ica(X, cycles=1000, tol=1e-5, test=True):
     X = whiten(center(X))
     nrows = X.shape[0]
     W = np.zeros((nrows, nrows))
+    distances = []
     for i in range(nrows):
         w = np.random.random((nrows))
-        for j in range(iterations):
+        for _ in range(cycles):
             w_new = (X * g(w.T @ X)).mean(axis=1) - dg(w.T @ X).mean() * w
+
[…]

### 2022-04-29 scripts/old.py

@@ -91,3 +91,14 @@ def pca():
 
 
 # pca()
+
+def data2():
+    X = np.array([
+        [1, 1, 2, 0, 5, 4, 5, 3],
+        [3, 2, 3, 3, 4, 5, 5, 4]
+        ])
+    return X
+
+def get_cov(X):
+    return (X @ X.T)/(X.shape[1] - 1)
+

### 2022-04-29 scripts/tests.py

@@ -1,0 +1,12 @@
+import numpy as np
+
+def test_gram_schmidt(w, W, i):
+    w_new1 = w.copy()
+    for j in range(i):
+        w_new1 -= (w_new1 @ W[j].T) * W[j]
+    w_new2 = w.copy()
+    w_new2 -= W[:i] @ w_new2 @ W[:i]
+    assert np.allclose(w_new1, w_new2)
+
+def test_identity(A):
+    assert np.allclose(A, np.eye(A.shape[0]))

### 2022-04-29 scripts/ex1.py
add sklearn

@@ -1,0 +1,60 @@
+from matplotlib import pyplot as plt
+import numpy as np
+from scipy import signal
+from scipy.io import wavfile
+from sklearn.decomposition import FastICA, PCA
+
+import fast_ica
+
+plt.rc('font', family='serif')
+plt.rc('text', usetex=True)
+
+def load_data():
+    N = 2000
+    time = np.linspace(0, 8, N)
+    s1 = np.sin(2*time) # sinusoidal
+    s2 = np.sign(np.sin(3*time)) # square signal
+    s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
+
+    S = np.array([s1, s2, s3])
+    A = np.array([
+        [1, 1, 1],
+        [0.5, 2, 1],
+        [1.5, 1, 2]
+        ])
+    X = A @ S
+    return X, S
+
+def plot_sources(X, S, S_predicted, S_sklearn, S_pca):
+    fig, ax = plt.subplots(nrows=2, ncols=3)
+
+    for x in X:
+        ax[0, 0].plot(x)
+        ax[0, 0].set_title('Mixture signals ($X$)')
+
+    for s in S:
+        ax[0, 1].plot(s)
+        ax[0, 1].set_title("Original sources ($S$)")
+
+    for S_predicted in S_predicted:
+        ax[1, 0].plot(S_predicted)
+        ax[1, 0].set_title("Predicted sources ($W X$) [Our FastICA]")
+
+    for S_sklearn in S_sklearn:
+        ax[1, 1].plot(S_sklearn)
+        ax[1, 1].set_title("Predicted sources ($W X$) [Sklearn FastICA]")
+
+    for S_pca in S_pca:
+        ax[1, 2].plot(S_pca)
+        ax[1, 2].set_title("Predicted sources ($W X$) [Sklearn PCA]")
+
+    fig.tight_layout()
+    plt.show()
+
+if __name__ == '__main__':
+    np.random.seed(0)
+    X, S = load_data()
+    S_predicted, distances = fast_ica.ica(X, cycles=100)
+    S_sklearn = FastICA(n_components=3).fit_transform(X.T).T
+    S_pca = PCA(n_components=3).fit_transform(X.T).T
+    plot_sources(X, S, S_predicted, S_sklearn, S_pca)

### 2022-04-29 scripts/fast_ica.py

@@ -1,29 +1,7 @@
-from matplotlib import pyplot as plt
 import numpy as np
 import numpy.linalg as la
-from scipy import signal
-from scipy.io import wavfile
 
 import tests
-
-plt.rc('font', family='serif')
-plt.rc('text', usetex=True)
-
-def load_data():
-    N = 2000
-    time = np.linspace(0, 8, N)
-    s1 = np.sin(2*time) # sinusoidal
-    s2 = np.sign(np.sin(3*time)) # square signal
-    s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
-
-    S = np.array([s1, s2, s3])
-    A = np.array([
-        [1, 1, 1],
-        [0.5, 2, 1],
-        [1.5, 1, 2]
-        ])
-    X = A @ S
-    return X, S
 
 def center(X, standardize=False):
     mu = X.mean(axis=1, keepdims=True)
@@ -39,6 +17,7 @@ def whiten(X, test=True):
 
     if test:
         tests.test_identity(np.cov(Z))
+
     return Z
 
 def g(x, a1=1):
@@ -47,7 +26,7 @@ def g(x, a1=1):
 def dg(x):
     return 1 - g(x)**2
 
-def ica(X, cycles=1000, tol=1e-5, test=True):
+def ica(X, cycles=1000, tol=1e-5, test=False):
     X = whiten(center(X))
     nrows = X.shape[0]
     W = np.zeros((nrows, nrows))
@@ -68,58 +47,5 @@ def ica(X, cycles=1000, tol=1e-5, test=True):
             if distance < tol:
                 break
         W[i, :] = w
-    S = W @ X
-    return S, distances
-
-def plot_sources(X, S, S_predicted):
-    fig, ax = plt.subplots(nrows=2, ncols=2)
-
-    for x in X:
-        ax[0, 0].plot(x)
-        ax[0, 0].set_title('mixture signals')
-
-    for s in S:
-        ax[0, 1].plot(s)
-        ax[0, 1].set_title("original sources")
-
-    for S_predicted in S_predicted:
-        ax[1, 0].plot(S_predicted)
-        ax[1, 0].set_title("predicted sources")
-    fig.tight_layout()
-    plt.show()
-
-# def mix_sources(mixtures, apply_noise=False):
-#     for i in range(len(mixtures)):
-#         max_val = np.max(mixtures[i])
-#         if max_val > 1 or np.min(mixtures[i]) < 1:
[…]

### 2022-04-29 scripts/archive/old.py
archive

@@ -1,0 +1,104 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import numpy.linalg as la
+
+def exmaple_1():
+
+    def get_s1(j):
+        return 0.5*np.sin(j)
+
+    def get_s2(N):
+        return np.random.uniform(0, 1, N) - 0.5
+
+    def get_x(s1, s2, a, b):
+        return a*s1 + b*ss
+
+    N = 500
+    lin = np.linspace(1, 30, N)
+    s1 = get_s1(lin)
+    s2 = get_s2(N)
+    x1 = get_x(s1, s2, 0.5, 0.5)
+    x2 = get_x(s1, s2, 0.5, -0.5)
+    # plt.plot(lin, x2)
+    plt.plot(x1, x2, 'o')
+    plt.show()
+
+def example_2_1_2():
+    s_1 = [1, 2, 1, 2]
+    s_2 = [1, 1, 2, 2]
+    s = np.array([s_1, s_2])
+    A = np.array([
+        [1, 2],
+        [1, -1]
+        ])
+    x = A @ s
+    W = la.inv(A)
+    print(W @ x)
+
+def get_cov(A):
+    return (A.T @ A)/(A.shape[0] - 1)
+
+def example_3_3():
+    X = np.array([
+        [1, 1, 2, 0, 5, 4, 5, 3],
+        [3, 2, 3, 3, 4, 5, 5, 4]
+        ]).T
+    mu = X.mean(axis=0)
+    D = X - mu
+    print(D)
+    sigma_D = get_cov(D)
+    print(sigma_D)
+    lamda, V = la.eigh(sigma_D)
+    U = (V @ D.T).T
+    sigma_U = get_cov(U)
+    print(sigma_U)
+
+    Z = np.einsum('i,ij -> ij', lamda**-0.5, U.T).T
+    print(lamda.shape)
+    print(U.shape)
+    print(Z.T)
+    sigma_Z = get_cov(Z)
+    print(sigma_Z)
+
+    # plt.plot(X[:, 0], X[:, 1], 'o')
+    # plt.plot(U[:, 0], U[:, 1], 'o')
+    plt.plot(Z[:, 0], Z[:, 1], 'o')
+    plt.show()
+
+example_3_3()
+
+def pca():
+
+    def get_s1(N):
+        lin = np.linspace(1, 30, N)
+        return 0.5*np.sin(lin)
+
+    def get_s2(N):
[…]

### 2022-04-29 paper.tex
added the sound wave example in script, added some texts in the paper.tex regarding information theoretic approach

@@
 
+Several necessay conditions for ICA: 
+1) The number of sensors is larger than number of sources. This ensures that the mixing matrix is full rank.
+2) THe sources at each sample (time) are mutually independent.
+3) At most one source is normally distributed.
+
 ALgorithms: FastICA, projection pursuit and Infomax. Drawbacks: over-complete or under-complete.
@@
 
+1) Non-Gaussianity maximization approach
+
 The non-Gaussianity thus hints the methodologies used in finding the independent component.
@@
 
-1) Optimizing Kurtosis (high-order based moment approach) (Mick)
-2) non-gaussianity  (Mick)
-3) Maximum likelihood estimation (Kannan)
-    - Minimizing mutual information (min KL divergence)
+%1) Optimizing Kurtosis (high-order based moment approach) (Mick)
+%2) non-gaussianity  (Mick)
+%3) Maximum likelihood estimation (Kannan)
+%    - Minimizing mutual information (min KL divergence)
+
+2) Minimize mutual information, max negentropy, and maximum likelihood
+
+Apart fomr the non-Gaussianity based approaches, there are several information theoetic formulations.
+Here in this section, we will review all different information theoretic frameworks and show that essentially they are equivalent.
+Then we will talk about the learning rules for these theoretical formulations.
+
+We can first consider the basic idea of minimizing mutual information, which is defined as the KL divergence of the multivariate distribution of the observed data $\bf{x}$ and product of all its marginal distributions, i.e.
+\begin{equation}
+I(\bf{x}) = \sum_{\bf{x}}p(\bf{x})\log\frac{p(\bf{x})}{\prod_{i}p_{i}(x_{i})}
+\end{equation}.
+$I(\bf{x})$ is non-negative and is 0 if an only if the $\bf{x}$ are independent.
+Within this framework, several different approaches have been formulated in history.
+One of them is a key finding in neural networks, that is maximizing information between inputs $\bf{x}$ and outputs $\bf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\bf{y}$.
+Maximizing the output entropy is to maximize $H(\bf{y}) = \sum_{i}H(y_{i}) - I(\bf{y})$.
+In a neural network, there is a non-linear activation function that generates the output $\bf{y} = g(\bf{u})$.
+This nonlinearity indeed brings the information of higher order statistics into the optimization process.
 

### 2022-04-29 paper.tex
add outlines

--- 2022-04-24 paper.tex
@@
 \section{Applications}
+1) sklearn example
 3) example: sound single decomposition
-4) tock price
-5) EGG
+4) stock price
+5) EEG
+% (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
+
+\section{How to check the robustness and }
+
+1) check if the performance matrix is identity (Te-won Lee, p. 51)
+\begin{align}
+W A &= I.
+\end{align}
+
+2) check how fast w changes (ICA_projectionPursuit1612.05445.pdf, p. 12)
+
+\section{PCA vs. ICA}
+% (https://www.cs.ubc.ca/~jnutini/documents/mlrg_pca.pdf)
+
+\section{what if two of the sources are dependent}
+s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
+s4 = s1 + s2
+s5 = s1 - s2
 

### 2022-04-29 scripts/ex1.py

@@ -15,8 +15,11 @@ def load_data():
     s1 = np.sin(2*time) # sinusoidal
     s2 = np.sign(np.sin(3*time)) # square signal
     s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
+    s4 = s1 + s2
+    s5 = s1 - s2
 
-    S = np.array([s1, s2, s3])
+    # S = np.array([s1, s2, s3])
+    S = np.array([s3, s4, s5])
     A = np.array([
         [1, 1, 1],
         [0.5, 2, 1],
@@ -37,7 +40,7 @@ def plot_sources(X, S, S_predicted, S_sklearn, S_pca):
         ax[0, 1].set_title("Original sources ($S$)")
 
     for S_predicted in S_predicted:
-        ax[1, 0].plot(S_predicted)
+        ax[1, 0].plot(-S_predicted)
         ax[1, 0].set_title("Predicted sources ($W X$) [Our FastICA]")
 
     for S_sklearn in S_sklearn:
@@ -55,6 +58,11 @@ if __name__ == '__main__':
     np.random.seed(0)
     X, S = load_data()
     S_predicted, distances = fast_ica.ica(X, cycles=100)
+    # print(distances)
+    # plt.plot(distances)
+    # plt.ylim((0, 0.1))
+    # plt.show()
+    # exit()
     S_sklearn = FastICA(n_components=3).fit_transform(X.T).T
     S_pca = PCA(n_components=3).fit_transform(X.T).T
     plot_sources(X, S, S_predicted, S_sklearn, S_pca)

### 2022-04-29 scripts/ex2.py

@@ -1,3 +1,17 @@
+import youtube_dl
+
+def download_sources():
+
+    ydl_opts = {}
+    with youtube_dl.YoutubeDL(ydl_opts) as ydl:
+        ydl.download(['https://www.youtube.com/watch?v=wuvXCC9xy48'])
+        ydl.download(['https://www.youtube.com/watch?v=ea8rkQFXEr4'])
+
+
+
+
+
+
 # def mix_sources(mixtures, apply_noise=False):
 #     for i in range(len(mixtures)):
 #         max_val = np.max(mixtures[i])
@@ -19,3 +33,4 @@
 #     wavfile.write('out1.wav', sampling_rate, S[0])
 #     wavfile.write('out2.wav', sampling_rate, S[1])
 
+if __name__ == '__main__':

### 2022-04-29 scripts/fast_ica.py

@@ -11,10 +11,11 @@ def center(X, standardize=False):
 
 def whiten(X, test=True):
     cov = np.cov(X)
+    print(cov)
     lamda, V = la.eigh(cov)
     lamda_inv = np.sqrt(la.inv(np.diag(lamda)))
     Z = lamda_inv @ V.T @ X
-
+    print(np.cov(Z))
     if test:
         tests.test_identity(np.cov(Z))
 
@@ -40,10 +41,12 @@ def ica(X, cycles=1000, tol=1e-5, test=False):
                 tests.test_gram_schmidt(w_new, W, i)
 
             w_new -= W[:i] @ w_new @ W[:i]
-            w_new /= (w_new ** 2).sum()**0.5
-            distance = np.abs(np.abs((w * w_new).sum()) - 1)
+            w_new /= (w_new**2).sum()**0.5
+            distance = np.abs(np.abs(w @ w_new) - 1)
             distances.append(distance)
             w = w_new
+
+            # check why it doesn't stop
             if distance < tol:
                 break
         W[i, :] = w

### 2022-04-29 paper.tex
edited some part in mathematical formulation

@@
 \begin{equation}
-L_{\alpha}(\bf{u}) := \alpha \gamma^{2}(\bf{u}^{t}\tilde{\bf{x}}) + (1-\alpha)\kappa^{2}(\bf{u}^{t}\tilde{\bf{x}})
+L_{\alpha}(\bf{u}) := \alpha \gamma^{2}(\bf{w}^{t}\tilde{\bf{x}}) + (1-\alpha)\kappa^{2}(\bf{w}^{t}\tilde{\bf{x}})
 \end{equation},
-where $\alpha \in [0, 1]$ and $\bf{u}^{t}\bf{u}$$ = 1$.
+where $\alpha \in [0, 1]$ and $\bf{w}^{t}\bf{w}$$ = 1$.
 
@@
 \end{equation}.
-$I(\bf{x})$ is non-negative and is 0 if an only if the $\bf{x}$ are independent.
-Within this framework, several different approaches have been formulated in history.
-One of them is a key finding in neural networks, that is maximizing information between inputs $\bf{x}$ and outputs $\bf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\bf{y}$.
+$I(\bf{x})$ is non-negative and is 0 if and only if the $\bf{x}$ are independent.
+Within this framework, several different information theoretic approaches have been formulated in history.
+
+One of them is a key finding in neural networks, that is, maximizing information between inputs $\bf{x}$ and outputs $\bf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\bf{y}$.
 Maximizing the output entropy is to maximize $H(\bf{y}) = \sum_{i}H(y_{i}) - I(\bf{y})$.
+The maximum is obtained if marginal distribution of each $y_{i}$ is uniform and independent.
+This implies that the activation function in the neural network $y_{i} = g(\bf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$.
+
 In a neural network, there is a non-linear activation function that generates the output $\bf{y} = g(\bf{u})$.
 This nonlinearity indeed brings the information of higher order statistics into the optimization process.
+
+Another way to formulate this is to maximize the negentropy, which is the KL divergence between $p(\bf{W}\bf{x})$ and Gaussian distribution $p_{G}(\bf{W}\bf{x})$.
+
+Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution.
 
@@
 \section{How to check the robustness and }
+% Amari metric, ref "ELements of Statistical Learning", p.570
 
-1) check if the performance matrix is identity (Te-won Lee, p. 51)
+%1) check if the performance matrix is identity (Te-won Lee, p. 51)
 \begin{align}
@@
 
-2) check how fast w changes (ICA_projectionPursuit1612.05445.pdf, p. 12)
+%2) check how fast w changes (ICA_projectionPursuit1612.05445.pdf, p. 12)
 
@@
 \section{what if two of the sources are dependent}
-s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
-s4 = s1 + s2
-s5 = s1 - s2
+%s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
+%s4 = s1 + s2
+%s5 = s1 - s2
 

### 2022-04-30 paper.tex
add ex2

@@
 %\graphicspath{{~/data/KannanLu/exp623/article/}}
+\usepackage{listings}
+\usepackage{color}
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+% \definecolor{backcolour}{rgb}{0.18,0.18,0.18}
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\footnotesize,
+    breakatwhitespace=false,
+    breaklines=true,
+    captionpos=b,
+    keepspaces=true,
+    numbers=left,
+    numbersep=5pt,
+    showspaces=false,
+    showstringspaces=false,
+    showtabs=false,
+    tabsize=2
+}
+\lstset{style=mystyle}
 
@@
 \end{align}
+\section{Example}
+We implement FastICA according to the formalism described in the previous section.
+The algorithm is applied to a following set of signal functions.
+\begin{align}
+s_1 &= \sin(2 t). \\
+s_2 &= \mathrm{sign} [\sin(3 t)]. \\
+s_3 &= \frac{1}{2\pi} (t~\mathrm{mod}~2\pi).
+\end{align}
+The mixing matrix is taken to be
+\begin{align}
+A &= \begin{bmatrix}
+1 & 1 & 1 \\
+0.5 & 2 & 1 \\
+1.5 & 1 & 2
+\end{bmatrix}.
+\end{align}
+The result from our implementation and FastICA are shown in Fig.~\ref{fig:sklearn}.
+\begin{figure}
+\includegraphics{scripts/sklearn.pdf}
+\caption{\label{fig:sklearn} FastICA results using our code and \lstinline{sklearn.decomposition.FastICA}}
+\end{figure}
 
 \section{Applications}
-1) sklearn example
 3) example: sound single decomposition
-4) stock price
-5) EEG
-% (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
+4) EEG % (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
+5) stock price
 
-\section{How to check the robustness and }
+\section{How to check the robustness and convergence}
 % Amari metric, ref "ELements of Statistical Learning", p.570

### 2022-04-30 scripts/ex1.py

@@ -1,7 +1,6 @@
 from matplotlib import pyplot as plt
 import numpy as np
 from scipy import signal
-from scipy.io import wavfile
 from sklearn.decomposition import FastICA, PCA
 
 import fast_ica
@@ -17,9 +16,10 @@ def load_data():
     s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
     s4 = s1 + s2
     s5 = s1 - s2
+    s6 = np.cos(2*time)
 
-    # S = np.array([s1, s2, s3])
-    S = np.array([s3, s4, s5])
+    S = np.array([s1, s2, s3])
+    # S = np.array([s3, s4, s5])
     A = np.array([
         [1, 1, 1],
         [0.5, 2, 1],
@@ -29,40 +29,38 @@ def load_data():
     return X, S
 
 def plot_sources(X, S, S_predicted, S_sklearn, S_pca):
-    fig, ax = plt.subplots(nrows=2, ncols=3)
+    fig, ax = plt.subplots(nrows=2, ncols=2, sharex=True, sharey=True)
 
     for x in X:
         ax[0, 0].plot(x)
         ax[0, 0].set_title('Mixture signals ($X$)')
+        ax[0, 0].set(ylabel='Amplitude')
 
     for s in S:
         ax[0, 1].plot(s)
         ax[0, 1].set_title("Original sources ($S$)")
 
     for S_predicted in S_predicted:
-        ax[1, 0].plot(-S_predicted)
+        ax[1, 0].plot(S_predicted)
         ax[1, 0].set_title("Predicted sources ($W X$) [Our FastICA]")
+        ax[1, 0].set(xlabel='Time', ylabel='Amplitude')
 
     for S_sklearn in S_sklearn:
         ax[1, 1].plot(S_sklearn)
         ax[1, 1].set_title("Predicted sources ($W X$) [Sklearn FastICA]")
+        ax[1, 1].set(xlabel='Time')
 
-    for S_pca in S_pca:
-        ax[1, 2].plot(S_pca)
-        ax[1, 2].set_title("Predicted sources ($W X$) [Sklearn PCA]")
+    # for S_pca in S_pca:
+    #     ax[1, 2].plot(S_pca)
+    #     ax[1, 2].set_title("Predicted sources ($W X$) [Sklearn PCA]")
 
     fig.tight_layout()
-    plt.show()
+    plt.savefig('sklearn.pdf')
 
 if __name__ == '__main__':
[…]

### 2022-04-30 scripts/ex2.py

@@ -1,16 +1,18 @@
-import youtube_dl
+from matplotlib import pyplot as plt
+import numpy as np
+import sounddevice
+from scipy.io import wavfile
 
-def download_sources():
+import fast_ica
 
-    ydl_opts = {}
-    with youtube_dl.YoutubeDL(ydl_opts) as ydl:
-        ydl.download(['https://www.youtube.com/watch?v=wuvXCC9xy48'])
-        ydl.download(['https://www.youtube.com/watch?v=ea8rkQFXEr4'])
+# import youtube_dl
 
+# def download_sources():
 
-
-
-
+#     ydl_opts = {}
+#     with youtube_dl.YoutubeDL(ydl_opts) as ydl:
+#         ydl.download(['https://www.youtube.com/watch?v=wuvXCC9xy48'])
+#         ydl.download(['https://www.youtube.com/watch?v=ea8rkQFXEr4'])
 
 # def mix_sources(mixtures, apply_noise=False):
 #     for i in range(len(mixtures)):
@@ -33,4 +35,16 @@ def download_sources():
 #     wavfile.write('out1.wav', sampling_rate, S[0])
 #     wavfile.write('out2.wav', sampling_rate, S[1])
 
+def play(sound, fs=11025):
+    for i in range(sound.shape[0]):
+        print(f'playing mixed track {i}')
+        sounddevice.play(sound[i, :], fs, blocking=True)
+
+def load():
+    X = np.loadtxt('mix.dat').T
+    play(X)
+    S_predicted, distances = fast_ica.ica(X, cycles=1000)
+    play(S_predicted)
+
 if __name__ == '__main__':
+    load()

### 2022-04-30 scripts/fast_ica.py

@@ -45,7 +45,7 @@ def ica(X, cycles=1000, tol=1e-5, test=False):
             distance = np.abs(np.abs(w @ w_new) - 1)
             distances.append(distance)
             w = w_new
-
+            print(distance)
             # check why it doesn't stop
             if distance < tol:
                 break

### 2022-04-30 paper.tex
edited the mathematical formulation part

--- 2022-04-29 paper.tex
@@
 
-Apart fomr the non-Gaussianity based approaches, there are several information theoetic formulations.
+Apart fomr the non-Gaussianity based approaches, there are several information theoetic formalisms.
 Here in this section, we will review all different information theoretic frameworks and show that essentially they are equivalent.
@@
 One of them is a key finding in neural networks, that is, maximizing information between inputs $\bf{x}$ and outputs $\bf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\bf{y}$.
-Maximizing the output entropy is to maximize $H(\bf{y}) = \sum_{i}H(y_{i}) - I(\bf{y})$.
-The maximum is obtained if marginal distribution of each $y_{i}$ is uniform and independent.
-This implies that the activation function in the neural network $y_{i} = g(\bf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$.
+Maximizing the information between inputs and outputs is to maximize the output joint entropy $H(\bf{y}) = \sum_{i}H(y_{i}) - I(\bf{y})$, where $I(\bf{y}) = \sum_{\bf{y}}p(\bf{y})\log\frac{p(\bf{y})}{\prod_{i}p_{i}(y_{i})}$ is the mutual information in the outputs $\bf{y}$.
+The mutual information is non-negative and is zero if and only if $\bf{y}$ are marginalized.
+Recall that in neural network, the output is given by the nonlinearity $y_{i} = g_{i}(\bf{w}_{i}^{t}\bf{x})$ $:= g_{i}(u_{i})$, thus $p(y_{i}) = \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|}$.
+The maximum is therefore obtained by considering the gradient,
+\begin{equation}
+\frac{\partial H(\bf{y})}{\partial \bf{W}} = \frac{\partial (-I(\bf{y}))}{\partial \bf{W}} - \frac{\partial}{\partial \bf{W}}\sum_{i} \mathbb{E}[\log \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|} ]
+\end{equation}
+ This implies that the nonlinear function $g_{i}$ in the neural network $y_{i} = g(\bf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$ in order to kill the second term.
+Together with the constraint that the outputs are marginalized, this gradient is zero.
+Therefore, a good estimation of the cdf of the source signal improves drastically the performance of ICA.
 
-In a neural network, there is a non-linear activation function that generates the output $\bf{y} = g(\bf{u})$.
+%In a neural network, there is a non-linear activation function that generates the output $\bf{y} = g(\bf{u})$.
 This nonlinearity indeed brings the information of higher order statistics into the optimization process.
 
-Another way to formulate this is to maximize the negentropy, which is the KL divergence between $p(\bf{W}\bf{x})$ and Gaussian distribution $p_{G}(\bf{W}\bf{x})$.
+Another way to formulate this is to maximize the negentropy $J(u_{i})$, which is the KL divergence $D(p(u_{i})||p_{G}(u_{i}))$ between $p(u_{i})$ and Gaussian distribution $p_{G}(u_{i})$ with the same mean and covariance as $p(u_{i})$.
+Requiring that the $\bf{u}$ can be factorized and decorrelated, the sum of negentropies can be written as
+\begin{align}
+\sum_{i}J(u_{i}) &= \sum_{i}D(p(u_{i})||p_{G}(u_{i})) \\
+&= \sum_{i}p(u_{i})\log\frac{p(u_{i})}{p_{G}(u_{i})} \\
+&= \sum_{\bf{u}}p(\bf{u}) \log\frac{\prod_{i}p(u_{i})}{\prod_{i}p_{G}(u_{i})} \\
+&= \sum_{\bf{u}}p(\bf{u}) \log\frac{\prod_{i}p(u_{i})}{p_{G}(\bf{u}_{i})} \\
+&= \sum_{\bf{u}}p(\bf{u}) \log\frac{\prod_{i}p(u_{i})}{p(\bf{u})} + \sum_{\bf{u}}p(\bf{u})\log\frac{p(\bf{u})}{p_{G}(\bf{u})} \\
+&= D(\prod_{i}p(u_{i})||p(\bf{u})) + J(\bf{u}) \\
+&= -I(\bf{u}) + J(\bf{u}) \\
+& = -I(\bf{u}) - H(\bf{u}) - \sum_{\bf{u}}p(\bf{u})\log p_{G}(\bf{u}) \\
+& = -I(\bf{u}) - H(\bf{x}) - \log(|det(\bf{W})|) - \frac{1}{2}\log((2\pi e)^{d}det(<\bf{u}, \bf{u}^{t}>))\\
+&= -I(\bf{u}) - H(\bf{x}) - \frac{1}{2}\log((2\pi e)^{d})
+\end{align}.
+In the derivation, the $\bf{u}$ are uncorrelated so the covariance matrix is identity.
+Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\bf{u} = \bf{W}\bf{x}$.
+The minimum condition is given by
 

### 2022-05-01 scripts/ex1.py
testing

@@ -1,5 +1,6 @@
 from matplotlib import pyplot as plt
 import numpy as np
+import numpy.linalg as la
 from scipy import signal
 from sklearn.decomposition import FastICA, PCA
 
@@ -19,6 +20,7 @@ def load_data():
     s6 = np.cos(2*time)
 
     S = np.array([s1, s2, s3])
+    S /= S.std(axis=1, keepdims=True)
     # S = np.array([s3, s4, s5])
     A = np.array([
         [1, 1, 1],
@@ -28,7 +30,7 @@ def load_data():
     X = A @ S
     return X, S
 
-def plot_sources(X, S, S_predicted, S_sklearn, S_pca):
+def plot_sources(X, S, S_predicted, S_predicted_std, S_sklearn, S_pca):
     fig, ax = plt.subplots(nrows=2, ncols=2, sharex=True, sharey=True)
 
     for x in X:
@@ -45,22 +47,44 @@ def plot_sources(X, S, S_predicted, S_sklearn, S_pca):
         ax[1, 0].set_title("Predicted sources ($W X$) [Our FastICA]")
         ax[1, 0].set(xlabel='Time', ylabel='Amplitude')
 
-    for S_sklearn in S_sklearn:
-        ax[1, 1].plot(S_sklearn)
-        ax[1, 1].set_title("Predicted sources ($W X$) [Sklearn FastICA]")
-        ax[1, 1].set(xlabel='Time')
+    for S_predicted_std in S_predicted_std:
+        ax[1, 1].plot(S_predicted_std)
+        ax[1, 1].set_title("Predicted sources ($W X$) [Our FastICA, std]")
+        ax[1, 1].set(xlabel='Time', ylabel='Amplitude')
+
+    # for S_sklearn in S_sklearn:
+    #     ax[1, 1].plot(S_sklearn)
+    #     ax[1, 1].set_title("Predicted sources ($W X$) [Sklearn FastICA]")
+    #     ax[1, 1].set(xlabel='Time')
 
     # for S_pca in S_pca:
     #     ax[1, 2].plot(S_pca)
     #     ax[1, 2].set_title("Predicted sources ($W X$) [Sklearn PCA]")
 
     fig.tight_layout()
-    plt.savefig('sklearn.pdf')
+    # plt.show()
+    plt.savefig('all.pdf', bbox_inches='tight')
 
 if __name__ == '__main__':
     np.random.seed(0)
[…]

### 2022-05-01 scripts/ex2.py

@@ -1,39 +1,8 @@
 from matplotlib import pyplot as plt
 import numpy as np
 import sounddevice
-from scipy.io import wavfile
 
 import fast_ica
-
-# import youtube_dl
-
-# def download_sources():
-
-#     ydl_opts = {}
-#     with youtube_dl.YoutubeDL(ydl_opts) as ydl:
-#         ydl.download(['https://www.youtube.com/watch?v=wuvXCC9xy48'])
-#         ydl.download(['https://www.youtube.com/watch?v=ea8rkQFXEr4'])
-
-# def mix_sources(mixtures, apply_noise=False):
-#     for i in range(len(mixtures)):
-#         max_val = np.max(mixtures[i])
-#         if max_val > 1 or np.min(mixtures[i]) < 1:
-#             mixtures[i] = mixtures[i] / (max_val / 2) - 0.5
-#     X = np.c_[[mix for mix in mixtures]]
-#     if apply_noise:
-#         X += 0.02 * np.random.normal(size=X.shape)
-#     return X
-
-# def example():
-#     sampling_rate, mix1 = wavfile.read('mix1.wav')
-#     sampling_rate, mix2 = wavfile.read('mix2.wav')
-#     sampling_rate, source1 = wavfile.read('source1.wav')
-#     sampling_rate, source2 = wavfile.read('source2.wav')
-#     X = mix_sources([mix1, mix2])
-#     S = ica(X)
-#     plot_mixture_sources_predictions(X, [source1, source2], S)
-#     wavfile.write('out1.wav', sampling_rate, S[0])
-#     wavfile.write('out2.wav', sampling_rate, S[1])
 
 def play(sound, fs=11025):
     for i in range(sound.shape[0]):
@@ -43,7 +12,7 @@ def play(sound, fs=11025):
 def load():
     X = np.loadtxt('mix.dat').T
     play(X)
-    S_predicted, distances = fast_ica.ica(X, cycles=1000)
+    S_predicted, distances = fast_ica.ica(X, cycles=1000, standardize=False)
     play(S_predicted)
 
 if __name__ == '__main__':

### 2022-05-01 scripts/fast_ica.py

@@ -11,11 +11,9 @@ def center(X, standardize=False):
 
 def whiten(X, test=True):
     cov = np.cov(X)
-    print(cov)
     lamda, V = la.eigh(cov)
     lamda_inv = np.sqrt(la.inv(np.diag(lamda)))
     Z = lamda_inv @ V.T @ X
-    print(np.cov(Z))
     if test:
         tests.test_identity(np.cov(Z))
 
@@ -27,8 +25,8 @@ def g(x, a1=1):
 def dg(x):
     return 1 - g(x)**2
 
-def ica(X, cycles=1000, tol=1e-5, test=False):
-    X = whiten(center(X))
+def ica(X, standardize=False, cycles=1000, tol=1e-5, test=False):
+    X = whiten(center(X, standardize=standardize))
     nrows = X.shape[0]
     W = np.zeros((nrows, nrows))
     distances = []
@@ -45,10 +43,9 @@ def ica(X, cycles=1000, tol=1e-5, test=False):
             distance = np.abs(np.abs(w @ w_new) - 1)
             distances.append(distance)
             w = w_new
-            print(distance)
-            # check why it doesn't stop
+
             if distance < tol:
                 break
         W[i, :] = w
     S_predicted = W @ X
-    return S_predicted, distances
+    return S_predicted, W, distances

### 2022-05-01 scripts/ex1.py
reproduce sklearn and original X

@@ -6,22 +6,20 @@ from sklearn.decomposition import FastICA, PCA
 
 import fast_ica
 
-plt.rc('font', family='serif')
-plt.rc('text', usetex=True)
+# plt.rc('font', family='serif')
+# plt.rc('text', usetex=True)
 
 def load_data():
     N = 2000
     time = np.linspace(0, 8, N)
-    s1 = np.sin(2*time) # sinusoidal
-    s2 = np.sign(np.sin(3*time)) # square signal
-    s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
+    s1 = 5*np.sin(2*time) # sinusoidal
+    s2 = 20*np.sign(np.sin(3*time)) # square signal
+    s3 = 100*signal.sawtooth(2*np.pi*time) + 50# saw tooth signal
     s4 = s1 + s2
     s5 = s1 - s2
     s6 = np.cos(2*time)
 
     S = np.array([s1, s2, s3])
-    S /= S.std(axis=1, keepdims=True)
-    # S = np.array([s3, s4, s5])
     A = np.array([
         [1, 1, 1],
         [0.5, 2, 1],
@@ -30,61 +28,54 @@ def load_data():
     X = A @ S
     return X, S
 
-def plot_sources(X, S, S_predicted, S_predicted_std, S_sklearn, S_pca):
-    fig, ax = plt.subplots(nrows=2, ncols=2, sharex=True, sharey=True)
+def plot_sources(sources, titles, output=None):
 
-    for x in X:
-        ax[0, 0].plot(x)
-        ax[0, 0].set_title('Mixture signals ($X$)')
-        ax[0, 0].set(ylabel='Amplitude')
-
-    for s in S:
-        ax[0, 1].plot(s)
-        ax[0, 1].set_title("Original sources ($S$)")
-
-    for S_predicted in S_predicted:
-        ax[1, 0].plot(S_predicted)
-        ax[1, 0].set_title("Predicted sources ($W X$) [Our FastICA]")
-        ax[1, 0].set(xlabel='Time', ylabel='Amplitude')
-
-    for S_predicted_std in S_predicted_std:
-        ax[1, 1].plot(S_predicted_std)
-        ax[1, 1].set_title("Predicted sources ($W X$) [Our FastICA, std]")
-        ax[1, 1].set(xlabel='Time', ylabel='Amplitude')
-
-    # for S_sklearn in S_sklearn:
[…]

### 2022-05-01 scripts/ex2.py

@@ -11,8 +11,8 @@ def play(sound, fs=11025):
 
 def load():
     X = np.loadtxt('mix.dat').T
-    play(X)
-    S_predicted, distances = fast_ica.ica(X, cycles=1000, standardize=False)
+    # play(X)
+    S_predicted, W, distances = fast_ica.ica(X, cycles=1000, standardize=False)
     play(S_predicted)
 
 if __name__ == '__main__':

### 2022-05-01 scripts/fast_ica.py

@@ -3,21 +3,29 @@ import numpy.linalg as la
 
 import tests
 
-def center(X, standardize=False):
+def center(X, divide_sd=False):
     mu = X.mean(axis=1, keepdims=True)
-    sd = X.std(axis=1, keepdims=True, ddof=0) if standardize else 1
+    sd = X.std(axis=1, keepdims=True, ddof=0) if divide_sd else 1
     D = (X - mu)/sd
     return D
 
 def whiten(X, test=True):
+    X = center(X)
     cov = np.cov(X)
     lamda, V = la.eigh(cov)
-    lamda_inv = np.sqrt(la.inv(np.diag(lamda)))
+    lamda_inv = la.inv(np.diag(lamda))**0.5
     Z = lamda_inv @ V.T @ X
+
     if test:
         tests.test_identity(np.cov(Z))
 
     return Z
+
+# def unstandardize(X_std, X_raw):
+#     mu = X_raw.mean(axis=1, keepdims=True)
+#     sd = X_raw.std(axis=1, keepdims=True, ddof=0)
+#     X_unstd = X_std*sd + mu
+#     return X_unstd
 
 def g(x, a1=1):
     return np.tanh(a1*x)
@@ -25,21 +33,28 @@ def g(x, a1=1):
 def dg(x):
     return 1 - g(x)**2
 
-def ica(X, standardize=False, cycles=1000, tol=1e-5, test=False):
-    X = whiten(center(X, standardize=standardize))
-    nrows = X.shape[0]
+def ica(X, cycles=1000, tol=1e-5, test=False):
+    # X1 = whiten(X)
+    X = center(X)
+    u, d, _ = la.svd(X, full_matrices=False)
+    K = u / d
+    X1 = K.T @ X
+
+    nrows, ncols = X.shape
+    X1 *= np.sqrt(ncols)
     W = np.zeros((nrows, nrows))
     distances = []
     for i in range(nrows):
-        w = np.random.random((nrows))
+        w = np.random.normal(size=(nrows))
+        w /= ((w**2).sum())**0.5
         for _ in range(cycles):
-            w_new = (X * g(w.T @ X)).mean(axis=1) - dg(w.T @ X).mean() * w
+            w_new = (X1 * g(w.T @ X1)).mean(axis=1) - dg(w.T @ X1).mean() * w
 
             if test:
                 tests.test_gram_schmidt(w_new, W, i)
 
-            w_new -= W[:i] @ w_new @ W[:i]
-            w_new /= (w_new**2).sum()**0.5
[…]

### 2022-05-01 scripts/ex1.py
fix K

@@ -63,7 +63,7 @@ if __name__ == '__main__':
     source_title = [
         (S, 'Original sources ($S$)'),
         (X, 'Mixture signals ($X$)'),
-        (fast_ica.whiten(X), 'Whiten $X$'),
+        (fast_ica.whiten(X)[0], 'Whiten $X$'),
         (S_predicted, 'Predicted $S$ [Our FastICA]'),
         (X_new, 'Retrieved $X = W^{-1} S_{\\mathrm{predicted}}$ [Our FastICA]'),
         # (S_sklearn, 'Predicted $S$ [Sklearn FastICA]'),

### 2022-05-01 scripts/fast_ica.py

@@ -14,18 +14,19 @@ def whiten(X, test=True):
     cov = np.cov(X)
     lamda, V = la.eigh(cov)
     lamda_inv = la.inv(np.diag(lamda))**0.5
-    Z = lamda_inv @ V.T @ X
+    K = lamda_inv @ V.T
+    Z = K @ X
 
     if test:
         tests.test_identity(np.cov(Z))
 
-    return Z
+    return Z, K.T
 
-# def unstandardize(X_std, X_raw):
-#     mu = X_raw.mean(axis=1, keepdims=True)
-#     sd = X_raw.std(axis=1, keepdims=True, ddof=0)
-#     X_unstd = X_std*sd + mu
-#     return X_unstd
+def whiten_svd(X):
+    U, D, _ = la.svd(X, full_matrices=False)
+    K = U / D
+    Z = K.T @ X
+    return Z, K
 
 def g(x, a1=1):
     return np.tanh(a1*x)
@@ -34,19 +35,20 @@ def dg(x):
     return 1 - g(x)**2
 
 def ica(X, cycles=1000, tol=1e-5, test=False):
-    # X1 = whiten(X)
+
     X = center(X)
-    u, d, _ = la.svd(X, full_matrices=False)
-    K = u / d
-    X1 = K.T @ X
+
 
     nrows, ncols = X.shape
-    X1 *= np.sqrt(ncols)
+
+    X1, K = whiten(X)
+
+
     W = np.zeros((nrows, nrows))
     distances = []
     for i in range(nrows):
-        w = np.random.normal(size=(nrows))
-        w /= ((w**2).sum())**0.5
+        w = np.random.random((nrows))
+
         for _ in range(cycles):
             w_new = (X1 * g(w.T @ X1)).mean(axis=1) - dg(w.T @ X1).mean() * w
 

### 2022-05-01 scripts/ex1.py
clean

@@ -31,7 +31,7 @@ def load_data():
 def plot_sources(sources, titles, output=None):
 
     n = len(sources)
-    fig, ax = plt.subplots(nrows=n, ncols=1, sharex=True, sharey=False)
+    fig, ax = plt.subplots(nrows=n, ncols=1, figsize=(7, 7), sharex=True, sharey=False)
     for i in range(n):
         for source in sources[i]:
             ax[i].plot(source)
@@ -44,38 +44,24 @@ def plot_sources(sources, titles, output=None):
         plt.savefig(output, bbox_inches='tight')
 
 if __name__ == '__main__':
-
     np.random.seed(0)
     X, S = load_data()
-    S_predicted, W, K, _ = fast_ica.ica(X)
-    tt= S_predicted.T @ la.inv(W @ K.T).T
-    print(X.mean(axis=1))
-    print(tt.shape)
-    # mean = np.einsum('ij,i->ij', np.ones(X.shape), X.mean(axis=1, keepdims=True))
-    X_new = ( S_predicted.T @ la.inv(W @ K.T).T + X.mean(axis=1)).T
-    sklearn_ica = FastICA()
-    S_sklearn = sklearn_ica.fit_transform(X.T).T
-    # S_sklearn = fast_ica.whiten(S_sklearn)
-    print(sklearn_ica.mean_)
-    W_sklearn = sklearn_ica.mixing_
-    X_new_sklearn = (S_sklearn.T @ W_sklearn.T + sklearn_ica.mean_).T
-
+    S_our, W, K, _ = fast_ica.ica(X)
+    X_our = la.inv(W @ K) @ S_our + X.mean(axis=1, keepdims=True)
+    ica_sk = FastICA()
+    S_sk = ica_sk.fit_transform(X.T).T
+    X_sk = la.inv(ica_sk._unmixing @ ica_sk.whitening_) @ S_sk + X.mean(axis=1, keepdims=True)
     source_title = [
         (S, 'Original sources ($S$)'),
         (X, 'Mixture signals ($X$)'),
         (fast_ica.whiten(X)[0], 'Whiten $X$'),
-        (S_predicted, 'Predicted $S$ [Our FastICA]'),
-        (X_new, 'Retrieved $X = W^{-1} S_{\\mathrm{predicted}}$ [Our FastICA]'),
-        # (S_sklearn, 'Predicted $S$ [Sklearn FastICA]'),
[…]

### 2022-05-01 scripts/ex2.py

@@ -12,7 +12,7 @@ def play(sound, fs=11025):
 def load():
     X = np.loadtxt('mix.dat').T
     # play(X)
-    S_predicted, W, distances = fast_ica.ica(X, cycles=1000, standardize=False)
+    S_predicted, W, distances = fast_ica.ica(X, standardize=False)
     play(S_predicted)
 
 if __name__ == '__main__':

### 2022-05-01 scripts/fast_ica.py

@@ -20,13 +20,13 @@ def whiten(X, test=True):
     if test:
         tests.test_identity(np.cov(Z))
 
-    return Z, K.T
+    return Z, K
 
 def whiten_svd(X):
     U, D, _ = la.svd(X, full_matrices=False)
     K = U / D
     Z = K.T @ X
-    return Z, K
+    return Z, K.T
 
 def g(x, a1=1):
     return np.tanh(a1*x)
@@ -34,21 +34,14 @@ def g(x, a1=1):
 def dg(x):
     return 1 - g(x)**2
 
-def ica(X, cycles=1000, tol=1e-5, test=False):
-
+def ica(X, cycles=200, tol=1e-5, test=False):
     X = center(X)
-
-
-    nrows, ncols = X.shape
-
     X1, K = whiten(X)
-
-
+    nrows, ncols = X.shape
     W = np.zeros((nrows, nrows))
     distances = []
     for i in range(nrows):
         w = np.random.random((nrows))
-
         for _ in range(cycles):
             w_new = (X1 * g(w.T @ X1)).mean(axis=1) - dg(w.T @ X1).mean() * w
 
@@ -65,7 +58,5 @@ def ica(X, cycles=1000, tol=1e-5, test=False):
                 break
         W[i, :] = w
 
-    print('my W', W)
-
-    S_predicted = W @ K.T @ X
+    S_predicted = W @ K @ X
     return S_predicted, W, K, distances

### 2022-05-01 paper.tex
testing
reproduce sklearn and original X
edited intro, math, sections of the paper

@@
 \frenchspacing
-%\graphicspath{{~/data/KannanLu/exp623/article/}}
+%\graphicspath{/pics/}
 
@@
 \section{Introduction}
-Why ICA: 1) In reality, source signals are often corrupted with noise, or in other words, data are composed by mutiple independent source signals.
-To isolate the signals from different sources are called 'Blind Source Separation' (BSS).
-Famous example is 'Cocktail party problem'.
-ICA is widely used in images, sounds, stock market and medicine, can be also viewed as a dimension reduction technique (or filtering operation).
-ICA is an extension of PCA, where the latter pertains to the second order statistics (covariance matrix of data).
-ICA optmizes higher-order statistics.
+Independent component analysis (ICA) is one of the classical unsupervised learning techniques that is used to seperate out latent variables or reduce dimensions.
+The motivation of the development of this technique is based on very commom problems.
+For instance, in reality, source signals are often corrupted with 'noise'.
+Or in other words, data are composed by mutiple independent source signals.
+To isolate the signals from different sources, ICA can be a good choice.
+The famous example is the 'cocktail party problem', where the sound detectors record the superposition of various sound signals from different sources in the party.
+The goal is to seperate out the observed signal into sounds from each different source.
+Since we do not know what the sources are and how these signals are mixed, this type of problem is called 'blind source separation' (BSS).
+ICA is widely used in images, sounds, stock market and medical signals.
+ICA can also be viewed as an extension of principle component analysis (PCA), where the latter maximizes the second order statistics (covariance matrix of data).
+ICA maximizes higher-order statistics or simply tries to look for independent components not just uncorrelated components.
+In Fig.~\ref{fig:ICA_vs_PCA}, we show the ICA components $IC_{i}$ and PCA components $PC_{i}$ learned from the observed data $x_{i}$ with the original sources $s_{i}$ sampled from uniform distribution independently.
+Clearly.
+the ICA learned the independent components correctly but the PCA did not in this case.
 
-Several necessay conditions for ICA: 
-1) The number of sensors is larger than number of sources. This ensures that the mixing matrix is full rank.
-2) THe sources at each sample (time) are mutually independent.
-3) At most one source is normally distributed.
+In this paper, we will review mathematical background of the ICA technique, the detailed algorithm implementation and several realizations in typical examples.
 
-ALgorithms: FastICA, projection pursuit and Infomax. Drawbacks: over-complete or under-complete.
-
+\begin{figure}[b]
+\includegraphics[width=0.8\textwidth]{ICA_vs_PCA.pdf}
+ \caption{\label{fig:ICA_vs_PCA} shows the PCA and ICA result of unmixing the observed data $x_{i}$ generated from two independent uniform variables $s_{i}$.
+The algorithm is based on maximizing the higher order moments.
+}
+\end{figure}
 % Put \label in argument of \section for cross-refe(WHAT DOES THIS LAST SENTENCE MEAN?)rencing
@@
 
-\section{Mathematical Formulation}
-0) Proof that gaussian signals cannot be separated (Kannan)
-//
-
-Before we delve into the detailed mathematical formulations, we first define the notations that we will use throughout the whole document.
+\section{Mathematical Background}
+\subsection{Definition of the problem and notations}
+Before we delve into the detailed mathematical formalisms, we first define the notations that we will use throughout the whole document.
 We will use boldface capital letters $\bf{A}$ for matrices, boldface lowercase letters $\bf{a}$ for vectors and ordinary lower case letters $a$ for scalars.
@@
 That is there is a linear map between the $\bf{s}$$^{(i)}$ and $\bf{x}$$^{(i)}$, $\bf{x}$$^{(i)}$ $=$ $\bf{A}$$\bf{s}$$^{(i)}$, where $\bf{A} \in \rm I\!R^{d\times d}$ is known as the mixing matrix.
-The whole idea of ICA is to learn the inverse of the mixing matrix $\bf{W} = \bf{A}$$^{-1}$ from the observed data $\bf{x}$$^{(i)}$, known as the un-mixing.
+The whole idea of ICA is to learn the inverse of the mixing matrix $\bf{W} = \bf{A}$$^{-1}$ from the observed data $\bf{x}$$^{(i)}$, known as the un-mixing process.
+The learned independent components are denoted by $\bf{u} = \bf{W}\bf{x}$.
 This can be done in several different ways based on ideas of seperating non-Gaussian signals, and the original signals being independent.
 We will in this section discuss various mathematical formulations.
+All these various theoretical frameworks are based on several constraints of the problem.
+These are:\\
+(1) The number of sensors is larger than number of sources. This ensures that the mixing matrix is full rank; \\
+(2) The sources at each sample (time) are mutually independent; \\
+(3) At most one source is normally distributed. \\
+Without these conditions, the BSS problem is ill-defined.
+For simplicity, we will also restrict the discussion to cases where the number of detectors and sources are the same.
+However, a generalization to condition (1) is feasible and the algorithms discussed in the following sections will still work.
 
-//
-
-First of all, we need to talk about some ambiguities embedded in the symmetry of the problem and justify that if the origianl signals are all Gaussians, they cannot be learned through the un-mixing.
-
+Based on these definitions, we need to talk about some ambiguities embedded in the symmetry of the problem and justify that if the origianl signals are all Gaussians, they cannot be learned through the un-mixing.
 There are two ambiguities that usually do not affect the practical application, namely the permutation ambiguity and the scaling ambiguity.
@@
 Due to the scale ambiguity, this is the same as considering normally distributed signals with unity variance.
-Since each column $j$ of the mixing matrix $\bf{A}$ can be scaled by $\sigma_{j}$, $\tilde{\bf{A}}_{:j} = \sigma_{j}\bf{A}_{:j}$ and correspondingly the signal needs to be scaled by $1/\sigma_{j}$, $\tilde{s_{j}} = \frac{s_{j}-\mu_{j}}{\sigma_{j}}$ (known as the whitening).
+Since each column $j$ of the mixing matrix $\bf{A}$ can be scaled by $\sigma_{j}$, $\tilde{\bf{A}}_{:j} = \sigma_{j}\bf{A}_{:j}$ and correspondingly the signal needs to be scaled by $1/\sigma_{j}$, $\tilde{s_{j}} = \frac{s_{j}-\mu_{j}}{\sigma_{j}}$.
 Then,
 \begin{equation}
-Cov[\bf{x}] = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t}
-\end{equation}.
-Now, if we consider a rotation $\bf{R} $$\in O(d)$ acting on the scaled (whitened) source signals $\bf{\tilde{s}}$.
+Cov[\bf{x}] = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t}.
+\end{equation}
+Now, if we consider a rotation $\bf{R} $$\in O(d)$ acting on the scaled source signals $\bf{\tilde{s}}$.
 The mixing matrix then changes to $\tilde{\bf{A}}\bf{R}$.
@@
 \begin{equation}
-Cov[\bf{x'}] = \tilde{\bf{A}}\bf{R}\bf{I}_{d\times d}\bf{R}^{t}\tilde{\bf{A}}^{t} = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t} 
-\end{equation}.
-This simply means that whether the sources are rotated or not the observed data will be distributed as $\mathcal{N}(0, \tilde{\bf{A}}\tilde{\bf{A}}^{t})$.
+Cov[\bf{x'}] = \tilde{\bf{A}}\bf{R}\bf{I}_{d\times d}\bf{R}^{t}\tilde{\bf{A}}^{t} = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t}. 
+\end{equation} This simply means that whether the sources are rotated or not the observed data will be distributed as $\mathcal{N}(0, \tilde{\bf{A}}\tilde{\bf{A}}^{t})$.
 Thus, because of the rotational symmetry of the multivariate Gaussian distribution, we cannot seperate and obtain the original source signals.
+These derivations also indicate that given the observed data $\bf{x}$, we can whiten the data and it will not affect the seperated signals (up to scalings and mean translations).
 
-//
+\subsection{Higher order statistics approach}
 
-1) Non-Gaussianity maximization approach
-
-The non-Gaussianity thus hints the methodologies used in finding the independent component.
+The non-Gaussianity thus directly provide us with the methodologies used in finding the independent component.
 Notice that the PCA only finds the uncorrelated components but not necessarily independent.
@@
 \begin{equation}
-L_{\alpha}(\bf{u}) := \alpha \gamma^{2}(\bf{w}^{t}\tilde{\bf{x}}) + (1-\alpha)\kappa^{2}(\bf{w}^{t}\tilde{\bf{x}})
+L_{\alpha}(\bf{w}) := \alpha \gamma^{2}(\bf{w}^{t}\tilde{\bf{x}}) + (1-\alpha)\kappa^{2}(\bf{w}^{t}\tilde{\bf{x}})
 \end{equation},
@@
 
-2) Minimize mutual information, max negentropy, and maximum likelihood
+\subsection{Information theoretic approach}
 
-Apart fomr the non-Gaussianity based approaches, there are several information theoetic formalisms.
+Apart from the moment based approaches, there are several information theoetic formalisms.
 Here in this section, we will review all different information theoretic frameworks and show that essentially they are equivalent.
-Then we will talk about the learning rules for these theoretical formulations.
+Then we will talk about the learning rules for these theoretical formulations in the subsequant setion.
 
-We can first consider the basic idea of minimizing mutual information, which is defined as the KL divergence of the multivariate distribution of the observed data $\bf{x}$ and product of all its marginal distributions, i.e.
+The centralized quantity in all these formalisms is the mutual information, which is defined as the KL divergence of the multivariate distribution of the observed data $\bf{x}$ and product of all its marginal distributions, i.e.
 \begin{equation}
-I(\bf{x}) = \sum_{\bf{x}}p(\bf{x})\log\frac{p(\bf{x})}{\prod_{i}p_{i}(x_{i})}
-\end{equation}.
-$I(\bf{x})$ is non-negative and is 0 if and only if the $\bf{x}$ are independent.
-Within this framework, several different information theoretic approaches have been formulated in history.
+I(\bf{x}) = \sum_{\bf{x}}p(\bf{x})\log\frac{p(\bf{x})}{\prod_{i}p_{i}(x_{i})}.
+\end{equation} $I(\bf{x})$ is non-negative and is 0 if and only if the $\bf{x}$ are independent.
+Several different information theoretic approaches have been formulated in history and can be unified in the concept of minimizing the mutual information of the seperated components $\bf{u} = \bf{W}\bf{x}$.
 
@@
 Therefore, a good estimation of the cdf of the source signal improves drastically the performance of ICA.
+Notice that, if $I(\bf{y}) = 0$, the $\bf{u} = \bf{W}\bf{x}$ should also satisfy $I(\bf{u}) = 0$ as $g_{i}$ is an invertible monotonic function.
+Hence, the neural network approach can be eventually reduced to minimizing the mutual information $I(\bf{u})$.
 
@@
 Another way to formulate this is to maximize the negentropy $J(u_{i})$, which is the KL divergence $D(p(u_{i})||p_{G}(u_{i}))$ between $p(u_{i})$ and Gaussian distribution $p_{G}(u_{i})$ with the same mean and covariance as $p(u_{i})$.
+Recall that the Gaussian distribution has maximum entropy constrained with the mean and covariance.
+The negentropy thus measures non-Gaussianity.
 Requiring that the $\bf{u}$ can be factorized and decorrelated, the sum of negentropies can be written as
@@
 & = -I(\bf{u}) - H(\bf{x}) - \log(|det(\bf{W})|) - \frac{1}{2}\log((2\pi e)^{d}det(<\bf{u}, \bf{u}^{t}>))\\
-&= -I(\bf{u}) - H(\bf{x}) - \frac{1}{2}\log((2\pi e)^{d})
-\end{align}.
-In the derivation, the $\bf{u}$ are uncorrelated so the covariance matrix is identity.
+&= -I(\bf{u}) - H(\bf{x}) - \frac{1}{2}\log((2\pi e)^{d}).
+\end{align} In the derivation, the $\bf{u}$ are uncorrelated so the covariance matrix is identity.
 Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\bf{u} = \bf{W}\bf{x}$.
-The minimum condition is given by
 
-Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution.
+Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution $\hat{p}_{s}(\bf{w}_{i}\bf{x})$ satisfying $p(\bf{x}) = \prod_{i}^{d}\hat{p}_{s}(\bf{w}_{i}\bf{x})|\bf{W}|$.
+The log-likelihood is
+\begin{equation}
+l(\bf{W}) = \frac{1}{N}\sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\bf{w}_{i}\bf{x}^{(j)}))+\log|\bf{W}|).
+\end{equation}   
+If the approximation $\hat{p}_{s}(\bf{w}_{i}\bf{x})$ are close to the actual pdf, the first term is approximately $-\sum_{i}H(\bf{w}_{i}\bf{x})$ and is equal to the negative of the mutual information up to an additive constant of the total entropy of $\bf{x}$.
+Hence, all these historical information theoretic approaches are equivalent and the key idea is to approximate the probability distribution of the sources correctly.
+Historically, there has been several proposed parametric distributions for approximating super-Gaussian or sub-Gaussian distributions.
+We will proceed to discuss the algorithm associated with the information based approach in the next section.
 
 \section{Algorithms}
+There are some detailed difference on how gradient algorithms are implemented, i.e. ordinary gradient descent or fixed point algorithm.
 \subsection{objective function based on maximum likelihood  (Kannan)}

### 2022-05-01 paper.tex

@@
 %\graphicspath{/pics/}
+%\graphicspath{{~/data/KannanLu/exp623/article/}}
+\usepackage{listings}
+\usepackage{color}
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+% \definecolor{backcolour}{rgb}{0.18,0.18,0.18}
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\footnotesize,
+    breakatwhitespace=false,
+    breaklines=true,
+    captionpos=b,
+    keepspaces=true,
+    numbers=left,
+    numbersep=5pt,
+    showspaces=false,
+    showstringspaces=false,
+    showtabs=false,
+    tabsize=2
+}
+\lstset{style=mystyle}
 
@@
 %\homepage[]{Your web page}
-%\thanks{}
-%\altaffiliation{}
-\author{K. Krongchon}
-\author{K. Lu}
+\author{Kittithat Krongchon}
+\author{Kannan Lu}
 \affiliation{Department of Physics, University of Illinois at Urbana-Champaign, Urbana, IL 61801, USA}
@@
 
-% insert suggested keywords - APS authors don't need to do this
-%\keywords{}
-
-%\maketitle must follow title, authors, abstract, and keywords
 \maketitle
 
-% body of paper here - Use proper section commands
-% References should be done using the \cite, \ref, and \label commands
 \section{Introduction}
@@
 \end{align}
+\section{Example}
+We implement FastICA according to the formalism described in the previous section.
+The algorithm is applied to a following set of signal functions.
+\begin{align}
+s_1 &= \sin(2 t). \\
+s_2 &= \mathrm{sign} [\sin(3 t)]. \\
+s_3 &= \frac{1}{2\pi} (t~\mathrm{mod}~2\pi).
+\end{align}
+The mixing matrix is taken to be
+\begin{align}
+A &= \begin{bmatrix}
+1 & 1 & 1 \\
+0.5 & 2 & 1 \\
+1.5 & 1 & 2
+\end{bmatrix}.
+\end{align}
+The result from our implementation and FastICA are shown in Fig.~\ref{fig:sklearn}.
+% \begin{figure}
+% \includegraphics{scripts/sklearn.pdf}
+% \caption{\label{fig:sklearn} FastICA results using our code and \lstinline{sklearn.decomposition.FastICA}}
+% \end{figure}
 
 \section{Applications}
-1) sklearn example
 3) example: sound single decomposition
-4) stock price
-5) EEG
-% (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
+4) EEG % (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
+5) stock price
 
-\section{How to check the robustness and }
+\section{How to check the robustness and convergence}
 % Amari metric, ref "ELements of Statistical Learning", p.570
@@
 
-\begin{acknowledgments}
-
-\end{acknowledgments}
-
-% If in two-column mode, this environment will change to single-column
-% format so that long equations can be displayed. Use
-% sparingly.
-%\begin{widetext}
-% put long equation here
-%\end{widetext}
-
-% figures should be put into the text as floats.
-% Use the graphics or graphicx packages (distributed with LaTeX2e)
-% and the \includegraphics macro defined in those packages.
-% See the LaTeX Graphics Companion by Michel Goosens, Sebastian Rahtz,
-% and Frank Mittelbach for instance.
-%
-% Here is an example of the general form of a figure:
-% Fill in the caption in the braces of the \caption{} command. Put the label
-% that you will use with \ref{} command in the braces of the \label{} command.
-% Use the figure* environment if the figure should span across the
-% entire page. There is no need to do explicit centering.
-
-% \begin{figure}
-% \includegraphics{}%
-% \caption{\label{}}
-% \end{figure}
-
-% Surround figure environment with turnpage environment for landscape
-% figure
-% \begin{turnpage}
-% \begin{figure}
-% \includegraphics{}%
-% \caption{\label{}}
-% \end{figure}
-% \end{turnpage}
-
-% tables should appear as floats within the text
-%
-% Here is an example of the general form of a table:
-% Fill in the caption in the braces of the \caption{} command. Put the label
-% that you will use with \ref{} command in the braces of the \label{} command.
-% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
-% \begin{tabular}{} command.
-% The ruledtabular enviroment adds doubled rules to table and sets a
-% reasonable default table settings.
-% Use the table* environment to get a full-width table in two-column
-% Add \usepackage{longtable} and the longtable (or longtable*}
-% environment for nicely formatted long tables. Or use the the [H]
-% placement option to break a long table (with less control than
-% in longtable).
-% \begin{table}%[H] add [H] placement to break table across pages
-% \caption{\label{}}
-% \begin{ruledtabular}
-% \begin{tabular}{}
-% Lines of table here ending with \\
-% \end{tabular}
-% \end{ruledtabular}
-% \end{table}
-
-% Surround table environment with turnpage environment for landscape
-% table
-% \begin{turnpage}
-% \begin{table}
-% \caption{\label{}}
-% \begin{ruledtabular}
-% \begin{tabular}{}
-% \end{tabular}
-% \end{ruledtabular}
-% \end{table}
-% \end{turnpage}
-
-% Specify following sections are appendices. Use \appendix* if there
-% only one appendix.
-%\appendix
-%\section{}
-
-% If you have acknowledgments, this puts in the proper section head.
-%\begin{acknowledgments}
-% put your acknowledgments here.
-%\end{acknowledgments}
-
-% Create the reference section using BibTeX:
-%\bibliography{****.bib}
-
 \end{document}
-%
-% ****** End of file apstemplate.tex ******
-
 \grid

### 2022-05-01 paper.tex
edited the algorithm section in paper

@@
 \section{Algorithms}
-There are some detailed difference on how gradient algorithms are implemented, i.e. ordinary gradient descent or fixed point algorithm.
-\subsection{objective function based on maximum likelihood  (Kannan)}
-
-\subsection{FastICA}
-\subsection{alternative objective function, skewness and kurtosis (Kannan)}
+In this section, we will discuss some of the common algorithms based on the previous mathematical formalisms.
+We will categorize them into moment-based approaches and information-based approaches.
+In either case, the problem is equivalent to an optimization problem.
+There are some detailed differences on how this optimization problem is implemented, i.e.
+ordinary gradient descent, Newton's method or fixed point algorithm.
+We will simply specify what we implemented for various examples that we will discuss later.
+\subsection{Moment-based approach}
 In this method, we maximize the non-Gaussianity given by
@@
 \end{align}
-\section{Example}
+\subsection{Information theoretic approach}
+
+\section{Applications}
 We implement FastICA according to the formalism described in the previous section.
@@
 
-\section{Applications}
-3) example: sound single decomposition
-4) EEG % (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
-5) stock price
+%3) example: sound single decomposition
+%4) EEG % (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
+%5) stock price
 
@@
 %2) check how fast w changes (ICA_projectionPursuit1612.05445.pdf, p. 12)
-
-\section{PCA vs. ICA}
-% (https://www.cs.ubc.ca/~jnutini/documents/mlrg_pca.pdf)
 

### 2022-05-01 scripts/ex1.py
update examples

@@ -31,7 +31,7 @@ def load_data():
 def plot_sources(sources, titles, output=None):
 
     n = len(sources)
-    fig, ax = plt.subplots(nrows=n, ncols=1, figsize=(7, 7), sharex=True, sharey=False)
+    fig, ax = plt.subplots(nrows=n, ncols=1, figsize=(9, 7), sharex=True, sharey=False)
     for i in range(n):
         for source in sources[i]:
             ax[i].plot(source)
@@ -43,25 +43,42 @@ def plot_sources(sources, titles, output=None):
     else:
         plt.savefig(output, bbox_inches='tight')
 
+def plot_distances(distances, output=None):
+    n = len(distances)
+    fig, ax = plt.subplots(nrows=1, ncols=1, figsize=(3, 3))
+    colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
+    for i in range(n):
+        ax.plot(distances[i], 'o-', mec='white', color=colors[i], label=f'Component {i}')
+    ax.set(xlabel='Cycle', ylabel='$||w_{n}^T w_{n-1}| - 1|$')
+    ax.legend(fancybox=False, edgecolor='k')
+
+    if output is None:
+        plt.show()
+    else:
+        plt.savefig(output, bbox_inches='tight')
+
 if __name__ == '__main__':
     np.random.seed(0)
     X, S = load_data()
-    S_our, W, K, _ = fast_ica.ica(X)
-    X_our = la.inv(W @ K) @ S_our + X.mean(axis=1, keepdims=True)
-    ica_sk = FastICA()
-    S_sk = ica_sk.fit_transform(X.T).T
-    X_sk = la.inv(ica_sk._unmixing @ ica_sk.whitening_) @ S_sk + X.mean(axis=1, keepdims=True)
+    S_out, W, K, X_out, distances = fast_ica.ica(X)
+    print(W)
+    plot_distances(distances, output='ex1_dist.pdf')
+
+    S_out_sk, W, K, X_out_sk, distances = fast_ica.ica_sk(X)
+    S_pca = PCA().fit_transform(X.T).T
+
     source_title = [
         (S, 'Original sources ($S$)'),
         (X, 'Mixture signals ($X$)'),
         (fast_ica.whiten(X)[0], 'Whiten $X$'),
[…]

### 2022-05-01 scripts/ex2.py

@@ -9,11 +9,24 @@ def play(sound, fs=11025):
         print(f'playing mixed track {i}')
         sounddevice.play(sound[i, :], fs, blocking=True)
 
-def load():
+def extract():
+    '''
+    Loads mixing sound samples from
+    - Godfather
+    - Southpark
+    - Beethoven 5th
+    - Austin Powers
+    - The Matrix
+    '''
     X = np.loadtxt('mix.dat').T
     # play(X)
-    S_predicted, W, distances = fast_ica.ica(X, standardize=False)
-    play(S_predicted)
+
+    S_out, W, K, X_out, distances = fast_ica.ica(X)
+    # play(S_out)
+    play(fast_ica.center(S_out, divide_sd=True))
+    # S_out, W, K, X_out, distances = fast_ica.ica_sk(X)
+    # play(fast_ica.center(S_out, divide_sd=True))
 
 if __name__ == '__main__':
-    load()
+    np.random.seed(0)
+    extract()

### 2022-05-01 scripts/ex3.py

@@ -2,7 +2,6 @@ import h5py
 import matplotlib.pyplot as plt
 import numpy as np
 import numpy.linalg as la
-import seaborn as sns
 
 import fast_ica
 
@@ -16,35 +15,38 @@ def loadmat(fn):
             d[k] = np.array(v)
     return d
 
-def plot_eeg(d, output='eeg.pdf'):
+def plot_eeg(d, d2=None, output='eeg.pdf'):
     ncomponents = d.shape[0]
     fig, ax = plt.subplots(nrows=8, ncols=8, sharex=True, sharey=False)
+    colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
     for comp in range(ncomponents):
         i = int(comp/8)
         j = int(comp%8)
-        ax[i, j].plot(d[comp])
-        # ax[i, j].set_title(f'{comp}', fontsize=6)
-        ax[i, j].set_yticks([])
+        ax[i, j].plot(d[comp], '-', color=colors[0])
+        if d2 is not None:
+            ax[i, j].plot(d2[comp], '-', color=colors[1])
+        ax[i, j].text(0.5, 1, f'{comp}', transform=ax[i, j].transAxes, fontsize=6)
+        # ax[i, j].set_yticks([])
         ax[i, j].set_xticks([])
     fig.tight_layout()
-    plt.savefig(output)
+    if output is None:
+        plt.show()
+    else:
+        plt.savefig(output, bbox_inches='tight')
+
+def remove_blinks():
+    np.random.seed(0)
+    X = loadmat('ex2_eeg.mat')['Data']
+    plot_eeg(X, output='ex2_raw.pdf')
+    S_out, W, K, X_out, distances = fast_ica.ica(X)
+    plot_eeg(S_out, output='ex2_ica.pdf')
+    S_out_noblinks = S_out.copy()
+    S_out_noblinks[3, :] = np.zeros(X.shape[1])
+    X_noblinks = fast_ica.retrieve_X_out(X, W, K, S_out_noblinks)
+    plot_eeg(X_noblinks, output='ex2_noblinks.pdf')
+    plot_eeg(X, d2=X_noblinks, output='ex2_noblinks.pdf')
 
 
-def std(x):
-    return (x - x.mean())/np.std(x)
 
 if __name__ == '__main__':
-    X = loadmat('eeg.mat')['Data']
-    X = X[:2]
[…]

### 2022-05-01 scripts/fast_ica.py

@@ -1,5 +1,6 @@
 import numpy as np
 import numpy.linalg as la
+from sklearn.decomposition import FastICA, PCA
 
 import tests
 
@@ -23,6 +24,7 @@ def whiten(X, test=True):
     return Z, K
 
 def whiten_svd(X):
+    X = center(X)
     U, D, _ = la.svd(X, full_matrices=False)
     K = U / D
     Z = K.T @ X
@@ -34,14 +36,18 @@ def g(x, a1=1):
 def dg(x):
     return 1 - g(x)**2
 
+def retrieve_X_out(X, W, K, S_out):
+    X_out = la.inv(W @ K) @ S_out + X.mean(axis=1, keepdims=True)
+    return X_out
+
 def ica(X, cycles=200, tol=1e-5, test=False):
-    X = center(X)
     X1, K = whiten(X)
     nrows, ncols = X.shape
     W = np.zeros((nrows, nrows))
     distances = []
     for i in range(nrows):
         w = np.random.random((nrows))
+        dd = []
         for _ in range(cycles):
             w_new = (X1 * g(w.T @ X1)).mean(axis=1) - dg(w.T @ X1).mean() * w
 
@@ -51,12 +57,23 @@ def ica(X, cycles=200, tol=1e-5, test=False):
             w_new -= w_new @ W[:i].T @ W[:i]
             w_new /= ((w_new**2).sum())**0.5
             distance = np.abs(np.abs(w @ w_new) - 1)
-            distances.append(distance)
+            dd.append(distance)
             w = w_new
 
             if distance < tol:
+                distances.append(np.array(dd))
                 break
+
         W[i, :] = w
 
-    S_predicted = W @ K @ X
-    return S_predicted, W, K, distances
+    S_out = W @ K @ X
+    X_out = retrieve_X_out(X, W, K, S_out)
+    return S_out, W, K, X_out, distances
+
+def ica_sk(X, cycles=200, tol=1e-5):
+    ica = FastICA(max_iter=cycles, tol=tol)
+    S_out = ica.fit_transform(X.T).T
+    W = ica._unmixing
+    K = ica.whitening_
+    X_out = retrieve_X_out(X, W, K, S_out)
+    return S_out, W, K, X_out, []

### 2022-05-01 paper.tex

@@
 \begin{abstract}
-
+Indenpendent component analysis (ICA) is an unsupervised learning technique that is widely used in extracting independent factors in image, sound and medical signals.
+In this document, we review the basic notions of the ICA including mathematical formulations and detailed algorithms.
+We also implement different algorithms based on higher-order statistics and information theoretic approaches.
+These algorithms are applied to sound signal disentanglement and EEG blinking signal removal with success.
 \end{abstract}
@@
 \begin{equation}
-\frac{\partial H(\bf{y})}{\partial \bf{W}} = \frac{\partial (-I(\bf{y}))}{\partial \bf{W}} - \frac{\partial}{\partial \bf{W}}\sum_{i} \mathbb{E}[\log \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|} ]
+\frac{\partial H(\bf{y})}{\partial \bf{W}} = \frac{\partial (-I(\bf{y}))}{\partial \bf{W}} -\frac{\partial}{\partial \bf{W}}\sum_{i} \mathbb{E}[\log \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|} ].
 \end{equation}
@@
 Hence, the neural network approach can be eventually reduced to minimizing the mutual information $I(\bf{u})$.
-
-%In a neural network, there is a non-linear activation function that generates the output $\bf{y} = g(\bf{u})$.
-This nonlinearity indeed brings the information of higher order statistics into the optimization process.
 
@@
 \subsection{Information theoretic approach}
-
+Based on what has been discussed in the previous section, we can approximate the source signal distribution with a parametrized pdf.
+There are multiple choices here tailored to different problems.
+A default choice is to assume the cdf of the source signal can be approximated by a sigmoid function $g(s_{i}) = \frac{1}{1+e^{-s_{s}}}$, thus the pdf is $p(s_{i}) = g'(s_{i})$.
+Notice that this pdf is super-Gaussian.
+This means that ICA based on this pdf works well when the distribution of original sources have heavy tails and sharp peak around center.
+Recall the log-likelihood (without normalization),
+\begin{align}
+l(\bf{W}) &= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\bf{w}_{i}\bf{x}^{(j)}))+\log|\bf{W}|) \\
+&= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(g'(\bf{w}_{i}\bf{x}^{(j)}))+\log|\bf{W}|)
+\end{align}    
 \section{Applications}

### 2022-05-01 scripts/ex1.py
fix ex3

@@ -63,7 +63,7 @@ if __name__ == '__main__':
     S_out, W, K, X_out, distances = fast_ica.ica(X)
     print(W)
     plot_distances(distances, output='ex1_dist.pdf')
-
+    X_whiten = fast_ica.retrieve_whiten(X, W, K, S_out)
     S_out_sk, W, K, X_out_sk, distances = fast_ica.ica_sk(X)
     S_pca = PCA().fit_transform(X.T).T
 
@@ -71,14 +71,13 @@ if __name__ == '__main__':
         (S, 'Original sources ($S$)'),
         (X, 'Mixture signals ($X$)'),
         (fast_ica.whiten(X)[0], 'Whiten $X$'),
-        (S_out, '$S_{\\mathrm{predicted}}$ [Our FastICA]'),
+        (fast_ica.center(S_out, divide_sd=True), '$S_{\\mathrm{predicted}}$ [Our FastICA]'),
         (X_out, 'Retrieved $X = (W K)^{-1} S_{\\mathrm{predicted}}$ [Our FastICA]'),
+        (X_whiten, '$X_{whiten}$ [Our FastICA]'),
         (S_out_sk, 'Predicted $S$ [Sklearn FastICA]'),
         (X_out_sk, 'Retrieved $X = (W K)^{-1} S_{\\mathrm{predicted}}$ [Sklearn FastICA]'),
-        (S_pca, '$S_{\\mathrm{predicted}}$ [PCA]')
+        (fast_ica.center(S_pca, divide_sd=True), '$S_{\\mathrm{predicted}}$ [PCA]')
         ]
     sources, titles = zip(*source_title)
     plot_sources(sources, titles, output='ex1_sources.pdf')
     exit()
-
-

### 2022-05-01 scripts/ex3.py

@@ -26,7 +26,7 @@ def plot_eeg(d, d2=None, output='eeg.pdf'):
         if d2 is not None:
             ax[i, j].plot(d2[comp], '-', color=colors[1])
         ax[i, j].text(0.5, 1, f'{comp}', transform=ax[i, j].transAxes, fontsize=6)
-        # ax[i, j].set_yticks([])
+        ax[i, j].set_yticks([])
         ax[i, j].set_xticks([])
     fig.tight_layout()
     if output is None:
@@ -38,15 +38,13 @@ def remove_blinks():
     np.random.seed(0)
     X = loadmat('ex3_eeg.mat')['Data']
     plot_eeg(X, output='ex3_raw.pdf')
-    S_out, W, K, X_out, distances = fast_ica.ica(X)
+    S_out, W, K, X_out, distances = fast_ica.ica_sk(X)
     plot_eeg(S_out, output='ex3_ica.pdf')
     S_out_noblinks = S_out.copy()
-    S_out_noblinks[3, :] = np.zeros(X.shape[1])
+    S_out_noblinks[48, :] = np.zeros(X.shape[1])
     X_noblinks = fast_ica.retrieve_X_out(X, W, K, S_out_noblinks)
     plot_eeg(X_noblinks, output='ex3_noblinks.pdf')
     plot_eeg(X, d2=X_noblinks, output='ex3_noblinks.pdf')
-
-
 
 if __name__ == '__main__':
     remove_blinks()

### 2022-05-01 scripts/fast_ica.py

@@ -40,9 +40,16 @@ def retrieve_X_out(X, W, K, S_out):
     X_out = la.inv(W @ K) @ S_out + X.mean(axis=1, keepdims=True)
     return X_out
 
+def retrieve_whiten(X, W, K, S_out):
+    X_whiten = la.inv(W) @ S_out
+    print()
+    return X_whiten
+
 def ica(X, cycles=200, tol=1e-5, test=False):
     X1, K = whiten(X)
+    # X1 = center(X, divide_sd=True)
     nrows, ncols = X.shape
+    # K = np.eye(nrows)
     W = np.zeros((nrows, nrows))
     distances = []
     for i in range(nrows):

### 2022-05-02 paper.tex
add bib
edited the paper
edited the paper, sound figures

@@
 \section{Introduction}
-Independent component analysis (ICA) is one of the classical unsupervised learning techniques that is used to seperate out latent variables or reduce dimensions.
+Independent component analysis (ICA) is one of the classical unsupervised learning techniques that is used to separate out latent variables or reduce dimensions.
 The motivation of the development of this technique is based on very commom problems.
@@
 The famous example is the 'cocktail party problem', where the sound detectors record the superposition of various sound signals from different sources in the party.
-The goal is to seperate out the observed signal into sounds from each different source.
+The goal is to separate out the observed signal into sounds from each different source.
 Since we do not know what the sources are and how these signals are mixed, this type of problem is called 'blind source separation' (BSS).
-ICA is widely used in images, sounds, stock market and medical signals.
-ICA can also be viewed as an extension of principle component analysis (PCA), where the latter maximizes the second order statistics (covariance matrix of data).
+ICA is widely used in images, sounds, stock market and medical signals, where latent independent variables are believed to exist.
+ICA can also be viewed as an extension of principal component analysis (PCA), where the latter maximizes the second order statistics (covariance matrix of data).
 ICA maximizes higher-order statistics or simply tries to look for independent components not just uncorrelated components.
@@
 \begin{figure}[b]
-\includegraphics[width=0.8\textwidth]{ICA_vs_PCA.pdf}
+\includegraphics[width=0.8\textwidth]{scripts/ICA_vs_PCA.pdf}
  \caption{\label{fig:ICA_vs_PCA} shows the PCA and ICA result of unmixing the observed data $x_{i}$ generated from two independent uniform variables $s_{i}$.
@@
 We will in this section discuss various mathematical formulations.
-All these various theoretical frameworks are based on several constraints of the problem.
-These are:\\
+All these various theoretical frameworks are constrained by the following:\\
 (1) The number of sensors is larger than number of sources. This ensures that the mixing matrix is full rank; \\
@@
 The other ambiguity that matters for the practical application is that the original sources cannot all be distributed as Gaussians.
-Or to be more precise, in order to seperate the independent components, we require that the original signals can only have at most one component to be sampled from a Gaussian distribution, i.e.
+Or to be more precise, in order to separate the independent components, we require that the original signals can only have at most one component to be sampled from a Gaussian distribution, i.e.
 $s_{j}$ $\sim \mathcal{N}(\mu,\,\sigma^{2})$ for at most one $j$.
@@
 \begin{equation}
-\mathbb{E}_{\bf{s}}[\bf{x}] = \bf{A}\mathbb{E}[\bf{s}] = \bf{A}\bm{\mu}
+\mathbb{E}_{\mathbf{s}}[\mathbf{x}] = \mathbf{A}\mathbb{E}[\mathbf{s}] = \mathbf{A}\bm{\mu}
 \end{equation}
@@
 \begin{equation}
-Cov[\bf{x}] = \mathbb{E}_{\bf{s}}[\bf{As}\bf{s}^t\bf{A}^{t}]-\mathbb{E}_{\bf{s}}[\bf{As}]\mathbb{E}[(\bf{As})^{t}] =  \bf{A}\bm{\Sigma}\bf{A}^{t}-\bf{A}\bf{M}\bf{A}^{t}
-\end{equation} 
-, where $\bf{M}$ and $\bm{\Sigma}$ are diagonal matrices of $\rm I\!R^{d\times d}$ with diagonal elements to be \{$\mu_{j}^{2}$\} and \{$\sigma_{j}^{2}$\}, respectively.
+Cov[\mathbf{x}] = \mathbb{E}_{\mathbf{s}}[\mathbf{As}\mathbf{s}^t\mathbf{A}^{t}]-\mathbb{E}_{\mathbf{s}}[\mathbf{As}]\mathbb{E}[(\mathbf{As})^{t}] =  \mathbf{A}\bm{\Sigma}\mathbf{A}^{t}-\mathbf{A}\mathbf{M}\mathbf{A}^{t},
+\end{equation} where $\mathbf{M}$ and $\bm{\Sigma}$ are diagonal matrices of $\rm I\!R^{d\times d}$ with diagonal elements to be \{$\mu_{j}^{2}$\} and \{$\sigma_{j}^{2}$\}, respectively.
 Due to the scale ambiguity, this is the same as considering normally distributed signals with unity variance.
-Since each column $j$ of the mixing matrix $\bf{A}$ can be scaled by $\sigma_{j}$, $\tilde{\bf{A}}_{:j} = \sigma_{j}\bf{A}_{:j}$ and correspondingly the signal needs to be scaled by $1/\sigma_{j}$, $\tilde{s_{j}} = \frac{s_{j}-\mu_{j}}{\sigma_{j}}$.
+Since each column $j$ of the mixing matrix $\mathbf{A}$ can be scaled by $\sigma_{j}$, $\tilde{\mathbf{A}}_{:j} = \sigma_{j}\mathbf{A}_{:j}$ and correspondingly the signal needs to be scaled by $1/\sigma_{j}$, $\tilde{s_{j}} = \frac{s_{j}-\mu_{j}}{\sigma_{j}}$.
 Then,
 \begin{equation}
-Cov[\bf{x}] = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t}.
+Cov[\mathbf{x}] = \tilde{\mathbf{A}}\mathbf{I}_{d\times d}\tilde{\mathbf{A}}^{t}.
 \end{equation}
-Now, if we consider a rotation $\bf{R} $$\in O(d)$ acting on the scaled source signals $\bf{\tilde{s}}$.
-The mixing matrix then changes to $\tilde{\bf{A}}\bf{R}$.
-Unpon this rotation, we would observe $\bf{x'}$ as $\bf{x'} = \tilde{\bf{A}}\bf{R}\tilde{\bf{s}}$.
-$\bf{x'}$ is again normally distributed and has covariance matrix,
+Now, if we consider a rotation $\mathbf{R} $$\in O(d)$ acting on the scaled source signals $\mathbf{\tilde{s}}$.
+The mixing matrix then changes to $\tilde{\mathbf{A}}\mathbf{R}$.
+Unpon this rotation, we would observe $\mathbf{x'}$ as $\mathbf{x'} = \tilde{\mathbf{A}}\mathbf{R}\tilde{\mathbf{s}}$.
+$\mathbf{x'}$ is again normally distributed and has covariance matrix,
 \begin{equation}
-Cov[\bf{x'}] = \tilde{\bf{A}}\bf{R}\bf{I}_{d\times d}\bf{R}^{t}\tilde{\bf{A}}^{t} = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t}. 
-\end{equation} This simply means that whether the sources are rotated or not the observed data will be distributed as $\mathcal{N}(0, \tilde{\bf{A}}\tilde{\bf{A}}^{t})$.
+Cov[\mathbf{x'}] = \tilde{\mathbf{A}}\mathbf{R}\mathbf{I}_{d\times d}\mathbf{R}^{t}\tilde{\mathbf{A}}^{t} = \tilde{\mathbf{A}}\mathbf{I}_{d\times d}\tilde{\mathbf{A}}^{t}. 
+\end{equation} This simply means that whether the sources are rotated or not the observed data will be distributed as $\mathcal{N}(0, \tilde{\mathbf{A}}\tilde{\mathbf{A}}^{t})$.
 Thus, because of the rotational symmetry of the multivariate Gaussian distribution, we cannot seperate and obtain the original source signals.
-These derivations also indicate that given the observed data $\bf{x}$, we can whiten the data and it will not affect the seperated signals (up to scalings and mean translations).
+These derivations also indicate that given the observed data $\bf{x}$, we can whiten (apply PCA rotation and normalize each principal component) the data and it will not affect the seperated signals (up to scalings and mean translations).
+We will assume the observed data $\bf{x}$ are whitened if not explicitly stated in later discussions.
+That is $\tilde{\mathbf{x}} = \mathbf{\Sigma}^{-1/2}(\mathbf{x}-\mathbf{\mu})$ (whitening).
+Then, to recover the source signal, we need to find $\mathbf{W}$ with orthonormal rows \{$\mathbf{w}_{i}^{t}$\} such that $\mathbf{s} = \mathbf{W}\mathbf{\Sigma}^{-1/2}\mathbf{x}$.
 
@@
 The ICA tries to find independent components where the joint distibutions can be factorized into marginal distributions by investigating into higher-order moments.
-There are multiple moment-based objective functions have been used along the development of ICA.
-We will define some of the notations and stick to one of the moment-based objective functions in this paper.
+There are multiple moment-based objective functions that have been used along the development of ICA.
+We will define some of the notations and discuss one of the moment-based objective functions in detail.
 
-As we discussed previously, we have the freedom to whiten the observed signals and we will denote the whitened observed data as $\tilde{\bf{x}}$, where each component is $\tilde{x_{j}} = \frac{x_{j}-\mathbb{E}[x_{j}]}{\sigma_{x,j}}$.
-The skewness, kurtosis and excess kutosis are defined as $$ \gamma(x) := \mathbb{E}[\tilde{x}^{3}], \beta(x) := \mathbb{E}[\tilde{x}^{4}], \kappa(x) := \beta(x)-3 $$.
-The objective function is a convex combination of the squared third and fourth cumulants defined as follows,
+As we discussed previously, we will whiten the observed signals and we will denote the whitened observed data as $\bf{x}$.
+The kurtosis and excess kurtosis of whitended variable $x$ are defined as $$ \beta(x) := \mathbb{E}[x^{4}], \kappa(x) := \beta(x)-3 $$.
+The standard normal randon varible has excess kurtosis to be 0.
+When excess kurtosis $\kappa(x) < 0$, $x$ is said to be sub-Gaussian (flat around center, e.g.
+uniform distribution).
+When $\kappa(x) > 0$, $x$ is super-Gaussian (fat tail, sharp peak around center, e.g.
+Laplace distribution).
+Motivated by this the objective function to maximize can be $|\kappa(x)|$ and $\kappa^{2}(x)$ etc.
+This is furtherly backed up by the following inequalities:
 \begin{equation}
-L_{\alpha}(\bf{w}) := \alpha \gamma^{2}(\bf{w}^{t}\tilde{\bf{x}}) + (1-\alpha)\kappa^{2}(\bf{w}^{t}\tilde{\bf{x}})
-\end{equation},
-where $\alpha \in [0, 1]$ and $\bf{w}^{t}\bf{w}$$ = 1$.
-
-%1) Optimizing Kurtosis (high-order based moment approach) (Mick)
-%2) non-gaussianity  (Mick)
-%3) Maximum likelihood estimation (Kannan)
-%    - Minimizing mutual information (min KL divergence)
+|\mathbb{E}[(\mathbf{w}^{t}\mathbf{x})^{4}]-3| \leq max\{|\mathbb{E}[x_{1}^{4}]-3|,...,|\mathbb{E}[x_{d}^{4}]-3|\}
+\end{equation}
+\begin{equation}
+|\mathbb{E}[(\mathbf{w}_{1}^{t}\mathbf{x})^{4}-3]|+...+|\mathbb{E}[(\mathbf{w}_{d}^{t}\mathbf{x})^{4}-3]| \leq |\mathbb{E}[x_{1}^{4}]-3|+...+|\mathbb{E}[x_{d}^{4}]-3|
+\end{equation}
+Thus the problem is equivalent to finding an orthogonal transformation $\mathbf{W}$ with orthonormal rows \{$\mathbf{w}_{i}^{t}$\} such that the objective function can be maximized.
+In the case of the objective function $L(\mathbf{w}) = |\kappa(\mathbf{w}^{t}\mathbf{x})|$, this amounts to iterating the $\mathbf{w}$ with gradient,
+\begin{equation}
+\frac{\partial |\kappa(\mathbf{w}^{t}\mathbf{x})|}{\partial \mathbf{w}} = 4 sign(\kappa(\mathbf{w}^{t}\mathbf{x}))(\mathbb{E}[\mathbf{x}(\mathbf{w}^{t}\mathbf{x}))^{3}])-3\mathbf{w}||\mathbf{w}||^{2}).
+\end{equation} 
+This objective function captures both sub-Gaussian and super-Gaussian distributions for the source signals but the kurtosis is sensitive to the outliers of the observed data.
+Hence, some adaptive objective functions are used in practice more extensively based on negentropy ideas (to be covered later) such as \begin{align}
+G_{1}(u) &= \frac{1}{a_{1}}\log(\cosh a_{1}u) \\
+G_{2}(u) &= -\exp(-u^{2}/2), 
+\end{align} where $u = \mathbf{w}^{t}\mathbf{x}$. We will use $G_{1}$ for our implementation of one of the algorithms in later sections.
 
@@
 \begin{equation}
-I(\bf{x}) = \sum_{\bf{x}}p(\bf{x})\log\frac{p(\bf{x})}{\prod_{i}p_{i}(x_{i})}.
+I(\mathbf{x}) = \sum_{\mathbf{x}}p(\mathbf{x})\log\frac{p(\mathbf{x})}{\prod_{i}p_{i}(x_{i})}.
 \end{equation} $I(\bf{x})$ is non-negative and is 0 if and only if the $\bf{x}$ are independent.
@@
 
-One of them is a key finding in neural networks, that is, maximizing information between inputs $\bf{x}$ and outputs $\bf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\bf{y}$.
-Maximizing the information between inputs and outputs is to maximize the output joint entropy $H(\bf{y}) = \sum_{i}H(y_{i}) - I(\bf{y})$, where $I(\bf{y}) = \sum_{\bf{y}}p(\bf{y})\log\frac{p(\bf{y})}{\prod_{i}p_{i}(y_{i})}$ is the mutual information in the outputs $\bf{y}$.
-The mutual information is non-negative and is zero if and only if $\bf{y}$ are marginalized.
-Recall that in neural network, the output is given by the nonlinearity $y_{i} = g_{i}(\bf{w}_{i}^{t}\bf{x})$ $:= g_{i}(u_{i})$, thus $p(y_{i}) = \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|}$.
+One of them is a key finding in neural networks, that is, maximizing information between inputs $\mathbf{x}$ and outputs $\mathbf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\mathbf{y}$.
+Maximizing the information between inputs and outputs is to maximize the output joint entropy $H(\mathbf{y}) = \sum_{i}H(y_{i}) - I(\mathbf{y})$, where $I(\mathbf{y}) = \sum_{\mathbf{y}}p(\mathbf{y})\log\frac{p(\mathbf{y})}{\prod_{i}p_{i}(y_{i})}$ is the mutual information in the outputs $\mathbf{y}$.
+The mutual information is non-negative and is zero if and only if $\mathbf{y}$ are marginalized.
+Recall that in neural network, the output is given by the nonlinearity $y_{i} = g_{i}(\mathbf{w}_{i}^{t}\mathbf{x}) := g_{i}(u_{i})$, thus $p(y_{i}) = \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|}$.
 The maximum is therefore obtained by considering the gradient,
 \begin{equation}
-\frac{\partial H(\bf{y})}{\partial \bf{W}} = \frac{\partial (-I(\bf{y}))}{\partial \bf{W}} -\frac{\partial}{\partial \bf{W}}\sum_{i} \mathbb{E}[\log \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|} ].
+\frac{\partial H(\mathbf{y})}{\partial \mathbf{W}} = \frac{\partial (-I(\mathbf{y}))}{\partial \mathbf{W}} -\frac{\partial}{\partial \mathbf{W}}\sum_{i} \mathbb{E}[\log \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|} ].
 \end{equation}
-This implies that the nonlinear function $g_{i}$ in the neural network $y_{i} = g(\bf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$ in order to kill the second term.
+This implies that the nonlinear function $g_{i}$ in the neural network $y_{i} = g_{i}(\mathbf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$ in order to kill the second term.
 Together with the constraint that the outputs are marginalized, this gradient is zero.
 Therefore, a good estimation of the cdf of the source signal improves drastically the performance of ICA.
-Notice that, if $I(\bf{y}) = 0$, the $\bf{u} = \bf{W}\bf{x}$ should also satisfy $I(\bf{u}) = 0$ as $g_{i}$ is an invertible monotonic function.
-Hence, the neural network approach can be eventually reduced to minimizing the mutual information $I(\bf{u})$.
+Notice that, if $I(\mathbf{y}) = 0$, the $\mathbf{u} = \bf{W}\bf{x}$ should also satisfy $I(\mathbf{u}) = 0$ as $g_{i}$ is an invertible monotonic function.
+Hence, the neural network approach can be eventually reduced to minimizing the mutual information $I(\mathbf{u})$.
 
@@
 Recall that the Gaussian distribution has maximum entropy constrained with the mean and covariance.
-The negentropy thus measures non-Gaussianity.
+The negentropy thus measures non-Gaussianity, which is equivalent to higher order moment approach in principle.
 Requiring that the $\bf{u}$ can be factorized and decorrelated, the sum of negentropies can be written as
@@
 &= \sum_{i}p(u_{i})\log\frac{p(u_{i})}{p_{G}(u_{i})} \\
-&= \sum_{\bf{u}}p(\bf{u}) \log\frac{\prod_{i}p(u_{i})}{\prod_{i}p_{G}(u_{i})} \\
-&= \sum_{\bf{u}}p(\bf{u}) \log\frac{\prod_{i}p(u_{i})}{p_{G}(\bf{u}_{i})} \\
-&= \sum_{\bf{u}}p(\bf{u}) \log\frac{\prod_{i}p(u_{i})}{p(\bf{u})} + \sum_{\bf{u}}p(\bf{u})\log\frac{p(\bf{u})}{p_{G}(\bf{u})} \\
-&= D(\prod_{i}p(u_{i})||p(\bf{u})) + J(\bf{u}) \\
-&= -I(\bf{u}) + J(\bf{u}) \\
-& = -I(\bf{u}) - H(\bf{u}) - \sum_{\bf{u}}p(\bf{u})\log p_{G}(\bf{u}) \\
-& = -I(\bf{u}) - H(\bf{x}) - \log(|det(\bf{W})|) - \frac{1}{2}\log((2\pi e)^{d}det(<\bf{u}, \bf{u}^{t}>))\\
-&= -I(\bf{u}) - H(\bf{x}) - \frac{1}{2}\log((2\pi e)^{d}).
-\end{align} In the derivation, the $\bf{u}$ are uncorrelated so the covariance matrix is identity.
-Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\bf{u} = \bf{W}\bf{x}$.
+&= \sum_{\mathbf{u}}p(\mathbf{u}) \log\frac{\prod_{i}p(u_{i})}{\prod_{i}p_{G}(u_{i})} \\
+&= \sum_{\mathbf{u}}p(\mathbf{u}) \log\frac{\prod_{i}p(u_{i})}{p_{G}(\mathbf{u}_{i})} \\
+&= \sum_{\mathbf{u}}p(\mathbf{u}) \log\frac{\prod_{i}p(u_{i})}{p(\mathbf{u})} + \sum_{\mathbf{u}}p(\mathbf{u})\log\frac{p(\mathbf{u})}{p_{G}(\mathbf{u})} \\
+&= D(\prod_{i}p(u_{i})||p(\mathbf{u})) + J(\mathbf{u}) \\
+&= -I(\mathbf{u}) + J(\mathbf{u}) \\
+& = -I(\mathbf{u}) - H(\mathbf{u}) - \sum_{\mathbf{u}}p(\mathbf{u})\log p_{G}(\mathbf{u}) \\
+& = -I(\mathbf{u}) - H(\mathbf{x}) - \log(|det(\mathbf{W})|) - \frac{1}{2}\log((2\pi e)^{d}det(<\mathbf{u}, \mathbf{u}^{t}>))\\
+&= -I(\mathbf{u}) - H(\mathbf{x}) - \frac{1}{2}\log((2\pi e)^{d}).
+\end{align} In the derivation, the $\mathbf{u}$ are uncorrelated so the covariance matrix is identity.
+Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\mathbf{u} = \mathbf{W}\mathbf{x}$.
+In practice, the negentropy is difficult to evaluate so approximation schemes have been developed and have been discussed in the previous section.
 
-Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution $\hat{p}_{s}(\bf{w}_{i}\bf{x})$ satisfying $p(\bf{x}) = \prod_{i}^{d}\hat{p}_{s}(\bf{w}_{i}\bf{x})|\bf{W}|$.
+Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution $\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})$ satisfying $p(\mathbf{x}) = \prod_{i}^{d}\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})|\mathbf{W}|$.
 The log-likelihood is
 \begin{equation}
-l(\bf{W}) = \frac{1}{N}\sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\bf{w}_{i}\bf{x}^{(j)}))+\log|\bf{W}|).
+l(\mathbf{W}) = \frac{1}{N}\sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x}^{(j)}))+\log|\mathbf{W}|).
 \end{equation}   
-If the approximation $\hat{p}_{s}(\bf{w}_{i}\bf{x})$ are close to the actual pdf, the first term is approximately $-\sum_{i}H(\bf{w}_{i}\bf{x})$ and is equal to the negative of the mutual information up to an additive constant of the total entropy of $\bf{x}$.
+If the approximation $\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})$ are close to the actual pdf, the first term is approximately $-\sum_{i}H(\mathbf{w}_{i}\mathbf{x})$ and is equal to the negative of the mutual information up to an additive constant of the total entropy of $\mathbf{x}$.
 Hence, all these historical information theoretic approaches are equivalent and the key idea is to approximate the probability distribution of the sources correctly.
@@
 In function forms of $G_1$ and $G_2$ should not grow too fast to be robust.
-The following choice of $G$ is proposed by Hyvärinen and Oja.
+The following choice of $G$ is proposed by Hyvärinen and Oja~\cite{hyvarinen2000independent}.
 \begin{align}
@@
 \begin{align}
-l(\bf{W}) &= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\bf{w}_{i}\bf{x}^{(j)}))+\log|\bf{W}|) \\
-&= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(g'(\bf{w}_{i}\bf{x}^{(j)}))+\log|\bf{W}|)
-\end{align}    
+l(\mathbf{W}) &= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\mathbf{w}_{i}^{t}\mathbf{x}^{(j)}))+\log|\mathbf{W}|) \\
+&= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(g'(\mathbf{w}_{i}^{t}\mathbf{x}^{(j)}))+\log|\mathbf{W}|),
+\end{align} where $g'(\mathbf{w}_{i}^{t}\mathbf{x}) = g'(\mathbf{u}_{i}) = g(1-g)$. Thus the gradient of log-likelihood is 
+\begin{equation}
+\frac{\partial l(\mathbf{W})}{\partial \mathbf{W}} = \sum_{j=1}^{N}(\mathbf{q}^{(j)}\mathbf{x}^{t, (j)} + (\mathbf{W}^{t})^{-1}),
+\end{equation} where $\mathbf{q}^{t, (j)} = [1-2g(u_{1}^{(j)}), ... ,1-2g(u_{d}^{(j)}) ]$. 
+In our implementation, we used the stochastic gradient descent method to update the $\mathbf{W}$:
+\begin{equation}
+\mathbf{W} = \mathbf{W} + \alpha (\mathbf{q}^{(j)}\mathbf{x}^{t, (j)} + (\mathbf{W}^{t})^{-1}).
+\end{equation} This means that we update the $\mathbf{W}$ for each sample of the observed data instead of calculating the exact summation for all samples at each step of iteration.
+This iteration schemes is accompanied by randomize the ordering of samples initially for each iteration (making the iteration more stochastic).
 \section{Applications}
+\subsection{Sklearn example}
 We implement FastICA according to the formalism described in the previous section.
@@
 The result from our implementation and FastICA are shown in Fig.~\ref{fig:sklearn}.
+\subsection{Sound waves}
+We employed the maximum likelihood algorithm to separate out the independent components of five mixed sound tracks (data obtained from Standford cs229 online course materials).
+The mixed sound waves are shown in the  Fig.~\ref{fig:sound} left panels (red).
+The separated independent sound tracks are shown in the right panels (blue) in Fig.~\ref{fig:sound}.
+The separated sound tracks can be played and listened one by one to verify that they are meaningful.
+\begin{figure}[b]
+\includegraphics[width=0.8\textwidth]{scripts/sound_example.pdf}
+ \caption{\label{fig:sound} shows the ICA results of separating the five mixed sound waves in the left panels (red).
+The separated independent components are shown in the right panels (blue).
+When the separated signals are played, one can clearly recognize these meaningful audio signals.}
+\end{figure}
+\subsection{Blinking removal in EEG signal}
 % \begin{figure}
@@
 
-%3) example: sound single decomposition
-%4) EEG % (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
-%5) stock price
-
-\section{How to check the robustness and convergence}
+%\section{How to check the robustness and convergence}
 % Amari metric, ref "ELements of Statistical Learning", p.570
@@
 %1) check if the performance matrix is identity (Te-won Lee, p. 51)
-\begin{align}
-W A &= I.
-\end{align}
+%\begin{align}
+%W A &= I.
+%\end{align}
 
@@
 
-\section{what if two of the sources are dependent}
+%\section{what if two of the sources are dependent}
 %s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
@@
 \section{Discussions and Conclusions}
-1) review the traditional development from 1990s
-2) recent advances and development from 2000s  
-3) what happens when there are multiple Gaussian components
+In this review paper of ICA, we have discussed the traditional and basic notions of this technique.
+We have also implemented the algorithms and utilized our codes to solve several different examples.
+We have noticed that the objective function plays an important role in the performance of ICA in various examples we examined.
+In the basic setting, the performance of the algorithm lies in choosing the correct and robust objective function.
+Several objective functions have been proposed to improve the robusteness and computation efficiency.
+However, if we know in advance what the probability distributions look like for the sources (latent variables), it would always be the best to tailor the objective function to include this information.
+Nonetheless, ICA has already shown reliable performance in examples such as sound waves identification and EEG signals separation using some default robust objective functions.
+The current research in this field focuses on cases where the number of detectors smaller than the number of sources, and non-linear mixing problems.
+
+\bibliography{paper}{}
 

### 2022-05-02 paper.tex
added the reference

@@
 The goal is to separate out the observed signal into sounds from each different source.
-Since we do not know what the sources are and how these signals are mixed, this type of problem is called 'blind source separation' (BSS).
+Since we do not know what the sources are and how these signals are mixed, this type of problem is called 'blind source separation' (BSS) \cite{lee1998independent}.
 ICA is widely used in images, sounds, stock market and medical signals, where latent independent variables are believed to exist.
@@
 We will in this section discuss various mathematical formulations.
-All these various theoretical frameworks are constrained by the following:\\
+All these various theoretical frameworks are constrained by the following \cite{lee1998independent}:\\
 (1) The number of sensors is larger than number of sources. This ensures that the mixing matrix is full rank; \\
@@
 That is, for a particular component $j$, if we scale the component by $c_{j}$ the corrsponding column of the mixing matrix can be scaled by $1/c_{j}$ to have the observed data unchanged (i.e.
-$x_{i} = \sum_{j}\frac{1}{c_{j}}A_{ij}c_{j}s_{j}$).
+$x_{i} = \sum_{j}\frac{1}{c_{j}}A_{ij}c_{j}s_{j}$) \cite{ng_cs229}.
 
-The other ambiguity that matters for the practical application is that the original sources cannot all be distributed as Gaussians.
+The other ambiguity that matters for the practical application is that the original sources cannot all be distributed as Gaussians \cite{lee1998independent, ng_cs229}.
 Or to be more precise, in order to separate the independent components, we require that the original signals can only have at most one component to be sampled from a Gaussian distribution, i.e.
@@
 Motivated by this the objective function to maximize can be $|\kappa(x)|$ and $\kappa^{2}(x)$ etc.
-This is furtherly backed up by the following inequalities:
+This is furtherly backed up by the following inequalities \cite{miettinen2015fourth}:
 \begin{equation}
@@
 Thus the problem is equivalent to finding an orthogonal transformation $\mathbf{W}$ with orthonormal rows \{$\mathbf{w}_{i}^{t}$\} such that the objective function can be maximized.
-In the case of the objective function $L(\mathbf{w}) = |\kappa(\mathbf{w}^{t}\mathbf{x})|$, this amounts to iterating the $\mathbf{w}$ with gradient,
+In the case of the objective function $L(\mathbf{w}) = |\kappa(\mathbf{w}^{t}\mathbf{x})|$, this amounts to iterating the $\mathbf{w}$ with gradient \citep{hyvarinen2000independent},
 \begin{equation}
@@
 G_{2}(u) &= -\exp(-u^{2}/2), 
-\end{align} where $u = \mathbf{w}^{t}\mathbf{x}$. We will use $G_{1}$ for our implementation of one of the algorithms in later sections.
+\end{align} where $u = \mathbf{w}^{t}\mathbf{x}$ \citep{hyvarinen2000independent}. We will use $G_{1}$ for our implementation of one of the algorithms in later sections.
 
@@
 Notice that, if $I(\mathbf{y}) = 0$, the $\mathbf{u} = \bf{W}\bf{x}$ should also satisfy $I(\mathbf{u}) = 0$ as $g_{i}$ is an invertible monotonic function.
-Hence, the neural network approach can be eventually reduced to minimizing the mutual information $I(\mathbf{u})$.
+Hence, the neural network approach can be eventually reduced to minimizing the mutual information $I(\mathbf{u})$ \cite{lee1998independent}.
 
@@
 \end{align} In the derivation, the $\mathbf{u}$ are uncorrelated so the covariance matrix is identity.
-Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\mathbf{u} = \mathbf{W}\mathbf{x}$.
+Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\mathbf{u} = \mathbf{W}\mathbf{x}$ \cite{lee1998independent}.
 In practice, the negentropy is difficult to evaluate so approximation schemes have been developed and have been discussed in the previous section.
 
-Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution $\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})$ satisfying $p(\mathbf{x}) = \prod_{i}^{d}\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})|\mathbf{W}|$.
+Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution $\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})$ satisfying $p(\mathbf{x}) = \prod_{i=1}^{d}\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})|\mathbf{W}|$ \cite{ng_cs229}.
 The log-likelihood is
@@
 This means that ICA based on this pdf works well when the distribution of original sources have heavy tails and sharp peak around center.
-Recall the log-likelihood (without normalization),
+Recall the log-likelihood (without normalization) \cite{ng_cs229},
 \begin{align}
@@
 \subsection{Sklearn example}
-We implement FastICA according to the formalism described in the previous section.
-The algorithm is applied to a following set of signal functions.
+We implement the moment-based approach according to the formalism described in the previous section.
+The algorithm is applied to a following set of signal functions (example from sklearn module):
 \begin{align}
-s_1 &= \sin(2 t). \\
-s_2 &= \mathrm{sign} [\sin(3 t)]. \\
+s_1 &= \sin(2 t) \\
+s_2 &= \mathrm{sign} [\sin(3 t)] \\
 s_3 &= \frac{1}{2\pi} (t~\mathrm{mod}~2\pi).
@@
 \end{align}
-The result from our implementation and FastICA are shown in Fig.~\ref{fig:sklearn}.
+The result from our implementation and sklearn module are shown in Fig.~\ref{fig:sklearn}.
 \subsection{Sound waves}
@@
 Nonetheless, ICA has already shown reliable performance in examples such as sound waves identification and EEG signals separation using some default robust objective functions.
-The current research in this field focuses on cases where the number of detectors smaller than the number of sources, and non-linear mixing problems.
+The ongoing research in this field focuses on cases where the number of detectors smaller than the number of sources, and non-linear mixing problems.
 

### 2022-05-02 paper.tex
edited the boldface style in equations

@@
 \begin{align}
-J(Y) \propto \{E[G(Y)] - E[G(V)]\}^2,
+J(Y) \propto \{\mathbb{E}[G(Y)] - \mathbb{E}[G(V)]\}^2,
 \end{align}
@@
 \end{align}
-We want to find the extrema of $E[G(Y)]$ to maximize $J(Y)$ under the constraints
+We want to find the extrema of $\mathbb{E}[G(Y)]$ to maximize $J(Y)$ under the constraints
 \begin{align}
-E[(w^T x)^2] = \| w\|^2 = 1.
+\mathbb{E}(\mathbf{w}^{t}\mathbf{x})^2] = \| \mathbf{w}\|^2 = 1.
 \end{align}
@@
 \begin{align}
-E[x g(w^Tx)] - \beta w &= 0.
+\mathbb{E}[\mathbf{x}g(\mathbf{w}^{t}\mathbf{x})^2)] - \beta\mathbf{w} &= \mathbf{0}.
 \end{align}
 This equation can be solved by using Newton's method.
-Let $F(w)$ denote the left-hand side of the equation, which we are trying to solve.
-We update the value of $w$ in each iteration according to the following equation.
+Let $\mathbf{F}(\mathbf{w})$ denote the left-hand side of the equation, which we are trying to solve.
+We update the value of $\mathbf{w}$ in each iteration according to the following equation.
 \begin{align}
-w_{n+1} &= w_n - \frac{F(w_n)}{F^{\prime}(w_n)} \\
-&= w_n - \frac{E[x g (w_n^T x)] - \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
-&= \frac{w_n E[g^{\prime}(w_n^T x)] - \beta w_n - E[x g (w_n^T x)] + \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
-&= \frac{w_n E[g^{\prime}(w_n^T x)] - E[x g (w_n^T x)]}{E[g^{\prime}(w_n^T x)] - \beta}.
+\mathbf{w}_{n+1} &= \mathbf{w}_n - \frac{\mathbf{F}(\mathbf{w}_n)}{F^{\prime}(\mathbf{w}_n)} \\
+&= \mathbf{w}_n - \frac{\mathbb{E}[\mathbf{x}g(\mathbf{w}_n^t\mathbf{x})] - \beta \mathbf{w}_n}{\mathbb{E}[g^{\prime}(\mathbf{w}_n^t\mathbf{x})] - \beta} \\
+&= \frac{\mathbf{w}_n\mathbb{E}[g^{\prime}(\mathbf{w}_n^t\mathbf{x})] - \beta\mathbf{w}_n - \mathbb{E}[\mathbf{x}g(\mathbf{w}_n^t\mathbf{x})] + \beta\mathbf{w}_n}{\mathbb{E}[g^{\prime}(\mathbf{w}_n^t\mathbf{x})] - \beta} \\
+&= \frac{\mathbf{w}_n\mathbb{E}[g^{\prime}(\mathbf{w}_n^t\mathbf{x})] - \mathbb{E}[\mathbf{x}g(\mathbf{w}_n^t\mathbf{x})]}{\mathbb{E}[g^{\prime}(\mathbf{w}_n^t\mathbf{x})] - \beta}.
 \end{align}
@@
 \begin{align}
-A &= \begin{bmatrix}
+\mathbf{A} &= \begin{bmatrix}
 1 & 1 & 1 \\

### 2022-05-03 paper.tex
add applications section

--- 2022-05-02 paper.tex
@@
 
-%Title of paper
 \title{Independent Component Analysis}
@@
 \section{Introduction}
-Independent component analysis (ICA) is one of the classical unsupervised learning techniques that is used to separate out latent variables or reduce dimensions.
+Independent component analysis (ICA) is one of the classical unsupervised learning techniques that is used to seperate out latent variables or reduce dimensions.
 The motivation of the development of this technique is based on very commom problems.
@@
 The famous example is the 'cocktail party problem', where the sound detectors record the superposition of various sound signals from different sources in the party.
-The goal is to separate out the observed signal into sounds from each different source.
-Since we do not know what the sources are and how these signals are mixed, this type of problem is called 'blind source separation' (BSS).
-ICA is widely used in images, sounds, stock market and medical signals, where latent independent variables are believed to exist.
-ICA can also be viewed as an extension of principal component analysis (PCA), where the latter maximizes the second order statistics (covariance matrix of data).
+The goal is to seperate out the observed signal into sounds from each different source.
+Since we do not know what the sources are and how these signals are mixed, this type of problem is called blind source separation (BSS).
+ICA is widely used in images, sounds, stock market and medical signals.
+ICA can also be viewed as an extension of principle component analysis (PCA), where the latter maximizes the second order statistics (covariance matrix of data).
 ICA maximizes higher-order statistics or simply tries to look for independent components not just uncorrelated components.
@@
 We will in this section discuss various mathematical formulations.
-All these various theoretical frameworks are constrained by the following:\\
+All these various theoretical frameworks are based on several constraints of the problem.
+These are:\\
 (1) The number of sensors is larger than number of sources. This ensures that the mixing matrix is full rank; \\
@@
 The other ambiguity that matters for the practical application is that the original sources cannot all be distributed as Gaussians.
-Or to be more precise, in order to separate the independent components, we require that the original signals can only have at most one component to be sampled from a Gaussian distribution, i.e.
+Or to be more precise, in order to seperate the independent components, we require that the original signals can only have at most one component to be sampled from a Gaussian distribution, i.e.
 $s_{j}$ $\sim \mathcal{N}(\mu,\,\sigma^{2})$ for at most one $j$.
@@
 \begin{equation}
-\mathbb{E}_{\mathbf{s}}[\mathbf{x}] = \mathbf{A}\mathbb{E}[\mathbf{s}] = \mathbf{A}\bm{\mu}
+\mathbb{E}_{\bf{s}}[\bf{x}] = \bf{A}\mathbb{E}[\bf{s}] = \bf{A}\bm{\mu}
 \end{equation}
@@
 \begin{equation}
-Cov[\mathbf{x}] = \mathbb{E}_{\mathbf{s}}[\mathbf{As}\mathbf{s}^t\mathbf{A}^{t}]-\mathbb{E}_{\mathbf{s}}[\mathbf{As}]\mathbb{E}[(\mathbf{As})^{t}] =  \mathbf{A}\bm{\Sigma}\mathbf{A}^{t}-\mathbf{A}\mathbf{M}\mathbf{A}^{t},
-\end{equation} where $\mathbf{M}$ and $\bm{\Sigma}$ are diagonal matrices of $\rm I\!R^{d\times d}$ with diagonal elements to be \{$\mu_{j}^{2}$\} and \{$\sigma_{j}^{2}$\}, respectively.
+Cov[\bf{x}] = \mathbb{E}_{\bf{s}}[\bf{As}\bf{s}^t\bf{A}^{t}]-\mathbb{E}_{\bf{s}}[\bf{As}]\mathbb{E}[(\bf{As})^{t}] =  \bf{A}\bm{\Sigma}\bf{A}^{t}-\bf{A}\bf{M}\bf{A}^{t}
+\end{equation} 
+, where $\bf{M}$ and $\bm{\Sigma}$ are diagonal matrices of $\rm I\!R^{d\times d}$ with diagonal elements to be \{$\mu_{j}^{2}$\} and \{$\sigma_{j}^{2}$\}, respectively.
 Due to the scale ambiguity, this is the same as considering normally distributed signals with unity variance.
-Since each column $j$ of the mixing matrix $\mathbf{A}$ can be scaled by $\sigma_{j}$, $\tilde{\mathbf{A}}_{:j} = \sigma_{j}\mathbf{A}_{:j}$ and correspondingly the signal needs to be scaled by $1/\sigma_{j}$, $\tilde{s_{j}} = \frac{s_{j}-\mu_{j}}{\sigma_{j}}$.
+Since each column $j$ of the mixing matrix $\bf{A}$ can be scaled by $\sigma_{j}$, $\tilde{\bf{A}}_{:j} = \sigma_{j}\bf{A}_{:j}$ and correspondingly the signal needs to be scaled by $1/\sigma_{j}$, $\tilde{s_{j}} = \frac{s_{j}-\mu_{j}}{\sigma_{j}}$.
 Then,
 \begin{equation}
-Cov[\mathbf{x}] = \tilde{\mathbf{A}}\mathbf{I}_{d\times d}\tilde{\mathbf{A}}^{t}.
+Cov[\bf{x}] = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t}.
 \end{equation}
-Now, if we consider a rotation $\mathbf{R} $$\in O(d)$ acting on the scaled source signals $\mathbf{\tilde{s}}$.
-The mixing matrix then changes to $\tilde{\mathbf{A}}\mathbf{R}$.
-Unpon this rotation, we would observe $\mathbf{x'}$ as $\mathbf{x'} = \tilde{\mathbf{A}}\mathbf{R}\tilde{\mathbf{s}}$.
-$\mathbf{x'}$ is again normally distributed and has covariance matrix,
+Now, if we consider a rotation $\bf{R} $$\in O(d)$ acting on the scaled source signals $\bf{\tilde{s}}$.
+The mixing matrix then changes to $\tilde{\bf{A}}\bf{R}$.
+Unpon this rotation, we would observe $\bf{x'}$ as $\bf{x'} = \tilde{\bf{A}}\bf{R}\tilde{\bf{s}}$.
+$\bf{x'}$ is again normally distributed and has covariance matrix,
 \begin{equation}
-Cov[\mathbf{x'}] = \tilde{\mathbf{A}}\mathbf{R}\mathbf{I}_{d\times d}\mathbf{R}^{t}\tilde{\mathbf{A}}^{t} = \tilde{\mathbf{A}}\mathbf{I}_{d\times d}\tilde{\mathbf{A}}^{t}. 
-\end{equation} This simply means that whether the sources are rotated or not the observed data will be distributed as $\mathcal{N}(0, \tilde{\mathbf{A}}\tilde{\mathbf{A}}^{t})$.
+Cov[\bf{x'}] = \tilde{\bf{A}}\bf{R}\bf{I}_{d\times d}\bf{R}^{t}\tilde{\bf{A}}^{t} = \tilde{\bf{A}}\bf{I}_{d\times d}\tilde{\bf{A}}^{t}. 
+\end{equation} This simply means that whether the sources are rotated or not the observed data will be distributed as $\mathcal{N}(0, \tilde{\bf{A}}\tilde{\bf{A}}^{t})$.
 Thus, because of the rotational symmetry of the multivariate Gaussian distribution, we cannot seperate and obtain the original source signals.
-These derivations also indicate that given the observed data $\bf{x}$, we can whiten (apply PCA rotation and normalize each principal component) the data and it will not affect the seperated signals (up to scalings and mean translations).
-We will assume the observed data $\bf{x}$ are whitened if not explicitly stated in later discussions.
-That is $\tilde{\mathbf{x}} = \mathbf{\Sigma}^{-1/2}(\mathbf{x}-\mathbf{\mu})$ (whitening).
-Then, to recover the source signal, we need to find $\mathbf{W}$ with orthonormal rows \{$\mathbf{w}_{i}^{t}$\} such that $\mathbf{s} = \mathbf{W}\mathbf{\Sigma}^{-1/2}\mathbf{x}$.
+These derivations also indicate that given the observed data $\bf{x}$, we can whiten the data and it will not affect the seperated signals (up to scalings and mean translations).
 
@@
 The ICA tries to find independent components where the joint distibutions can be factorized into marginal distributions by investigating into higher-order moments.
-There are multiple moment-based objective functions that have been used along the development of ICA.
-We will define some of the notations and discuss one of the moment-based objective functions in detail.
+There are multiple moment-based objective functions have been used along the development of ICA.
+We will define some of the notations and stick to one of the moment-based objective functions in this paper.
 
-As we discussed previously, we will whiten the observed signals and we will denote the whitened observed data as $\bf{x}$.
-The kurtosis and excess kurtosis of whitended variable $x$ are defined as $$ \beta(x) := \mathbb{E}[x^{4}], \kappa(x) := \beta(x)-3 $$.
-The standard normal randon varible has excess kurtosis to be 0.
-When excess kurtosis $\kappa(x) < 0$, $x$ is said to be sub-Gaussian (flat around center, e.g.
-uniform distribution).
-When $\kappa(x) > 0$, $x$ is super-Gaussian (fat tail, sharp peak around center, e.g.
-Laplace distribution).
-Motivated by this the objective function to maximize can be $|\kappa(x)|$ and $\kappa^{2}(x)$ etc.
-This is furtherly backed up by the following inequalities:
+As we discussed previously, we have the freedom to whiten the observed signals and we will denote the whitened observed data as $\tilde{\bf{x}}$, where each component is $\tilde{x_{j}} = \frac{x_{j}-\mathbb{E}[x_{j}]}{\sigma_{x,j}}$.
+The skewness, kurtosis and excess kutosis are defined as $$ \gamma(x) := \mathbb{E}[\tilde{x}^{3}], \beta(x) := \mathbb{E}[\tilde{x}^{4}], \kappa(x) := \beta(x)-3 $$.
+The objective function is a convex combination of the squared third and fourth cumulants defined as follows,
 \begin{equation}
-|\mathbb{E}[(\mathbf{w}^{t}\mathbf{x})^{4}]-3| \leq max\{|\mathbb{E}[x_{1}^{4}]-3|,...,|\mathbb{E}[x_{d}^{4}]-3|\}
-\end{equation}
-\begin{equation}
-|\mathbb{E}[(\mathbf{w}_{1}^{t}\mathbf{x})^{4}-3]|+...+|\mathbb{E}[(\mathbf{w}_{d}^{t}\mathbf{x})^{4}-3]| \leq |\mathbb{E}[x_{1}^{4}]-3|+...+|\mathbb{E}[x_{d}^{4}]-3|
-\end{equation}
-Thus the problem is equivalent to finding an orthogonal transformation $\mathbf{W}$ with orthonormal rows \{$\mathbf{w}_{i}^{t}$\} such that the objective function can be maximized.
-In the case of the objective function $L(\mathbf{w}) = |\kappa(\mathbf{w}^{t}\mathbf{x})|$, this amounts to iterating the $\mathbf{w}$ with gradient,
-\begin{equation}
-\frac{\partial |\kappa(\mathbf{w}^{t}\mathbf{x})|}{\partial \mathbf{w}} = 4 sign(\kappa(\mathbf{w}^{t}\mathbf{x}))(\mathbb{E}[\mathbf{x}(\mathbf{w}^{t}\mathbf{x}))^{3}])-3\mathbf{w}||\mathbf{w}||^{2}).
-\end{equation} 
-This objective function captures both sub-Gaussian and super-Gaussian distributions for the source signals but the kurtosis is sensitive to the outliers of the observed data.
-Hence, some adaptive objective functions are used in practice more extensively based on negentropy ideas (to be covered later) such as \begin{align}
-G_{1}(u) &= \frac{1}{a_{1}}\log(\cosh a_{1}u) \\
-G_{2}(u) &= -\exp(-u^{2}/2), 
-\end{align} where $u = \mathbf{w}^{t}\mathbf{x}$. We will use $G_{1}$ for our implementation of one of the algorithms in later sections.
+L_{\alpha}(\bf{w}) := \alpha \gamma^{2}(\bf{w}^{t}\tilde{\bf{x}}) + (1-\alpha)\kappa^{2}(\bf{w}^{t}\tilde{\bf{x}})
+\end{equation},
+where $\alpha \in [0, 1]$ and $\bf{w}^{t}\bf{w}$$ = 1$.
+
+%1) Optimizing Kurtosis (high-order based moment approach) (Mick)
+%2) non-gaussianity  (Mick)
+%3) Maximum likelihood estimation (Kannan)
+%    - Minimizing mutual information (min KL divergence)
 
@@
 \begin{equation}
-I(\mathbf{x}) = \sum_{\mathbf{x}}p(\mathbf{x})\log\frac{p(\mathbf{x})}{\prod_{i}p_{i}(x_{i})}.
+I(\bf{x}) = \sum_{\bf{x}}p(\bf{x})\log\frac{p(\bf{x})}{\prod_{i}p_{i}(x_{i})}.
 \end{equation} $I(\bf{x})$ is non-negative and is 0 if and only if the $\bf{x}$ are independent.
@@
 
-One of them is a key finding in neural networks, that is, maximizing information between inputs $\mathbf{x}$ and outputs $\mathbf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\mathbf{y}$.
-Maximizing the information between inputs and outputs is to maximize the output joint entropy $H(\mathbf{y}) = \sum_{i}H(y_{i}) - I(\mathbf{y})$, where $I(\mathbf{y}) = \sum_{\mathbf{y}}p(\mathbf{y})\log\frac{p(\mathbf{y})}{\prod_{i}p_{i}(y_{i})}$ is the mutual information in the outputs $\mathbf{y}$.
-The mutual information is non-negative and is zero if and only if $\mathbf{y}$ are marginalized.
-Recall that in neural network, the output is given by the nonlinearity $y_{i} = g_{i}(\mathbf{w}_{i}^{t}\mathbf{x}) := g_{i}(u_{i})$, thus $p(y_{i}) = \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|}$.
+One of them is a key finding in neural networks, that is, maximizing information between inputs $\bf{x}$ and outputs $\bf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\bf{y}$.
+Maximizing the information between inputs and outputs is to maximize the output joint entropy $H(\bf{y}) = \sum_{i}H(y_{i}) - I(\bf{y})$, where $I(\bf{y}) = \sum_{\bf{y}}p(\bf{y})\log\frac{p(\bf{y})}{\prod_{i}p_{i}(y_{i})}$ is the mutual information in the outputs $\bf{y}$.
+The mutual information is non-negative and is zero if and only if $\bf{y}$ are marginalized.
+Recall that in neural network, the output is given by the nonlinearity $y_{i} = g_{i}(\bf{w}_{i}^{t}\bf{x})$ $:= g_{i}(u_{i})$, thus $p(y_{i}) = \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|}$.
 The maximum is therefore obtained by considering the gradient,
 \begin{equation}
-\frac{\partial H(\mathbf{y})}{\partial \mathbf{W}} = \frac{\partial (-I(\mathbf{y}))}{\partial \mathbf{W}} -\frac{\partial}{\partial \mathbf{W}}\sum_{i} \mathbb{E}[\log \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|} ].
+\frac{\partial H(\bf{y})}{\partial \bf{W}} = \frac{\partial (-I(\bf{y}))}{\partial \bf{W}} -\frac{\partial}{\partial \bf{W}}\sum_{i} \mathbb{E}[\log \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|} ].
 \end{equation}
-This implies that the nonlinear function $g_{i}$ in the neural network $y_{i} = g_{i}(\mathbf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$ in order to kill the second term.
+This implies that the nonlinear function $g_{i}$ in the neural network $y_{i} = g(\bf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$ in order to kill the second term.
 Together with the constraint that the outputs are marginalized, this gradient is zero.
 Therefore, a good estimation of the cdf of the source signal improves drastically the performance of ICA.
-Notice that, if $I(\mathbf{y}) = 0$, the $\mathbf{u} = \bf{W}\bf{x}$ should also satisfy $I(\mathbf{u}) = 0$ as $g_{i}$ is an invertible monotonic function.
-Hence, the neural network approach can be eventually reduced to minimizing the mutual information $I(\mathbf{u})$.
+Notice that, if $I(\bf{y}) = 0$, the $\bf{u} = \bf{W}\bf{x}$ should also satisfy $I(\bf{u}) = 0$ as $g_{i}$ is an invertible monotonic function.
+Hence, the neural network approach can be eventually reduced to minimizing the mutual information $I(\bf{u})$.
 
@@
 Recall that the Gaussian distribution has maximum entropy constrained with the mean and covariance.
-The negentropy thus measures non-Gaussianity, which is equivalent to higher order moment approach in principle.
+The negentropy thus measures non-Gaussianity.
 Requiring that the $\bf{u}$ can be factorized and decorrelated, the sum of negentropies can be written as
@@
 &= \sum_{i}p(u_{i})\log\frac{p(u_{i})}{p_{G}(u_{i})} \\
-&= \sum_{\mathbf{u}}p(\mathbf{u}) \log\frac{\prod_{i}p(u_{i})}{\prod_{i}p_{G}(u_{i})} \\
-&= \sum_{\mathbf{u}}p(\mathbf{u}) \log\frac{\prod_{i}p(u_{i})}{p_{G}(\mathbf{u}_{i})} \\
-&= \sum_{\mathbf{u}}p(\mathbf{u}) \log\frac{\prod_{i}p(u_{i})}{p(\mathbf{u})} + \sum_{\mathbf{u}}p(\mathbf{u})\log\frac{p(\mathbf{u})}{p_{G}(\mathbf{u})} \\
-&= D(\prod_{i}p(u_{i})||p(\mathbf{u})) + J(\mathbf{u}) \\
-&= -I(\mathbf{u}) + J(\mathbf{u}) \\
-& = -I(\mathbf{u}) - H(\mathbf{u}) - \sum_{\mathbf{u}}p(\mathbf{u})\log p_{G}(\mathbf{u}) \\
-& = -I(\mathbf{u}) - H(\mathbf{x}) - \log(|det(\mathbf{W})|) - \frac{1}{2}\log((2\pi e)^{d}det(<\mathbf{u}, \mathbf{u}^{t}>))\\
-&= -I(\mathbf{u}) - H(\mathbf{x}) - \frac{1}{2}\log((2\pi e)^{d}).
-\end{align} In the derivation, the $\mathbf{u}$ are uncorrelated so the covariance matrix is identity.
-Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\mathbf{u} = \mathbf{W}\mathbf{x}$.
-In practice, the negentropy is difficult to evaluate so approximation schemes have been developed and have been discussed in the previous section.
+&= \sum_{\bf{u}}p(\bf{u}) \log\frac{\prod_{i}p(u_{i})}{\prod_{i}p_{G}(u_{i})} \\
+&= \sum_{\bf{u}}p(\bf{u}) \log\frac{\prod_{i}p(u_{i})}{p_{G}(\bf{u}_{i})} \\
+&= \sum_{\bf{u}}p(\bf{u}) \log\frac{\prod_{i}p(u_{i})}{p(\bf{u})} + \sum_{\bf{u}}p(\bf{u})\log\frac{p(\bf{u})}{p_{G}(\bf{u})} \\
+&= D(\prod_{i}p(u_{i})||p(\bf{u})) + J(\bf{u}) \\
+&= -I(\bf{u}) + J(\bf{u}) \\
+& = -I(\bf{u}) - H(\bf{u}) - \sum_{\bf{u}}p(\bf{u})\log p_{G}(\bf{u}) \\
+& = -I(\bf{u}) - H(\bf{x}) - \log(|det(\bf{W})|) - \frac{1}{2}\log((2\pi e)^{d}det(<\bf{u}, \bf{u}^{t}>))\\
+&= -I(\bf{u}) - H(\bf{x}) - \frac{1}{2}\log((2\pi e)^{d}).
+\end{align} In the derivation, the $\bf{u}$ are uncorrelated so the covariance matrix is identity.
+Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\bf{u} = \bf{W}\bf{x}$.
 
-Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution $\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})$ satisfying $p(\mathbf{x}) = \prod_{i}^{d}\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})|\mathbf{W}|$.
+Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution $\hat{p}_{s}(\bf{w}_{i}\bf{x})$ satisfying $p(\bf{x}) = \prod_{i}^{d}\hat{p}_{s}(\bf{w}_{i}\bf{x})|\bf{W}|$.
 The log-likelihood is
 \begin{equation}
-l(\mathbf{W}) = \frac{1}{N}\sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x}^{(j)}))+\log|\mathbf{W}|).
+l(\bf{W}) = \frac{1}{N}\sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\bf{w}_{i}\bf{x}^{(j)}))+\log|\bf{W}|).
 \end{equation}   
-If the approximation $\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})$ are close to the actual pdf, the first term is approximately $-\sum_{i}H(\mathbf{w}_{i}\mathbf{x})$ and is equal to the negative of the mutual information up to an additive constant of the total entropy of $\mathbf{x}$.
+If the approximation $\hat{p}_{s}(\bf{w}_{i}\bf{x})$ are close to the actual pdf, the first term is approximately $-\sum_{i}H(\bf{w}_{i}\bf{x})$ and is equal to the negative of the mutual information up to an additive constant of the total entropy of $\bf{x}$.
 Hence, all these historical information theoretic approaches are equivalent and the key idea is to approximate the probability distribution of the sources correctly.
@@
 where $Y$ is a random variable of zero mean and unit variance, which can be achieved by whitening the data, and $V$ is a Gaussian variable also of zero mean and unit variance.
-In function forms of $G_1$ and $G_2$ should not grow too fast to be robust.
+THe function forms of $G$ should not grow too fast to be robust.
 The following choice of $G$ is proposed by Hyvärinen and Oja~\cite{hyvarinen2000independent}.
 \begin{align}
-G_1(u) &= \frac{1}{a_1} \log (\cosh a_1 u). \\
-G_1^{\prime}(u) &= \tanh(a_1 u).
+G(u) &= \frac{1}{a_1} \log (\cosh a_1 u). \\
+\partial_u G(u) &= \tanh(a_1 u).
 \end{align}
@@
 \begin{align}
-l(\mathbf{W}) &= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\mathbf{w}_{i}^{t}\mathbf{x}^{(j)}))+\log|\mathbf{W}|) \\
-&= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(g'(\mathbf{w}_{i}^{t}\mathbf{x}^{(j)}))+\log|\mathbf{W}|),
-\end{align} where $g'(\mathbf{w}_{i}^{t}\mathbf{x}) = g'(\mathbf{u}_{i}) = g(1-g)$. Thus the gradient of log-likelihood is 
-\begin{equation}
-\frac{\partial l(\mathbf{W})}{\partial \mathbf{W}} = \sum_{j=1}^{N}(\mathbf{q}^{(j)}\mathbf{x}^{t, (j)} + (\mathbf{W}^{t})^{-1}),
-\end{equation} where $\mathbf{q}^{t, (j)} = [1-2g(u_{1}^{(j)}), ... ,1-2g(u_{d}^{(j)}) ]$. 
-In our implementation, we used the stochastic gradient descent method to update the $\mathbf{W}$:
-\begin{equation}
-\mathbf{W} = \mathbf{W} + \alpha (\mathbf{q}^{(j)}\mathbf{x}^{t, (j)} + (\mathbf{W}^{t})^{-1}).
-\end{equation} This means that we update the $\mathbf{W}$ for each sample of the observed data instead of calculating the exact summation for all samples at each step of iteration.
-This iteration schemes is accompanied by randomize the ordering of samples initially for each iteration (making the iteration more stochastic).
+l(\bf{W}) &= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\bf{w}_{i}\bf{x}^{(j)}))+\log|\bf{W}|) \\
+&= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(g'(\bf{w}_{i}\bf{x}^{(j)}))+\log|\bf{W}|)
+\end{align}    
 \section{Applications}
-\subsection{Sklearn example}
-We implement FastICA according to the formalism described in the previous section.
-The algorithm is applied to a following set of signal functions.
+We implement FastICA according to the formalism described in the previous section and apply it to the observed data $X$ in order to find the source matrix $S$ given by
 \begin{align}
-s_1 &= \sin(2 t). \\
-s_2 &= \mathrm{sign} [\sin(3 t)]. \\
-s_3 &= \frac{1}{2\pi} (t~\mathrm{mod}~2\pi).
+S &= \begin{bmatrix}
+s_1(t_1) & s_1(t_2) & \cdots & s_1(t_N) \\
+s_2(t_1) & s_2(t_2) & \cdots & s_2(t_N) \\
+\vdots & \vdots & \ddots & \vdots \\
+s_d(t_1) & s_d(t_2) & \cdots & s_d(t_N)
+\end{bmatrix},
 \end{align}
-The mixing matrix is taken to be
+where $N$ is the number of time steps, $d$ is the number of detectors, and $s_i$'s are independent signal sources.
+
+\subsection{Application 1: sine, square, and sawtooth waves}
+In this first example, we define the independent signal sources according to the following equations and plot them in Fig.~\ref{fig:ex1_sources}(a).
+\begin{align}
+s_1 &= \sin(2 t). \nonumber \\
+s_2 &= 2~\mathrm{sign} [\sin(3 t)]. \nonumber \\
+s_3 &= 4~\left[\frac{1}{\pi} (2 \pi t~\mathrm{mod}~2\pi) - 1\right]. \label{eq:ex1_sources}
+\end{align}
+The mixture signals $X$ (Fig.~\ref{fig:ex1_sources}(b)) that we want to apply FastICA on is constructed by the equation
+\begin{align}
+X &= AS,
+\end{align}
+where $A$ is the mixing matrix, which in this example, is set to be
 \begin{align}
@@
 1 & 1 & 1 \\
-0.5 & 2 & 1 \\
-1.5 & 1 & 2
-\end{bmatrix}.
+\frac{1}{2} & 2 & 1 \\
+\frac{3}{2} & 1 & 2
+\end{bmatrix}. \label{eq:ex1_mixing}
 \end{align}
-The result from our implementation and FastICA are shown in Fig.~\ref{fig:sklearn}.
-\subsection{Sound waves}
-We employed the maximum likelihood algorithm to separate out the independent components of five mixed sound tracks (data obtained from Standford cs229 online course materials).
-The mixed sound waves are shown in the  Fig.~\ref{fig:sound} left panels (red).
-The separated independent sound tracks are shown in the right panels (blue) in Fig.~\ref{fig:sound}.
-The separated sound tracks can be played and listened one by one to verify that they are meaningful.
-\begin{figure}[b]
-\includegraphics[width=0.8\textwidth]{scripts/sound_example.pdf}
- \caption{\label{fig:sound} shows the ICA results of separating the five mixed sound waves in the left panels (red).
-The separated independent components are shown in the right panels (blue).
-When the separated signals are played, one can clearly recognize these meaningful audio signals.}
+Newton's method converages rather quickly within only four cycles for each row of $W$ as shown in Fig.~\ref{fig:ex1_dist}.
+\begin{figure}
+\includegraphics{scripts/ex1_dist.pdf}
+\caption{\label{fig:ex1_dist} The convergence criterion $||w_{n}^T w_{n-1}| - 1|$ as a function of cycle for each row of $W$.}
 \end{figure}
-\subsection{Blinking removal in EEG signal}
-% \begin{figure}
-% \includegraphics{scripts/sklearn.pdf}
-% \caption{\label{fig:sklearn} FastICA results using our code and \lstinline{sklearn.decomposition.FastICA}}
-% \end{figure}
+The sources predicted by the algorithm as implemented in our code and in \lstinline{sklearn.decomposition.FastICA} are shown in Fig.~\ref{fig:ex1_sources}(d, e).
+Both of the results are very similar in terms of the ability to separate out independent sources.
+The main difference is the sign of the sawtooth wave because independence is up to a minus sign as discussed previously in the formalism section.
+From the predicted sources $S_{\mathrm{predicted}}$, we can verify the accuracy of FastICA results by inverting the mixing and whitening transformations and adding back the mean of each mixture to retrieve the input signals.
+\begin{align}
+(X_{\mathrm{retrived}})_{i:} &= [(W K)^{-1} S_{\mathrm{predicted}}]_{i:} + E[X_{i:}], \label{eq:ex1_retrieved}
+\end{align}
+where $K$ is the whitening transformation given by...
+The retrieved signals $X_{\mathrm{retrieved}}$ is shown to be equal to the original input signal $X$ (Fig.~\ref{fig:ex1_sources}(c)), which confirms that we can exactly reproduce the observed data from the output of ICA.
+\begin{figure*}
+\includegraphics{scripts/ex1_sources.pdf}
+\caption{\label{fig:ex1_sources}
+(a) Original sources as defined in Eq.~(\ref{eq:ex1_sources}).
+(b) Mixture signals whose mixing matrix is defined by Eq.~(\ref{eq:ex1_mixing}).
+(c) Retrieved input signals from the predicted sources according to Eq.~(\ref{eq:ex1_retrieved}).
+(d) Separated sources using our implementation of FastICA.
+(e) Separated sources using \lstinline{sklearn.decomposition.FastICA}.
+(f) Separated sources using \lstinline{sklearn.decomposition.PCA}.}
+\end{figure*}
+In some applications, such as separation of independent sound sources, we might not need to retrieve the original signal.
+However, this procedure becomes crucial in other applications, such as blink removals of the electroencephalogram data.
+The discussion of both mentioned applications follows.
 
-%\section{How to check the robustness and convergence}
+In this specific example, the result from \lstinline{sklearn.decomposition.PCA} is also plotted in Fig.~\ref{fig:ex1_sources}(f).
+The figure shows that PCA is not suitable for separating independent sources.
+
+\subsection{Application 2: audio source separation}
+
+%3) example: sound single decomposition
+%4) EEG % (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
+
+\section{How to check the robustness and convergence}
 % Amari metric, ref "ELements of Statistical Learning", p.570
@@
 %1) check if the performance matrix is identity (Te-won Lee, p. 51)
-%\begin{align}
-%W A &= I.
-%\end{align}
+\begin{align}
+W A &= I.
+\end{align}
 
@@
 
-%\section{what if two of the sources are dependent}
+\section{what if two of the sources are dependent}
 %s3 = signal.sawtooth(2*np.pi*time) # saw tooth signal
@@
 \section{Discussions and Conclusions}
-In this review paper of ICA, we have discussed the traditional and basic notions of this technique.
-We have also implemented the algorithms and utilized our codes to solve several different examples.
-We have noticed that the objective function plays an important role in the performance of ICA in various examples we examined.
-In the basic setting, the performance of the algorithm lies in choosing the correct and robust objective function.
-Several objective functions have been proposed to improve the robusteness and computation efficiency.
-However, if we know in advance what the probability distributions look like for the sources (latent variables), it would always be the best to tailor the objective function to include this information.
-Nonetheless, ICA has already shown reliable performance in examples such as sound waves identification and EEG signals separation using some default robust objective functions.
-The current research in this field focuses on cases where the number of detectors smaller than the number of sources, and non-linear mixing problems.
+1) review the traditional development from 1990s
+2) recent advances and development from 2000s  
+3) what happens when there are multiple Gaussian components
 

### 2022-05-03 scripts/ex1.py

@@ -6,15 +6,16 @@ from sklearn.decomposition import FastICA, PCA
 
 import fast_ica
 
-# plt.rc('font', family='serif')
-# plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+plt.rc('text', usetex=True)
 
 def load_data():
     N = 2000
     time = np.linspace(0, 8, N)
-    s1 = 5*np.sin(2*time) # sinusoidal
-    s2 = 20*np.sign(np.sin(3*time)) # square signal
-    s3 = 100*signal.sawtooth(2*np.pi*time) + 50# saw tooth signal
+    s1 = np.sin(2*time) # sinusoidal
+    s2 = 2*np.sign(np.sin(3*time)) # square signal
+    s3 = 4*signal.sawtooth(2*np.pi*time)  # saw tooth signal
+    # s3 = 4*(1/(np.pi)*(2*np.pi*time%(2*np.pi)) - 1)
     s4 = s1 + s2
     s5 = s1 - s2
     s6 = np.cos(2*time)
@@ -28,14 +29,26 @@ def load_data():
     X = A @ S
     return X, S
 
-def plot_sources(sources, titles, output=None):
+def plot_sources(sources, titles, output=None, colwrap=2, figsize=(7, 4)):
 
     n = len(sources)
-    fig, ax = plt.subplots(nrows=n, ncols=1, figsize=(9, 7), sharex=True, sharey=False)
-    for i in range(n):
-        for source in sources[i]:
-            ax[i].plot(source)
-        ax[i].set_title(titles[i], fontsize=8)
+    nrows = int(np.ceil(n/colwrap))
+    ncols = colwrap
+    fig, ax = plt.subplots(nrows=nrows, ncols=ncols, figsize=figsize, sharex=True, sharey='row')
+    for i in range(nrows):
+        for j in range(ncols):
+            idx = i*colwrap + j
+            if idx >= n:
+                break
+            source = sources[idx]
+            title = titles[idx]
+            for component in source:
+                ax[i, j].plot(component)
+            ax[i, j].set_title(title, fontsize=8)
+            if i == nrows - 1:
+                ax[i, j].set_xlabel('Time')
+            if j == 0:
+                ax[i, j].set_ylabel('Value')
 
     fig.tight_layout()
     if output is None:
[…]

### 2022-05-03 scripts/fast_ica.py

@@ -28,6 +28,7 @@ def whiten_svd(X):
     U, D, _ = la.svd(X, full_matrices=False)
     K = U / D
     Z = K.T @ X
+    Z *= X.shape[0]**0.5
     return Z, K.T
 
 def g(x, a1=1):
@@ -42,21 +43,19 @@ def retrieve_X_out(X, W, K, S_out):
 
 def retrieve_whiten(X, W, K, S_out):
     X_whiten = la.inv(W) @ S_out
-    print()
     return X_whiten
 
 def ica(X, cycles=200, tol=1e-5, test=False):
-    X1, K = whiten(X)
-    # X1 = center(X, divide_sd=True)
+    Z, K = whiten(X)
     nrows, ncols = X.shape
-    # K = np.eye(nrows)
     W = np.zeros((nrows, nrows))
     distances = []
     for i in range(nrows):
         w = np.random.random((nrows))
+        w /= ((w**2).sum())**0.5
         dd = []
         for _ in range(cycles):
-            w_new = (X1 * g(w.T @ X1)).mean(axis=1) - dg(w.T @ X1).mean() * w
+            w_new = (Z * g(w.T @ Z)).mean(axis=1) - dg(w.T @ Z).mean() * w
 
             if test:
                 tests.test_gram_schmidt(w_new, W, i)

### 2022-05-03 paper.tex

--- 2022-05-02 paper.tex
@@
-%% ****** Start of file apstemplate.tex ****** %
-%%
-%%
-%%   This file is part of the APS files in the REVTeX 4.2 distribution.
-%%   Version 4.2a of REVTeX, January, 2015
-%%
-%%
-%%   Copyright (c) 2015 The American Physical Society.
-%%
-%%   See the REVTeX 4 README file for restrictions and more information.
-%%
-%
-% This is a template for producing manuscripts for use with REVTEX 4.2
-% Copy this file to another name and then work on that file.
-% That way, you always have this original template file to use.
-%
-% Group addresses by affiliation; use superscriptaddress for long
-% author lists, or if there are many overlapping affiliations.
-% For Phys. Rev. appearance, change preprint to twocolumn.
-% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
-%  Add 'draft' option to mark overfull boxes with black boxes
-%  Add 'showkeys' option to make keywords appear
-%\documentclass[aps,prb,twocolumn,superscriptaddress]{revtex4-1}
 \documentclass[aps,prl,preprint,superscriptaddress]{revtex4-2}
-%\documentclass[aps,prl,reprint,groupedaddress]{revtex4-2}
 
@@
 
-% Use the \preprint command to place your local institutional report
-% number in the upper righthand corner of the title page in preprint mode.
-% Multiple \preprint commands are allowed.
-% Use the 'preprintnumbers' class option to override journal defaults
-% to display numbers if necessary
-%\preprint{}
-
-%Title of paper
 \title{Independent Component Analysis}
 
-% repeat the \author .. \affiliation  etc. as needed
-% \email, \thanks, \homepage, \altaffiliation all apply to the current
-% author. Explanatory text should go in the []'s, actual e-mail
-% address or url should go in the {}'s for \email and \homepage.
-% Please use the appropriate macro foreach each type of information
-
-% \affiliation command applies to all authors since the last
-% \affiliation command. The \affiliation command should follow the
-% other information
-% \affiliation can be followed by \email, \homepage, \thanks as well.
-
-%\email[]{Your e-mail address}
-%\homepage[]{Your web page}
+\author{Kannan Lu}
 \author{Kittithat Krongchon}
-\author{Kannan Lu}
 \affiliation{Department of Physics, University of Illinois at Urbana-Champaign, Urbana, IL 61801, USA}
-
-%Collaboration name if desired (requires use of superscriptaddress
-%option in \documentclass). \noaffiliation is required (may also be
-%used with the \author command).
-%\collaboration can be followed by \email, \homepage, \thanks as well.
-%\collaboration{}
-%\noaffiliation
 
@@
 \begin{abstract}
-Indenpendent component analysis (ICA) is an unsupervised learning technique that is widely used in extracting independent factors in image, sound and medical signals.
+Indenpendent component analysis (ICA) is an unsupervised learning technique that is widely used in extracting independent factors in image, sound, and medical signals.
 In this document, we review the basic notions of the ICA including mathematical formulations and detailed algorithms.
 We also implement different algorithms based on higher-order statistics and information theoretic approaches.
-These algorithms are applied to sound signal disentanglement and EEG blinking signal removal with success.
+These algorithms are applied to sound signal disentanglement and electroencephalogram blinking signal removal with success.
 \end{abstract}
@@
 The motivation of the development of this technique is based on very commom problems.
-For instance, in reality, source signals are often corrupted with 'noise'.
-Or in other words, data are composed by mutiple independent source signals.
+For instance, in reality, source signals are often corrupted with noise.
+In other words, data are composed of mutiple independent source signals.
 To isolate the signals from different sources, ICA can be a good choice.
-The famous example is the 'cocktail party problem', where the sound detectors record the superposition of various sound signals from different sources in the party.
+The famous example is the cocktail party problem, where the sound detectors record the superposition of various sound signals from different sources in the party.
 The goal is to separate out the observed signal into sounds from each different source.
-Since we do not know what the sources are and how these signals are mixed, this type of problem is called 'blind source separation' (BSS) \cite{lee1998independent}.
-ICA is widely used in images, sounds, stock market and medical signals, where latent independent variables are believed to exist.
+Since we do not know what the sources are and how these signals are mixed, this type of problem is called blind source separation (BSS) \cite{lee1998independent}.
+ICA is widely used in images, sounds, stock market, and medical signals, where latent independent variables are believed to exist.
 ICA can also be viewed as an extension of principal component analysis (PCA), where the latter maximizes the second order statistics (covariance matrix of data).
-ICA maximizes higher-order statistics or simply tries to look for independent components not just uncorrelated components.
+ICA maximizes higher-order statistics or simply tries to look for independent components, not just uncorrelated components.
 In Fig.~\ref{fig:ICA_vs_PCA}, we show the ICA components $IC_{i}$ and PCA components $PC_{i}$ learned from the observed data $x_{i}$ with the original sources $s_{i}$ sampled from uniform distribution independently.
-Clearly.
-the ICA learned the independent components correctly but the PCA did not in this case.
+Clearly, the ICA learned the independent components correctly, but the PCA did not in this case.
 
-In this paper, we will review mathematical background of the ICA technique, the detailed algorithm implementation and several realizations in typical examples.
+In this paper, we review mathematical background of the ICA technique, the detailed algorithm implementation and several realizations in typical examples.
 
@@
 \end{figure}
-% Put \label in argument of \section for cross-refe(WHAT DOES THIS LAST SENTENCE MEAN?)rencing
-%\section{\label{}}
-%\subsection{}
-%\subsubsection{}
 
@@
 \subsection{Definition of the problem and notations}
-Before we delve into the detailed mathematical formalisms, we first define the notations that we will use throughout the whole document.
-We will use boldface capital letters $\bf{A}$ for matrices, boldface lowercase letters $\bf{a}$ for vectors and ordinary lower case letters $a$ for scalars.
-We will use subscripts for denoting the components of matrices (i.e.
+Before we delve into the detailed mathematical formalism, we first define the notations that we use throughout the whole document.
+We use boldface capital letters $\bf{A}$ for matrices, boldface lowercase letters $\bf{a}$ for vectors and ordinary lower case letters $a$ for scalars.
+We use subscripts to denote the components of matrices (i.e.
 $A_{ij})$ and vectors (i.e.
@@
 Superscripts are reserved for indexing different samples.
-For instance, $m$ samples of vector $\bf{a}$ can be denoted as $\bf{a}$$^{(k)}$ where $k = 1, 2, ..., m$.
+For instance, $m$ samples of vector $\bf{a}$ can be denoted as $\bf{a}$$^{(k)}$ where $k = 1, 2, \ldots, m$.
 Following these conventions, we denote the original signal to be $\bf{s}$$^{(i)} \in \rm I\!R^{d}$.
@@
 That is there is a linear map between the $\bf{s}$$^{(i)}$ and $\bf{x}$$^{(i)}$, $\bf{x}$$^{(i)}$ $=$ $\bf{A}$$\bf{s}$$^{(i)}$, where $\bf{A} \in \rm I\!R^{d\times d}$ is known as the mixing matrix.
-The whole idea of ICA is to learn the inverse of the mixing matrix $\bf{W} = \bf{A}$$^{-1}$ from the observed data $\bf{x}$$^{(i)}$, known as the un-mixing process.
+The whole idea of ICA is to learn the inverse of the mixing matrix $\bf{W} = \bf{A}$$^{-1}$ from the observed data $\bf{x}$$^{(i)}$, known as the unmixing process.
 The learned independent components are denoted by $\bf{u} = \bf{W}\bf{x}$.
-This can be done in several different ways based on ideas of seperating non-Gaussian signals, and the original signals being independent.
+This can be done in several different ways based on ideas of separating non-Gaussian signals and the original signals being independent.
 We will in this section discuss various mathematical formulations.
-All these various theoretical frameworks are constrained by the following \cite{lee1998independent}:\\
-(1) The number of sensors is larger than number of sources. This ensures that the mixing matrix is full rank; \\
-(2) The sources at each sample (time) are mutually independent; \\
-(3) At most one source is normally distributed. \\
-Without these conditions, the BSS problem is ill-defined.
-For simplicity, we will also restrict the discussion to cases where the number of detectors and sources are the same.
-However, a generalization to condition (1) is feasible and the algorithms discussed in the following sections will still work.
+All these various theoretical frameworks are constrained by the following conditions~\cite{lee1998independent}.
+\begin{enumerate}
+\item The number of sensors is larger than number of sources. This ensures that the mixing matrix is full rank.
+\item The sources at each sample (time) are mutually independent.
+\item At most one source is normally distributed.
+\end{enumerate}
+Without these conditions, the BSS problem is ill-defined. For simplicity, we will also restrict the discussion to cases where the number of detectors and sources are the same.
+However, a generalization to the first condition is feasible, and the algorithms discussed in the following sections will still work.
 
-Based on these definitions, we need to talk about some ambiguities embedded in the symmetry of the problem and justify that if the origianl signals are all Gaussians, they cannot be learned through the un-mixing.
+Based on these definitions, we need to talk about some ambiguities embedded in the symmetry of the problem and justify that if the original signals are all Gaussians, they cannot be learned through the unmixing.
 There are two ambiguities that usually do not affect the practical application, namely the permutation ambiguity and the scaling ambiguity.
 It is trivial to see that the ordering of the original sources \{$s_{j}$\} is ambiguous.
-In the case of scaling, if the original signal $\bf{s}$ is scaled by a non-zero constant to be $c\bf{s}$ where $c \neq 0$.
-Then the mixing matrix $\bf{A}$ can be scaled by $1/c$ and resulting in the same observed data $\bf{x}$$ = \frac{1}{c}\bf{A}$$c\bf{s}$.
-This scaling ambiguity can be furtherly extended with regard to each component of the original signal.
-That is, for a particular component $j$, if we scale the component by $c_{j}$ the corrsponding column of the mixing matrix can be scaled by $1/c_{j}$ to have the observed data unchanged (i.e.
+In the case of scaling, if the original signal $\bf{s}$ is scaled by a non-zero constant to be $c\bf{s}$ where $c \neq 0$, then the mixing matrix $\bf{A}$ can be scaled by $1/c$ and will result in the same observed data $\bf{x}$$ = \frac{1}{c}\bf{A}$$c\bf{s}$.
+This scaling ambiguity can be further extended to each component of the original signal.
+That is, for a particular component $j$, if we scale the component by $c_{j}$, the corresponding column of the mixing matrix can be scaled by $1/c_{j}$ to have the observed data unchanged (i.e.
 $x_{i} = \sum_{j}\frac{1}{c_{j}}A_{ij}c_{j}s_{j}$) \cite{ng_cs229}.
@@
 The other ambiguity that matters for the practical application is that the original sources cannot all be distributed as Gaussians \cite{lee1998independent, ng_cs229}.
-Or to be more precise, in order to separate the independent components, we require that the original signals can only have at most one component to be sampled from a Gaussian distribution, i.e.
+To be more precise, in order to separate the independent components, we require that the original signals have at most only one component to be sampled from a Gaussian distribution, i.e.
 $s_{j}$ $\sim \mathcal{N}(\mu,\,\sigma^{2})$ for at most one $j$.
 Let's consider the case when all the original independent signals are Gaussians.
-That is, $s_{j}$ $\sim \mathcal{N}(\mu_{j},\,\sigma_{j}^{2})$ for $j = 1, 2, ..., d$.
-Then given the mixing matrix $\bf{A}$, we have
+That is, $s_{j}$ $\sim \mathcal{N}(\mu_{j},\,\sigma_{j}^{2})$ for $j = 1, 2, \ldots, d$.
+Then, given the mixing matrix $\bf{A}$, we have
 \begin{equation}
-\mathbb{E}_{\mathbf{s}}[\mathbf{x}] = \mathbf{A}\mathbb{E}[\mathbf{s}] = \mathbf{A}\bm{\mu}
+\mathbb{E}[\mathbf{x}] = \mathbf{A}\mathbb{E}[\mathbf{s}] = \mathbf{A}\bm{\mu}
 \end{equation}
@@
 \begin{equation}
-Cov[\mathbf{x}] = \mathbb{E}_{\mathbf{s}}[\mathbf{As}\mathbf{s}^t\mathbf{A}^{t}]-\mathbb{E}_{\mathbf{s}}[\mathbf{As}]\mathbb{E}[(\mathbf{As})^{t}] =  \mathbf{A}\bm{\Sigma}\mathbf{A}^{t}-\mathbf{A}\mathbf{M}\mathbf{A}^{t},
+\mathrm{Cov}[\mathbf{x}] = \mathbb{E}[\mathbf{As}\mathbf{s}^t\mathbf{A}^{t}]-\mathbb{E}[\mathbf{As}]\mathbb{E}[(\mathbf{As})^{t}] =  \mathbf{A}\bm{\Sigma}\mathbf{A}^{t}-\mathbf{A}\mathbf{M}\mathbf{A}^{t},
 \end{equation} where $\mathbf{M}$ and $\bm{\Sigma}$ are diagonal matrices of $\rm I\!R^{d\times d}$ with diagonal elements to be \{$\mu_{j}^{2}$\} and \{$\sigma_{j}^{2}$\}, respectively.
-Due to the scale ambiguity, this is the same as considering normally distributed signals with unity variance.
-Since each column $j$ of the mixing matrix $\mathbf{A}$ can be scaled by $\sigma_{j}$, $\tilde{\mathbf{A}}_{:j} = \sigma_{j}\mathbf{A}_{:j}$ and correspondingly the signal needs to be scaled by $1/\sigma_{j}$, $\tilde{s_{j}} = \frac{s_{j}-\mu_{j}}{\sigma_{j}}$.
+Due to the scaling ambiguity, this is the same as considering normally distributed signals with unity variance.
+Since each column $j$ of the mixing matrix $\mathbf{A}$ can be scaled by $\sigma_{j}$, $\tilde{\mathbf{A}}_{:j} = \sigma_{j}\mathbf{A}_{:j}$, and correspondingly, the signal needs to be scaled by $1/\sigma_{j}$, $\tilde{s_{j}} = \frac{s_{j}-\mu_{j}}{\sigma_{j}}$.
 Then,
 \begin{equation}
-Cov[\mathbf{x}] = \tilde{\mathbf{A}}\mathbf{I}_{d\times d}\tilde{\mathbf{A}}^{t}.
+\mathrm{Cov}[\mathbf{x}] = \tilde{\mathbf{A}}\mathbf{I}_{d\times d}\tilde{\mathbf{A}}^{t}.
 \end{equation}
-Now, if we consider a rotation $\mathbf{R} $$\in O(d)$ acting on the scaled source signals $\mathbf{\tilde{s}}$.
-The mixing matrix then changes to $\tilde{\mathbf{A}}\mathbf{R}$.
-Unpon this rotation, we would observe $\mathbf{x'}$ as $\mathbf{x'} = \tilde{\mathbf{A}}\mathbf{R}\tilde{\mathbf{s}}$.
-$\mathbf{x'}$ is again normally distributed and has covariance matrix,
+Now, if we consider a rotation $\mathbf{R} \in O(d)$ acting on the scaled source signals $\mathbf{\tilde{s}}$, the mixing matrix then changes to $\tilde{\mathbf{A}}\mathbf{R}$.
+Upon this rotation, we would observe $\mathbf{x'}$ as $\mathbf{x'} = \tilde{\mathbf{A}}\mathbf{R}\tilde{\mathbf{s}}$, and $\mathbf{x'}$ is again normally distributed and has covariance matrix,
 \begin{equation}
-Cov[\mathbf{x'}] = \tilde{\mathbf{A}}\mathbf{R}\mathbf{I}_{d\times d}\mathbf{R}^{t}\tilde{\mathbf{A}}^{t} = \tilde{\mathbf{A}}\mathbf{I}_{d\times d}\tilde{\mathbf{A}}^{t}. 
+\mathrm{Cov}[\mathbf{x'}] = \tilde{\mathbf{A}}\mathbf{R}\mathbf{I}_{d\times d}\mathbf{R}^{t}\tilde{\mathbf{A}}^{t} = \tilde{\mathbf{A}}\mathbf{I}_{d\times d}\tilde{\mathbf{A}}^{t}.
 \end{equation} This simply means that whether the sources are rotated or not the observed data will be distributed as $\mathcal{N}(0, \tilde{\mathbf{A}}\tilde{\mathbf{A}}^{t})$.
-Thus, because of the rotational symmetry of the multivariate Gaussian distribution, we cannot seperate and obtain the original source signals.
-These derivations also indicate that given the observed data $\bf{x}$, we can whiten (apply PCA rotation and normalize each principal component) the data and it will not affect the seperated signals (up to scalings and mean translations).
+Thus, because of the rotational symmetry of the multivariate Gaussian distribution, we cannot separate and obtain the original source signals.
+These derivations also indicate that given the observed data $\bf{x}$, we can whiten (apply PCA rotation and normalize each principal component) the data, and it does not affect the separated signals (up to scalings and mean translations).
 We will assume the observed data $\bf{x}$ are whitened if not explicitly stated in later discussions.
-That is $\tilde{\mathbf{x}} = \mathbf{\Sigma}^{-1/2}(\mathbf{x}-\mathbf{\mu})$ (whitening).
-Then, to recover the source signal, we need to find $\mathbf{W}$ with orthonormal rows \{$\mathbf{w}_{i}^{t}$\} such that $\mathbf{s} = \mathbf{W}\mathbf{\Sigma}^{-1/2}\mathbf{x}$.
+That is $\tilde{\mathbf{x}} = \mathbf{K}(\mathbf{x}-\bm{\mu})$, where $\mathbf{K}$ is the whitening transformation.
+Then, to recover the source signals, we need to find $\mathbf{W}$ with orthonormal rows \{$\mathbf{w}_{i}^{t}$\} such that $\mathbf{s} = \mathbf{W}\mathbf{K}\mathbf{x}$.
 
@@
 
-The non-Gaussianity thus directly provide us with the methodologies used in finding the independent component.
+The non-Gaussianity directly provides us with the methodologies used in finding the independent component.
 Notice that the PCA only finds the uncorrelated components but not necessarily independent.
-The ICA tries to find independent components where the joint distibutions can be factorized into marginal distributions by investigating into higher-order moments.
+The ICA tries to find independent components where the joint distributions can be factorized into marginal distributions by investigating higher-order moments.
 There are multiple moment-based objective functions that have been used along the development of ICA.
-We will define some of the notations and discuss one of the moment-based objective functions in detail.
+We define some of the notations and discuss one of the moment-based objective functions in detail.
 
-As we discussed previously, we will whiten the observed signals and we will denote the whitened observed data as $\bf{x}$.
-The kurtosis and excess kurtosis of whitended variable $x$ are defined as $$ \beta(x) := \mathbb{E}[x^{4}], \kappa(x) := \beta(x)-3 $$.
-The standard normal randon varible has excess kurtosis to be 0.
-When excess kurtosis $\kappa(x) < 0$, $x$ is said to be sub-Gaussian (flat around center, e.g.
+As we discussed previously, we whiten the observed signals and we denote the whitened observed data as $\bf{x}$. The kurtosis and excess kurtosis of whitened variable $x$ are defined as
+\begin{align}
+\beta(x) := \mathbb{E}[x^{4}],~\kappa(x) := \beta(x)-3.
+\end{align}
+The standard normal random variable has excess kurtosis to be 0.
+When excess kurtosis $\kappa(x) < 0$, $x$ is said to be sub-Gaussian (flat around the center, e.g.
 uniform distribution).
-When $\kappa(x) > 0$, $x$ is super-Gaussian (fat tail, sharp peak around center, e.g.
+When $\kappa(x) > 0$, $x$ is super-Gaussian (heavy tail and sharp peak around the center, e.g.
 Laplace distribution).
-Motivated by this the objective function to maximize can be $|\kappa(x)|$ and $\kappa^{2}(x)$ etc.
-This is furtherly backed up by the following inequalities \cite{miettinen2015fourth}:
+The objective function to maximize can be $|\kappa(x)|$ and $\kappa^{2}(x)$ etc. This is further backed up by the following inequalities \cite{miettinen2015fourth}.
+\begin{align}
+\forall i \in \{1, \ldots, d\}, ~|\kappa(\mathbf{w}_{i}^{t}\mathbf{x})| &\leq \mathrm{max}\{|\kappa(z_1)|, \ldots ,|\kappa(z_d)|\}. \\
+|\kappa(\mathbf{w}_{1}^{t}\mathbf{x})|+ \cdots + |\kappa(\mathbf{w}_{d}^{t}\mathbf{x})| &\leq |\kappa(z_1)|+ \cdots +|\kappa(z_d)|,
+\end{align}
+where $z_{i}$'s are the standardized independent components.
+
+Thus, the problem is equivalent to finding an orthogonal transformation $\mathbf{W}$ with orthonormal rows \{$\mathbf{w}_{i}^{t}$\} such that the objective function can be maximized.
+In the case of the objective function $L(\mathbf{w}) = |\kappa(\mathbf{w}^{t}\mathbf{x})|$, this amounts to iterating the $\mathbf{w}$ with gradient~\cite{hyvarinen2000independent},
 \begin{equation}
-|\mathbb{E}[(\mathbf{w}^{t}\mathbf{x})^{4}]-3| \leq max\{|\mathbb{E}[x_{1}^{4}]-3|,...,|\mathbb{E}[x_{d}^{4}]-3|\}
-\end{equation}
-\begin{equation}
-|\mathbb{E}[(\mathbf{w}_{1}^{t}\mathbf{x})^{4}-3]|+...+|\mathbb{E}[(\mathbf{w}_{d}^{t}\mathbf{x})^{4}-3]| \leq |\mathbb{E}[x_{1}^{4}]-3|+...+|\mathbb{E}[x_{d}^{4}]-3|
-\end{equation}
-Thus the problem is equivalent to finding an orthogonal transformation $\mathbf{W}$ with orthonormal rows \{$\mathbf{w}_{i}^{t}$\} such that the objective function can be maximized.
-In the case of the objective function $L(\mathbf{w}) = |\kappa(\mathbf{w}^{t}\mathbf{x})|$, this amounts to iterating the $\mathbf{w}$ with gradient \citep{hyvarinen2000independent},
-\begin{equation}
-\frac{\partial |\kappa(\mathbf{w}^{t}\mathbf{x})|}{\partial \mathbf{w}} = 4 sign(\kappa(\mathbf{w}^{t}\mathbf{x}))(\mathbb{E}[\mathbf{x}(\mathbf{w}^{t}\mathbf{x}))^{3}])-3\mathbf{w}||\mathbf{w}||^{2}).
+\frac{\partial |\kappa(\mathbf{w}^{t}\mathbf{x})|}{\partial \mathbf{w}} = 4~\mathrm{sign} [\kappa(\mathbf{w}^{t}\mathbf{x})]\{\mathbb{E}[\mathbf{x}(\mathbf{w}^{t}\mathbf{x})^{3}]-3\mathbf{w}||\mathbf{w}||^{2}\}.
 \end{equation}
 This objective function captures both sub-Gaussian and super-Gaussian distributions for the source signals but the kurtosis is sensitive to the outliers of the observed data.
-Hence, some adaptive objective functions are used in practice more extensively based on negentropy ideas (to be covered later) such as \begin{align}
-G_{1}(u) &= \frac{1}{a_{1}}\log(\cosh a_{1}u) \\
-G_{2}(u) &= -\exp(-u^{2}/2), 
-\end{align} where $u = \mathbf{w}^{t}\mathbf{x}$ \citep{hyvarinen2000independent}. We will use $G_{1}$ for our implementation of one of the algorithms in later sections.
+Hence, some adaptive objective functions are used in practice more extensively based on negentropy~\cite{hyvarinen2000independent}.
 
@@
 
-Apart from the moment based approaches, there are several information theoetic formalisms.
-Here in this section, we will review all different information theoretic frameworks and show that essentially they are equivalent.
-Then we will talk about the learning rules for these theoretical formulations in the subsequant setion.
-
-The centralized quantity in all these formalisms is the mutual information, which is defined as the KL divergence of the multivariate distribution of the observed data $\bf{x}$ and product of all its marginal distributions, i.e.
+Apart from the moment-based approaches, there are several information theoretic methods.
+In this section, we will review all different information theoretic frameworks and show that essentially they are equivalent.
+Then we will talk about the learning rules for these theoretical formulations in the subsequent section.
+The centralized quantity in this formalism is the mutual information, which is defined as the Kullback--Leibler (KL) divergence of the multivariate distribution of the observed data $\bf{x}$ and product of all its marginal distributions, i.e.
 \begin{equation}
 I(\mathbf{x}) = \sum_{\mathbf{x}}p(\mathbf{x})\log\frac{p(\mathbf{x})}{\prod_{i}p_{i}(x_{i})}.
-\end{equation} $I(\bf{x})$ is non-negative and is 0 if and only if the $\bf{x}$ are independent.
-Several different information theoretic approaches have been formulated in history and can be unified in the concept of minimizing the mutual information of the seperated components $\bf{u} = \bf{W}\bf{x}$.
+\end{equation} $I(\bf{x})$ is non-negative and is zero if and only if the $\bf{x}$ are independent.
+Several different information theoretic approaches have been formulated in history and can be unified in the concept of minimizing the mutual information of the separated components $\bf{u} = \bf{W}\bf{x}$.
 
-One of them is a key finding in neural networks, that is, maximizing information between inputs $\mathbf{x}$ and outputs $\mathbf{y}$ implies that the output distributions are factorized, thus minimizes the mutual information in the outputs $\mathbf{y}$.
-Maximizing the information between inputs and outputs is to maximize the output joint entropy $H(\mathbf{y}) = \sum_{i}H(y_{i}) - I(\mathbf{y})$, where $I(\mathbf{y}) = \sum_{\mathbf{y}}p(\mathbf{y})\log\frac{p(\mathbf{y})}{\prod_{i}p_{i}(y_{i})}$ is the mutual information in the outputs $\mathbf{y}$.
+One formulation in neural networks is to maximize information between inputs $\mathbf{x}$ and outputs $\mathbf{y}$, which implies that the output distributions are factorized and thus minimize the mutual information in the outputs $\mathbf{y}$.
+Maximizing the information between inputs and outputs is to maximize the output joint entropy $H(\mathbf{y}) = \sum_{i}H(y_{i}) - I(\mathbf{y})$, where $I(\mathbf{y})$ is the mutual information in the outputs $\mathbf{y}$.
 The mutual information is non-negative and is zero if and only if $\mathbf{y}$ are marginalized.
-Recall that in neural network, the output is given by the nonlinearity $y_{i} = g_{i}(\mathbf{w}_{i}^{t}\mathbf{x}) := g_{i}(u_{i})$, thus $p(y_{i}) = \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|}$.
+Recall that in neural networks, the output is given by the nonlinearity $y_{i} = g_{i}(\mathbf{w}_{i}^{t}\mathbf{x}) := g_{i}(u_{i})$. Thus,
+\begin{align}
+p(y_{i}) = \left|\frac{\partial g_{i}}{\partial u_{i}} \right|^{-1} p(u_{i}).
+\end{align}
 The maximum is therefore obtained by considering the gradient,
 \begin{equation}
-\frac{\partial H(\mathbf{y})}{\partial \mathbf{W}} = \frac{\partial (-I(\mathbf{y}))}{\partial \mathbf{W}} -\frac{\partial}{\partial \mathbf{W}}\sum_{i} \mathbb{E}[\log \frac{p(u_{i})}{|\frac{\partial g_{i}}{\partial u_{i}}|} ].
+\frac{\partial H(\mathbf{y})}{\partial \mathbf{W}} = \frac{\partial (-I(\mathbf{y}))}{\partial \mathbf{W}}
+-\frac{\partial}{\partial \mathbf{W}}\sum_{i} \mathbb{E} \left[\log  \left\{\left|\frac{\partial g_{i}}{\partial u_{i}}\right|^{-1} p(u_{i}) \right\} \right].
 \end{equation}
-This implies that the nonlinear function $g_{i}$ in the neural network $y_{i} = g_{i}(\mathbf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$ in order to kill the second term.
+This implies that the nonlinear function $g_{i}$ in the neural networks, $y_{i} = g_{i}(\mathbf{w}_{i}^{t}\bf{x})$ is a cdf of the source distribution $s_{i}$ in order to kill the second term.
 Together with the constraint that the outputs are marginalized, this gradient is zero.
@@
 &= \sum_{\mathbf{u}}p(\mathbf{u}) \log\frac{\prod_{i}p(u_{i})}{p(\mathbf{u})} + \sum_{\mathbf{u}}p(\mathbf{u})\log\frac{p(\mathbf{u})}{p_{G}(\mathbf{u})} \\
-&= D(\prod_{i}p(u_{i})||p(\mathbf{u})) + J(\mathbf{u}) \\
+&= D \left (\prod_{i}p(u_{i})||p(\mathbf{u}) \right) + J(\mathbf{u}) \\
 &= -I(\mathbf{u}) + J(\mathbf{u}) \\
 & = -I(\mathbf{u}) - H(\mathbf{u}) - \sum_{\mathbf{u}}p(\mathbf{u})\log p_{G}(\mathbf{u}) \\
-& = -I(\mathbf{u}) - H(\mathbf{x}) - \log(|det(\mathbf{W})|) - \frac{1}{2}\log((2\pi e)^{d}det(<\mathbf{u}, \mathbf{u}^{t}>))\\
+& = -I(\mathbf{u}) - H(\mathbf{x}) - \log(|\mathrm{det}(\mathbf{W})|) - \frac{1}{2}\log((2\pi e)^{d} \mathrm{det}(\langle \mathbf{u}, \mathbf{u}^{t} \rangle))\\
 &= -I(\mathbf{u}) - H(\mathbf{x}) - \frac{1}{2}\log((2\pi e)^{d}).
 \end{align} In the derivation, the $\mathbf{u}$ are uncorrelated so the covariance matrix is identity.
-Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\mathbf{u} = \mathbf{W}\mathbf{x}$ \cite{lee1998independent}.
+Therefore, maximizing the negentropy is equivalent to minimizing mutual information in $\mathbf{u} = \mathbf{W}\mathbf{x}$~\cite{lee1998independent}.
 In practice, the negentropy is difficult to evaluate so approximation schemes have been developed and have been discussed in the previous section.
 
-Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution $\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})$ satisfying $p(\mathbf{x}) = \prod_{i=1}^{d}\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})|\mathbf{W}|$ \cite{ng_cs229}.
+Lastly, for the maximum likelihood estimation approach, we want to maximize the log likelihood over all samples observed upon choosing a parametrized distribution $\hat{p}_{s}(\mathbf{w}_{i}^t\mathbf{x})$ satisfying $p(\mathbf{x}) = \prod_{i=1}^{d}\hat{p}_{s}(\mathbf{w}_{i}^t\mathbf{x})|\mathbf{W}|$ \cite{ng_cs229}.
 The log-likelihood is
 \begin{equation}
-l(\mathbf{W}) = \frac{1}{N}\sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x}^{(j)}))+\log|\mathbf{W}|).
+l(\mathbf{W}) = \frac{1}{N}\sum_{j=1}^{N} \left(\sum_{i=1}^{d}\log(\hat{p}_{s}(\mathbf{w}_{i}^t\mathbf{x}^{(j)}))+\log|\mathbf{W}|\right).
 \end{equation}
-If the approximation $\hat{p}_{s}(\mathbf{w}_{i}\mathbf{x})$ are close to the actual pdf, the first term is approximately $-\sum_{i}H(\mathbf{w}_{i}\mathbf{x})$ and is equal to the negative of the mutual information up to an additive constant of the total entropy of $\mathbf{x}$.
+If the approximation $\hat{p}_{s}(\mathbf{w}_{i}^t\mathbf{x})$ are close to the actual pdf, the first term is approximately $-\sum_{i}H(\mathbf{w}_{i}^t\mathbf{x})$ and is equal to the negative of the mutual information up to an additive constant of the total entropy of $\mathbf{x}$.
 Hence, all these historical information theoretic approaches are equivalent and the key idea is to approximate the probability distribution of the sources correctly.
-Historically, there has been several proposed parametric distributions for approximating super-Gaussian or sub-Gaussian distributions.
-We will proceed to discuss the algorithm associated with the information based approach in the next section.
+Historically, there has been several proposed parametric distributions for approximating super-Gaussian or sub-Gaussian distributions~\cite{lee1998independent}.
+In the following section, we will discuss the algorithms associated with the moment-based and information-based approaches.
 
 \section{Algorithms}
-In this section, we will discuss some of the common algorithms based on the previous mathematical formalisms.
-We will categorize them into moment-based approaches and information-based approaches.
+In this section, we discuss some of the common algorithms based on the previous mathematical formalism.
+We categorize them into moment-based approaches and information-based approaches.
 In either case, the problem is equivalent to an optimization problem.
@@
 ordinary gradient descent, Newton's method or fixed point algorithm.
-We will simply specify what we implemented for various examples that we will discuss later.
+We simply specify what we implemented for various examples that we will discuss later.
 \subsection{Moment-based approach}
@@
 \begin{align}
-J(Y) \propto \{\mathbb{E}[G(Y)] - \mathbb{E}[G(V)]\}^2,
+J(u) \propto \{\mathbb{E}[G(u)] - \mathbb{E}[G(\nu)]\}^2,
 \end{align}
-where $Y$ is a random variable of zero mean and unit variance, which can be achieved by whitening the data, and $V$ is a Gaussian variable also of zero mean and unit variance.
-In function forms of $G_1$ and $G_2$ should not grow too fast to be robust.
+where $u$ is a random variable of zero mean and unit variance, which can be achieved by whitening the data, and $\nu$ is a Gaussian variable also of zero mean and unit variance.
+The function forms of $G$ should not grow too fast to be robust.
 The following choice of $G$ is proposed by Hyvärinen and Oja~\cite{hyvarinen2000independent}.
 \begin{align}
-G_1(u) &= \frac{1}{a_1} \log (\cosh a_1 u). \\
-G_1^{\prime}(u) &= \tanh(a_1 u).
+G(u) &= \frac{1}{a_1} \log (\cosh a_1 u). \\
+g(u) &= \partial_u G(u) = \tanh(a_1 u).
 \end{align}
-We want to find the extrema of $\mathbb{E}[G(Y)]$ to maximize $J(Y)$ under the constraints
+We want to find the extrema of $\mathbb{E}[G(u)]$ to maximize $J(u)$ under the constraints
 \begin{align}
-\mathbb{E}(\mathbf{w}^{t}\mathbf{x})^2] = \| \mathbf{w}\|^2 = 1.
+\mathbb{E}[(\mathbf{w}^{t}\mathbf{x})^2] = \| \mathbf{w}\|^2 = 1.
 \end{align}
@@
 \begin{align}
-\mathbb{E}[\mathbf{x}g(\mathbf{w}^{t}\mathbf{x})^2)] - \beta\mathbf{w} &= \mathbf{0}.
+\mathbb{E}[\mathbf{x}g(\mathbf{w}^{t}\mathbf{x})] - \beta\mathbf{w} &= \mathbf{0}.
 \end{align}
@@
 There are multiple choices here tailored to different problems.
-A default choice is to assume the cdf of the source signal can be approximated by a sigmoid function $g(s_{i}) = \frac{1}{1+e^{-s_{s}}}$, thus the pdf is $p(s_{i}) = g'(s_{i})$.
+A default choice is to assume the cdf of the source signal can be approximated by a sigmoid function
+\begin{align}
+g(s_{i}) = \frac{1}{1+e^{-s_{i}}}.
+\end{align}
+Thus, the pdf is $p(s_{i}) = g'(s_{i})$.
 Notice that this pdf is super-Gaussian.
@@
 \begin{align}
-l(\mathbf{W}) &= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(\hat{p}_{s}(\mathbf{w}_{i}^{t}\mathbf{x}^{(j)}))+\log|\mathbf{W}|) \\
-&= \sum_{j=1}^{N}(\sum_{i=1}^{d}\log(g'(\mathbf{w}_{i}^{t}\mathbf{x}^{(j)}))+\log|\mathbf{W}|),
-\end{align} where $g'(\mathbf{w}_{i}^{t}\mathbf{x}) = g'(\mathbf{u}_{i}) = g(1-g)$. Thus the gradient of log-likelihood is 
+l(\mathbf{W}) &= \sum_{j=1}^{N} \left(\sum_{i=1}^{d}\log(\hat{p}_{s}(\mathbf{w}_{i}^{t}\mathbf{x}^{(j)}))+\log|\mathbf{W}|\right) \\
+&= \sum_{j=1}^{N} \left(\sum_{i=1}^{d}\log(g'(\mathbf{w}_{i}^{t}\mathbf{x}^{(j)}))+\log|\mathbf{W}| \right),
+\end{align}
+where $g'(\mathbf{w}_{i}^{t}\mathbf{x}) = g'(\mathbf{u}_{i}) = g(1-g)$.
+Thus, the gradient of log-likelihood is
 \begin{equation}
-\frac{\partial l(\mathbf{W})}{\partial \mathbf{W}} = \sum_{j=1}^{N}(\mathbf{q}^{(j)}\mathbf{x}^{t, (j)} + (\mathbf{W}^{t})^{-1}),
-\end{equation} where $\mathbf{q}^{t, (j)} = [1-2g(u_{1}^{(j)}), ... ,1-2g(u_{d}^{(j)}) ]$. 
-In our implementation, we used the stochastic gradient descent method to update the $\mathbf{W}$:
+\frac{\partial l(\mathbf{W})}{\partial \mathbf{W}} = \sum_{j=1}^{N} [\mathbf{q}^{(j)}\mathbf{x}^{t, (j)} + (\mathbf{W}^{t})^{-1}],
+\end{equation}
+where $\mathbf{q}^{t, (j)} = [1-2g(u_{1}^{(j)}), \ldots ,1-2g(u_{d}^{(j)}) ]$.
+In our implementation, we used the stochastic gradient descent method to update $\mathbf{W}$:
 \begin{equation}
-\mathbf{W} = \mathbf{W} + \alpha (\mathbf{q}^{(j)}\mathbf{x}^{t, (j)} + (\mathbf{W}^{t})^{-1}).
-\end{equation} This means that we update the $\mathbf{W}$ for each sample of the observed data instead of calculating the exact summation for all samples at each step of iteration.
-This iteration schemes is accompanied by randomize the ordering of samples initially for each iteration (making the iteration more stochastic).
+\mathbf{W} = \mathbf{W} + \alpha [\mathbf{q}^{(j)}\mathbf{x}^{t, (j)} + (\mathbf{W}^{t})^{-1}].
+\end{equation}
+This means that we update $\mathbf{W}$ for each sample of the observed data instead of calculating the exact summation for all samples at each step of iteration.
+This iteration scheme is accompanied by randomizing the order of samples for each iteration (making the iteration more stochastic).
+
 \section{Applications}
-\subsection{Sklearn example}
-We implement the moment-based approach according to the formalism described in the previous section.
-The algorithm is applied to a following set of signal functions (example from sklearn module):
+We implement the moment-based approach according to the formalism described in the previous section and apply it to the observed data $\mathbf{x}^{(i)}$ in order to find the source matrix $\mathbf{s}^{(i)}$ given by
 \begin{align}
-s_1 &= \sin(2 t) \\
-s_2 &= \mathrm{sign} [\sin(3 t)] \\
-s_3 &= \frac{1}{2\pi} (t~\mathrm{mod}~2\pi).
+\mathbf{S} &= \begin{bmatrix}
+s_1(t_1) & s_1(t_2) & \cdots & s_1(t_N) \\
+s_2(t_1) & s_2(t_2) & \cdots & s_2(t_N) \\
+\vdots & \vdots & \ddots & \vdots \\
+s_d(t_1) & s_d(t_2) & \cdots & s_d(t_N)
+\end{bmatrix},
 \end{align}
-The mixing matrix is taken to be
+where $N$ is the number of time steps, $d$ is the number of detectors, and $s_i$'s are independent signal sources.
+
+\subsection{Application 1: sine, square, and sawtooth waves}
+In this first example, we define the independent signal sources according to the following equations and plot them in Fig.~\ref{fig:ex1_sources}(a).
+\begin{align}
+s_1 &= \sin(2 t). \nonumber \\
+s_2 &= 2~\mathrm{sign} [\sin(3 t)]. \nonumber \\
+s_3 &= 4~\left[\frac{1}{\pi} (2 \pi t~\mathrm{mod}~2\pi) - 1\right]. \label{eq:ex1_sources}
+\end{align}
+The mixture signals $\mathbf{x}^{(i)}$ (Fig.~\ref{fig:ex1_sources}(b)) that we want to apply the moment-based algorithm on is constructed by the equation
+\begin{align}
+\mathbf{x} &= \mathbf{A} \mathbf{s},
+\end{align}
+where $\mathbf{A}$ is the mixing matrix, which in this example, is set to be
 \begin{align}
@@
 1 & 1 & 1 \\
-0.5 & 2 & 1 \\
-1.5 & 1 & 2
-\end{bmatrix}.
+\frac{1}{2} & 2 & 1 \\
+\frac{3}{2} & 1 & 2
+\end{bmatrix}. \label{eq:ex1_mixing}
 \end{align}
-The result from our implementation and sklearn module are shown in Fig.~\ref{fig:sklearn}.
-\subsection{Sound waves}
-We employed the maximum likelihood algorithm to separate out the independent components of five mixed sound tracks (data obtained from Standford cs229 online course materials).
-The mixed sound waves are shown in the  Fig.~\ref{fig:sound} left panels (red).
+Newton's method converages rather quickly within only four cycles for each row of $\mathbf{W}$ as shown in Fig.~\ref{fig:ex1_dist}.
+\begin{figure}
+\includegraphics{scripts/ex1_dist.pdf}
+\caption{\label{fig:ex1_dist} The convergence criterion $||\mathbf{w}_{n}^t \mathbf{w}_{n-1}| - 1|$ as a function of cycle for each row of $W$.}
+\end{figure}
+The sources predicted by the algorithm as implemented in our code and in \lstinline{sklearn.decomposition.FastICA} are shown in Fig.~\ref{fig:ex1_sources}(d, e).
+Both of the results are very similar in terms of the ability to separate out independent sources.
+The main difference is the sign of the sawtooth wave because independence is up to a minus sign as discussed previously in the formalism section.
+From the predicted sources $\mathbf{s}_{\mathrm{predicted}}$, we can verify the accuracy of FastICA results by inverting the mixing and whitening transformations and adding back the mean of each mixture to retrieve the input signals.
+\begin{align}
+\mathbf{x}_{\mathrm{retrived}} &= (\mathbf{W} \mathbf{K})^{-1} \mathbf{s}_{\mathrm{predicted}} + \mathbb{E}[\mathbf{x}], \label{eq:ex1_retrieved}
+\end{align}
+where $\mathbf{K}$ is the whitening transformation.
+The retrieved signals $\mathbf{x}_{\mathrm{retrieved}}$ are shown to be equal to the original input signals $\mathbf{x}$ (Fig.~\ref{fig:ex1_sources}(c)), which confirms that we can exactly reproduce the observed data from the output of ICA.
+\begin{figure*}
+\includegraphics{scripts/ex1_sources.pdf}
+\caption{\label{fig:ex1_sources}
+(a) Original sources as defined in Eq.~(\ref{eq:ex1_sources}).
+(b) Mixture signals whose mixing matrix is defined by Eq.~(\ref{eq:ex1_mixing}).
+(c) Retrieved input signals from the predicted sources according to Eq.~(\ref{eq:ex1_retrieved}).
+(d) Separated sources using our implementation of FastICA.
+(e) Separated sources using \lstinline{sklearn.decomposition.FastICA}.
+(f) Separated sources using \lstinline{sklearn.decomposition.PCA}.
+}
+\end{figure*}
+In some applications, such as separation of independent sound sources, we might not need to retrieve the original mixtures.
+However, this procedure becomes crucial in other applications, such as blink removals of the electroencephalogram (EEG) data. The discussion of both mentioned applications follows.
+
+In this specific example, the result from \lstinline{sklearn.decomposition.PCA} is also plotted in Fig.~\ref{fig:ex1_sources}(f).
+The figure shows that PCA is not suitable for separating independent sources.
+
+\subsection{Application 2: audio source separation}
+We employ the maximum likelihood algorithm to separate out the independent components of five mixed sound tracks~\cite{ng_cs229}.
+The mixed sound waves are shown in Fig.~\ref{fig:sound} left panels (red).
 The separated independent sound tracks are shown in the right panels (blue) in Fig.~\ref{fig:sound}.
@@
 \end{figure}
-\subsection{Blinking removal in EEG signal}
-% \begin{figure}
-% \includegraphics{scripts/sklearn.pdf}
-% \caption{\label{fig:sklearn} FastICA results using our code and \lstinline{sklearn.decomposition.FastICA}}
-% \end{figure}
+
+\subsection{Application 3: blinking removal in EEG signals}
+
+In Fig.~\ref{fig:ex3_ica}, we decompose the EEG signals~\cite{talebi2021eeg} from 64 detectors into 64 independent components.
+We are able to identity the 48th component to be associated with blinking signals because of the four visible peaks.
+The 48th component shows a different time structure in Fig....
+We reconstruct the signals $\mathbf{x}_{\mathrm{retrived}}$ using Eq.~\ref{eq:ex1_retrieved} with the 48th component set to zero.
+The final result is shown in Fig.~\ref{fig:ex3_noblinks} in orange.
+\begin{figure*}
+\includegraphics{scripts/ex3_ica.pdf}
+\caption{\label{fig:ex3_ica} All 64 independent components separated by the moment-based algorithm.}
+\end{figure*}
+\begin{figure*}
+\includegraphics{scripts/ex3_noblinks.pdf}
+\caption{\label{fig:ex3_noblinks} Original signals (blue) and reconstructed signals with the blinking component removed (orange).}
+\end{figure*}
+
+% (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
 
@@
 
-\section{Discussions and Conclusions}
-In this review paper of ICA, we have discussed the traditional and basic notions of this technique.
-We have also implemented the algorithms and utilized our codes to solve several different examples.
-We have noticed that the objective function plays an important role in the performance of ICA in various examples we examined.
+\section{Conclusions}
+In this review paper of ICA, we discuss the traditional and basic notions of this technique. We also implement the algorithms and utilize our codes in three different applications.
+We notice that the objective function plays an important role in the performance of ICA in various examples we examined.
 In the basic setting, the performance of the algorithm lies in choosing the correct and robust objective function.
-Several objective functions have been proposed to improve the robusteness and computation efficiency.
-However, if we know in advance what the probability distributions look like for the sources (latent variables), it would always be the best to tailor the objective function to include this information.
-Nonetheless, ICA has already shown reliable performance in examples such as sound waves identification and EEG signals separation using some default robust objective functions.
-The ongoing research in this field focuses on cases where the number of detectors smaller than the number of sources, and non-linear mixing problems.
+Several objective functions have been proposed to improve the robustness and computation efficiency.
+However, if we know in advance what the probability distributions look like for the sources or latent variables, it is always the best to tailor the objective function to include this information.
+Nonetheless, ICA has already shown reliable performance in examples such as sound waves identification and EEG signal separation using some default robust objective functions.
+The ongoing research in this field include cases where the number of detectors is smaller than the number of sources and non-linear mixing problems.
 
